### 2024-05-28 paper/main.tex
Final mag + gcc model

\documentclass{article}

\usepackage[english]{babel}
\usepackage{authblk}
\usepackage[a4paper,top=2cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{gensymb}

\title{Automatic Prediction of Ensemble Width and Location in Binaural Recordings of Music Using Convolutional Neural Network}
\author[1]{Paweł Antoniuk*}
\author[1]{Sławomir Zieliński}
\affil[1]{Faculty of Computer Science, Białystok University of Technology, 15-351 Białystok, Poland; pawel.antoniuk@sd.pb.edu.pl}
\affil[*]{Corresponding author: pawel.antoniuk@sd.pb.edu.pl}

\date{}

\begin{document}
\maketitle

\begin{abstract}
Binaural audio technology has been around for many years, but its popularity has dramatically increased over the past decade due to advancements in virtual reality and streaming technologies. Along with its growing popularity, the quantity of publicly accessible binaural audio materials has also expanded. Consequently, there is now a need for automated and objective measurements of spatial content information, with ensemble width and location being the most important. In this study, 30 head-related transfer functions and 192 binaural music recordings from publicly accessible multi-track recording repositories were used to synthesize 23,040 binaural recordings. The synthesised recordings were then used to train a convolutional neural network prediction model, the aim of which was to predict the width and location of the ensemble for unseen recordings. The results indicate that models of ensemble breadth and position can be successfully constructed with low prediction errors—9.3 degrees for ensemble location and 8.4 degrees for ensemble width. This approach is the first of its kind to predict both ensemble width and location, offering a more accurate representation of spatial properties. This suggests significant potential for advancing spatial audio applications in virtual reality and streaming technologies, by providing audio engineers with tools that can leverage these methods to enhance spatial audio experiences.
\end{abstract}

\section{Introduction}
\section{Related studies}
\section{Methodology}

Experiments in this study were conducted on 2340 binaural recordings of music. The binaural recordings were synthesized semi-automatically using 192 multi-track publicly-available music recordings and 30 HRTF databases. For each multi-track recording and HRTF database pair, four binaural recordings were synthesized for different random ensemble parameters --- its location $\phi$ and width $\theta$ --- as defined in figure X. Both parameters were drawn from uniform random distribution.

\subsection{Ensemble location and width definition}
\subsection{Synthesis of binaural music recordings}
\subsection{Spectrogram conversions of binaural music recordings}

\begin{figure}
\centering
\includegraphics[width=\linewidth]{sample.png}
\caption{\label{fig:sample}Example spectrograms}
\end{figure}

Figure \ref{fig:sample} shows spectrograms for four different binaural recordings of music. The spectrograms were used as the input of the convolutional neural network that used in this study.

\subsection{Convolutional neural network topology}
\subsection{Model training and evaluation}
\section{Results}

\begin{figure}[ht]
  \centering
  \begin{minipage}[t]{0.45\linewidth}
    \includegraphics[width=\linewidth]{actual_vs_predicted_width.pdf}
    \caption{\label{fig:actual_vs_predicted_width}A comparison between the actual and the predicted ensemble width for a single iteration (of the total five) }
  \end{minipage}
  \hspace{0.5cm} 
  \begin{minipage}[t]{0.45\linewidth}
    \includegraphics[width=\linewidth]{mae_width.pdf}
    \caption{\label{fig:mae_width}The impact of the actual ensemble width on the mean absolute prediction error, averaged across all five iterations, with indicated standard deviation.}
  \end{minipage}
\end{figure}

Figure \ref{fig:actual_vs_predicted_width} illustrates the comparison between the actual and the predicted ensemble width. The results demonstrate that the model exhibits optimal prediction quality for narrower ensemble widths ($X\degree$ for $Y\degree$) and that its performance deteriorates with the increase of the ensemble width ($Z\degree$ for $W\degree$). Figure \ref{fig:mae_width} further demonstrates that the relationship between the prediction error and the actual width is not linear, exhibiting a depression between 60 and 75 degrees. This suggests, that the ensemble width has a significant impact on the ensemble width estimation.

\begin{figure}[ht]
  \centering
  \begin{minipage}[t]{0.45\linewidth}
    \includegraphics[width=\linewidth]{actual_vs_predicted_location.pdf}
    \caption{\label{fig:actual_vs_predicted_location}A comparison between the actual and the predicted ensemble location for a single iteration (of the total five) }
  \end{minipage}
  \hspace{0.5cm} 
  \begin{minipage}[t]{0.45\linewidth}
    \includegraphics[width=\linewidth]{mae_location.pdf}
    \caption{\label{fig:mae_location}The impact of the actual ensemble width on the mean absolute prediction error, averaged across all five iterations, with indicated standard deviation.}
  \end{minipage}
\end{figure}

In contrast to the correlation between ensemble width and its prediction error, there is no relationship between the actual location and its prediction error, as illustrated in Figures \ref{fig:actual_vs_predicted_location} and \ref{fig:mae_location}. This finding indicates that the model's capabilities for localizing the center of the ensemble is robust, unaffected by the actual spatial positioning of the ensemble, including lateral locations.

% \begin{figure}
% \centering
% \includegraphics[width=0.5\linewidth]{topology.png}
% \caption{\label{fig:topology}Topology of model}
% \end{figure}

% \begin{figure}
% \centering
% \includegraphics[width=\linewidth]{train-history.png}
% \caption{\label{fig:train-history}History of model training}
% \end{figure}

% \begin{figure}
% \centering
% \includegraphics[width=\linewidth]{width-err.png}
% \caption{\label{fig:width-err}Ensemble width vs. error}
% \end{figure}

\section{Conclusions}
\section{References}

\bibliographystyle{alpha}
\bibliography{references}
\end{document}

% Your introduction goes here! Simply start writing your document and use the Recompile button to view the updated PDF preview. Examples of commonly used commands and features are listed below, to help you get started.

% Once you're familiar with the editor, you can find various project settings in the Overleaf menu, accessed via the button in the very top left of the editor. To view tutorials, user guides, and further documentation, please visit our \href{https://www.overleaf.com/learn}{help library}, or head to our plans page to \href{https://www.overleaf.com/user/subscription/plans}{choose your plan}.

% \section{Some examples to get started}

% \subsection{How to create Sections and Subsections}

% Simply use the section and subsection commands, as in this example document! With Overleaf, all the formatting and numbering is handled automatically according to the template you've chosen. If you're using the Visual Editor, you can also create new section and subsections via the buttons in the editor toolbar.

% \subsection{How to include Figures}

% First you have to upload the image file from your computer using the upload link in the file-tree menu. Then use the includegraphics command to include it in your document. Use the figure environment and the caption command to add a number and a caption to your figure. See the code for Figure \ref{fig:frog} in this section for an example.

% Note that your figure will automatically be placed in the most appropriate place for it, given the surrounding text and taking into account other figures or tables that may be close by. You can find out more about adding images to your documents in this help article on \href{https://www.overleaf.com/learn/how-to/Including_images_on_Overleaf}{including images on Overleaf}.

% \begin{figure}
% \centering
% \includegraphics[width=0.25\linewidth]{frog.jpg}
% \caption{\label{fig:frog}This frog was uploaded via the file-tree menu.}
% \end{figure}

% \subsection{How to add Tables}

% Use the table and tabular environments for basic tables --- see Table~\ref{tab:widgets}, for example. For more information, please see this help article on \href{https://www.overleaf.com/learn/latex/tables}{tables}.

% \begin{table}
% \centering
% \begin{tabular}{l|r}
% Item & Quantity \\\hline
% Widgets & 42 \\
% Gadgets & 13
% \end{tabular}
% \caption{\label{tab:widgets}An example table.}
% \end{table}

% \subsection{How to add Comments and Track Changes}

% Comments can be added to your project by highlighting some text and clicking ``Add comment'' in the top right of the editor pane. To view existing comments, click on the Review menu in the toolbar above. To reply to a comment, click on the Reply button in the lower right corner of the comment. You can close the Review pane by clicking its name on the toolbar when you're done reviewing for the time being.

% Track changes are available on all our \href{https://www.overleaf.com/user/subscription/plans}{premium plans}, and can be toggled on or off using the option at the top of the Review pane. Track changes allow you to keep track of every change made to the document, along with the person making the change.

% \subsection{How to add Lists}

% You can make lists with automatic numbering \dots

% \begin{enumerate}
% \item Like this,
% \item and like this.
% \end{enumerate}
% \dots or bullet points \dots
% \begin{itemize}
% \item Like this,
% \item and like this.
% \end{itemize}

% \subsection{How to write Mathematics}

% \LaTeX{} is great at typesetting mathematics. Let $X_1, X_2, \ldots, X_n$ be a sequence of independent and identically distributed random variables with $\text{E}[X_i] = \mu$ and $\text{Var}[X_i] = \sigma^2 < \infty$, and let
% \[S_n = \frac{X_1 + X_2 + \cdots + X_n}{n}
%       = \frac{1}{n}\sum_{i}^{n} X_i\]
% denote their mean. Then as $n$ approaches infinity, the random variables $\sqrt{n}(S_n - \mu)$ converge in distribution to a normal $\mathcal{N}(0, \sigma^2)$.

% \subsection{How to change the margins and paper size}

% Usually the template you're using will have the page margins and paper size set correctly for that use-case. For example, if you're using a journal article template provided by the journal publisher, that template will be formatted according to their requirements. In these cases, it's best not to alter the margins directly.

% If however you're using a more general template, such as this one, and would like to alter the margins, a common way to do so is via the geometry package. You can find the geometry package loaded in the preamble at the top of this example file, and if you'd like to learn more about how to adjust the settings, please visit this help article on \href{https://www.overleaf.com/learn/latex/page_size_and_margins}{page size and margins}.

% \subsection{How to change the document language and spell check settings}

% Overleaf supports many different languages, including multiple different languages within one document.

% To configure the document language, simply edit the option provided to the babel package in the preamble at the top of this example project. To learn more about the different options, please visit this help article on \href{https://www.overleaf.com/learn/latex/International_language_support}{international language support}.

% To change the spell check language, simply open the Overleaf menu at the top left of the editor window, scroll down to the spell check setting, and adjust accordingly.

% \subsection{How to add Citations and a References List}

% You can simply upload a \verb|.bib| file containing your BibTeX entries, created with a tool such as JabRef. You can then cite entries from it, like this: \cite{greenwade93}. Just remember to specify a bibliography style, as well as the filename of the \verb|.bib|. You can find a \href{https://www.overleaf.com/help/97-how-to-include-a-bibliography-using-bibtex}{video tutorial here} to learn more about BibTeX.

% If you have an \href{https://www.overleaf.com/user/subscription/plans}{upgraded account}, you can also import your Mendeley or Zotero library directly as a \verb|.bib| file, via the upload menu in the file-tree.

% \subsection{Good luck!}

% We hope you find Overleaf useful, and do take a look at our \href{https://www.overleaf.com/learn}{help library} for more tutorials and user guides! Please also let us know if you have any feedback using the Contact Us link at the bottom of the Overleaf menu --- or use the contact form at \url{https://www.overleaf.com/contact}.

% \bibliographystyle{alpha}
% \bibliography{sample}

% \end{document}

### 2024-05-28 analyze_plot.py

@@ -11,69 +11,105 @@ import tensorflow as tf
 
 from model import Model
 
+model_name = "multi_inout_1_model_final_2"
+model_path = Path("models") / model_name
+out_img_format = "pdf"
+
 
 def load_results():
     prediction_results = []
-    src_model_path = Path('models') / 'simple2'
-    for path_rep_it in src_model_path.glob('*'):
-        request = joblib.load(path_rep_it / 'request.pkl')
+    src_model_path = model_path
+    for path_rep_it in src_model_path.glob("*"):
+        request = joblib.load(path_rep_it / "request.pkl")
         model = Model(request).try_load_or_compute_input_data()
         prediction_result = model.predict_test_data()
-        errors = prediction_result.predicted_width - prediction_result.actual_width
-        score_test = np.mean(np.abs(errors))
+        errors_width = (
+            prediction_result.predicted_width - prediction_result.actual_width
+        )
+        errors_location = (
+            prediction_result.predicted_location - prediction_result.actual_location
+        )
+        score_test_width = np.mean(np.abs(errors_width))
+        score_test_location = np.mean(np.abs(errors_location))
 
         # assertions
-        score_path = next(path_rep_it.glob('score_*.csv'))
+        score_path = next(path_rep_it.glob("score_*.csv"))
         score_df = pd.read_csv(score_path)
-        score_file_test = score_df['test_score']
-        assert np.all(np.abs(score_file_test - score_test) < 1e-6)
+        score_file_test_width = score_df["test_score_width"]
+        score_file_test_location = score_df["test_score_location"]
+        assert np.all(np.abs(score_file_test_width - score_test_width) < 1e-6)
+        assert np.all(np.abs(score_file_test_location - score_test_location) < 1e-6)
 
         test_filenames = prediction_result.filename
-        actual_widths_test = [float(x.split('_')[-4].replace('width', '')) for x in test_filenames]
-        stored_actual_widths_test = np.array(model.store_split.y_test * 45.0)
-        assert np.all(abs(stored_actual_widths_test - actual_widths_test) < 1e-6)
+        actual_widths_test = [
+            float(x.split("_")[-4].replace("width", "")) for x in test_filenames
+        ]
[…]

### 2024-05-28 architecture/multi_inout_1.py

@@ -1,0 +1,40 @@
+from keras import layers as l
+from keras import models as m
+
+
+def cnn_block(input_layer):
+    x = l.Conv2D(32, kernel_size=(2, 2), activation='relu', padding='same')(input_layer)
+    x = l.MaxPooling2D(pool_size=(2, 3))(x)
+    x = l.Conv2D(64, kernel_size=(2, 2), activation='relu', padding='same')(x)
+    x = l.MaxPooling2D(pool_size=(2, 3))(x)
+    x = l.Conv2D(128, kernel_size=(2, 2), activation='relu', padding='same')(x)
+    x = l.MaxPooling2D(pool_size=(2, 2))(x)
+    x = l.Conv2D(256, kernel_size=(2, 2), activation='relu', padding='same')(x)
+    x = l.MaxPooling2D(pool_size=(2, 2))(x)
+    x = l.GlobalAveragePooling2D()(x)
+    return x
+
+
+def gcc_phat_block(input_layer):
+    x = l.Dense(32, activation='relu')(input_layer)
+    return x
+
+
+def architecture(input_shape):
+    mag_input_layer = l.Input(shape=input_shape, name='in_mag')
+    # phase_input_layer = l.Input(shape=input_shape, name='in_phase')
+    # gcc_input_layer = l.Input(shape=(66,), name='in_gcc')
+
+    mag_model = cnn_block(mag_input_layer)
+    # phase_model = cnn_block(phase_input_layer)
+    # gcc_model = gcc_phat_block(gcc_input_layer)
+    merged = l.concatenate([mag_model])
+
+    x = l.Dense(64, activation='relu')(merged)
+    x = l.Dense(32, activation='relu')(x)
+    x = l.Dense(16, activation='relu')(x)
+
+    output_width = l.Dense(1, name='out_width')(x)
+    output_location = l.Dense(1, name='out_location')(x)
+
+    return m.Model(inputs=[mag_input_layer], outputs=[output_width, output_location])

### 2024-05-28 gcc_phat.py

@@ -1,0 +1,72 @@
+import numpy as np
+import numpy.fft as fft
+import matplotlib.pyplot as plt
+import soundfile as sf
+import os
+import random
+
+
+def gcc_phat(x1: np.ndarray, x2: np.ndarray, fs=1, interp=1):
+    n = x1.shape[0] + x2.shape[0] - 1
+    if n % 2 != 0:
+        n += 1
+    X1 = fft.rfft(x1, n)
+    X2 = fft.rfft(x2, n)
+    X1 /= np.abs(X1)
+    X2 /= np.abs(X2)
+    cc = fft.irfft(X1 * np.conj(X2), n=interp * n)
+    t_max = n // 2 + 1
+    cc = np.concatenate((cc[-t_max:], cc[:t_max]))
+    tau = np.argmax(np.abs(cc))
+    tau -= t_max
+    return tau / (fs * interp), cc
+
+
+def gcc_phat_feature(xy: np.ndarray, fs: int, len: float) -> np.ndarray:
+    _, cc = gcc_phat(xy[:, 0], xy[:, 1], fs=fs)
+    c = cc.shape[0]//2
+    o = int(len * fs)
+    ccs = cc[c-o:c+o]
+    assert ccs.shape[0] == gcc_phat_feature_nsamples(fs, len)
+    return ccs
+
+
+def gcc_phat_feature_nsamples(fs: int, len: float) -> int:
+    o = int(len * fs)
+    return int(o * 2)
+
+
+def plot_gcc_phat_feature(filename: str, base_path: str, len: float = 0.0007):
+    full_filename = os.path.join(base_path, filename)
+    xy, fs = sf.read(full_filename)
+    fvec = gcc_phat_feature(xy, fs)
+    num_samples = len(fvec)
+    time_lags = np.linspace(-num_samples // 2,
+                            num_samples // 2, num_samples) / fs * 1e6
+
+    plt.figure(figsize=(6, 4))
+    plt.plot(time_lags, fvec)
+    plt.title(filename)
+    plt.xlabel('Time Lag [µs]')
+    plt.ylabel('Normalized Correlation Values')
+    plt.xlim(-700, 700)
+    plt.ylim(-1, 1)
+    plt.grid(True)
+    plt.savefig(f'gcc_phat/gcc_phat_{filename}.png')
+    plt.close()
+
+
+if __name__ == "__main__":
+    base_path = '/run/media/pawel/alpha/spatresults/spat'
+
+    # List all files in the directory
+    all_files = os.listdir(base_path)
+
[…]

### 2024-05-28 model.py

@@ -24,8 +24,9 @@ from dotmap import DotMap
 from scipy.signal import spectrogram
 from scipy.signal.windows import hamming
 from sklearn.model_selection import GroupShuffleSplit
+from keras import mixed_precision
 
-from gcc_phat import gcc_phat_feature, gcc_phat_feature_len
+from gcc_phat import gcc_phat_feature, gcc_phat_feature_nsamples
 
 
 @dataclass
@@ -48,20 +49,22 @@ class Request:
     train_batch_size: int = 8
     train_epochs: int = 128
     train_patience: int = 15
-    train_loss: str = "mean_squared_error"
+    train_loss: str = "mae"
     dev_test_ratio: float = 1 / 3
     train_val_ratio: float = 1 / 8
     split_seed: int = 2
     train_seed: int = 3
     repetition_iteration: int = 1
     spectrogram_type: str = "MAGNITUDE"
+    dtype: str = "float32"
+    gcc_phat_len: float = 0.0007
 
 
 class Model:
     def __init__(self, request: Request):
         self.result_best_val_mae: Optional[float] = None
         self.result_best_epoch: Optional[int] = None
-        self.result_test_mae: Optional[float] = None
+        self.result_test_mae: Optional[dict] = None
         self.model_n_params: Optional[int] = None
         self.result_history: Optional[keras.callbacks.History] = None
         self.time_extract: Optional[float] = None
@@ -86,8 +89,7 @@ class Model:
         self.input_dir = Path(request.input_dir)
         self.num_samples = request.num_samples
         self.sample_rate = request.sample_rate
-        self.filepaths = [p for p in self.input_dir.iterdir()
-                          ][0: self.num_samples]
+        self.filepaths = [p for p in self.input_dir.iterdir()][0 : self.num_samples]
         self.target_bands = request.target_bands
         self.time_window_len = request.time_window_len
         self.time_window_overlap = request.time_window_overlap
@@ -95,20 +97,21 @@ class Model:
         self.spectrogram_max_freq = request.spectrogram_max_freq
         self.recording_time_length = request.recording_time_length
         self.n_time_frames = int(
-            self.recording_time_length / self.time_window_len * 2 - 1
+            (self.recording_time_length - self.time_window_len)
+            / (self.time_window_len * (1 - self.time_window_overlap))
+            + 1
         )
         self.model_architecture = request.model_architecture
         self.train_batch_size = request.train_batch_size
         self.train_epochs = request.train_epochs
[…]

### 2024-05-28 run.py

@@ -19,7 +19,7 @@ def execute_request(evaluation_request, iteration):
         evaluation_request.split_seed = iteration
         evaluation_request.train_seed = iteration + 1
         m = Model(Request(**evaluation_request))
-        m.try_load_or_compute_input_data().plot_features().train().save_results()
+        m.try_load_or_compute_input_data().train().save_results()
     except Exception as e:
         print(f'[{evaluation_request.name}] Error while executing request: {e}')
         traceback.print_exc()

### 2024-06-04 paper/main.tex
Refactor repo structure

@@
 \usepackage{gensymb}
+\usepackage{textcomp}
 
-\title{Automatic Prediction of Ensemble Width and Location in Binaural Recordings of Music Using Convolutional Neural Network}
+\title{Predicting Ensemble Width and Location in Binaural Recordings of Music with Convolutional Neural Networks}
 \author[1]{Paweł Antoniuk*}
@@
 The synthesised recordings were then used to train a convolutional neural network prediction model, the aim of which was to predict the width and location of the ensemble for unseen recordings.
-The results indicate that models of ensemble breadth and position can be successfully constructed with low prediction errors—9.3 degrees for ensemble location and 8.4 degrees for ensemble width.
+The results indicate that models of ensemble breadth and position can be successfully constructed with low prediction errors --- $4.62\degree$ ($\pm0.09\degree$) degrees for ensemble location and $8.63\degree$ ($\pm0.29\degree$) degrees for ensemble width.
 This approach is the first of its kind to predict both ensemble width and location, offering a more accurate representation of spatial properties.
@@
 \section{Related studies}
+
 \section{Methodology}
@@
 The binaural recordings were synthesized semi-automatically using 192 multi-track publicly-available music recordings and 30 HRTF databases.
-For each multi-track recording and HRTF database pair, four binaural recordings were synthesized for different random ensemble parameters --- its location $\phi$ and width $\theta$ --- as defined in figure X.
+For each multi-track recording and HRTF database pair, four binaural recordings were synthesized for different random ensemble parameters --- its location $\phi$ and width $\theta$ --- as defined in Section \ref{subsec:ensemble-definition}.
 Both parameters were drawn from uniform random distribution.
@@
 \subsection{Ensemble location and width definition}
+\label{subsec:ensemble-definition}
+
+The primary objective of the model developed in this study is to predict the ensemble location ($\theta$) and width ($\omega$), as illustrated in Figure \ref{fig:scene}.
+The ensemble is defined as a group of audio point sources.
+The ensemble width $\omega$ is defined as the angular width between two extreme point sources, while the ensemble location $\theta$ represents the middle angle between two extreme sound sources.
+
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{../pictures/scene.png}
+\caption{\label{fig:scene}Illustration of ensemble width ($\omega$) and ensemble location ($\theta$) relative to the direction of the head.
+Black dots represent the positions of audio sources.
+The ensemble location ($\theta$) is defined as the angular position of the center of the ensemble relative to the direction the head is turned.
+The ensemble width ($\omega$) is defined as angular distance between two extreme audio sources.
+}
+\end{figure}
+
 \subsection{Synthesis of binaural music recordings}
@@
 \centering
-\includegraphics[width=\linewidth]{sample.png}
-\caption{\label{fig:sample}Example spectrograms}
+\includegraphics[width=\linewidth]{../figures/samples.pdf}
+\caption{\label{fig:samples}An illustrative example of the magnitude spectrograms generated from synthesized binaural recordings of music.
+Each spectrogram was derived from a different randomly selected binaural music recording.
+These spectrograms were subsequently employed as input data for training and assessing the convolutional neural network (CNN) in this study.
+}
 \end{figure}
 
-Figure \ref{fig:sample} shows spectrograms for four different binaural recordings of music. The spectrograms were used as the input of the convolutional neural network that used in this study.
+Prior to being fed into the model, the binaural recordings of music were transformed into magnitude spectrograms.
+Figure \ref{fig:samples} illustrates example spectrograms for four distinct binaural recordings of music.
+It should be noted that the original spectrograms were stored in a floating-point-precision matrix format.
 
@@
   \begin{minipage}[t]{0.45\linewidth}
-    \includegraphics[width=\linewidth]{actual_vs_predicted_width.pdf}
+    \includegraphics[width=\linewidth]{../figures/actual_vs_predicted_width.pdf}
     \caption{\label{fig:actual_vs_predicted_width}A comparison between the actual and the predicted ensemble width for a single iteration (of the total five) }
@@
   \begin{minipage}[t]{0.45\linewidth}
-    \includegraphics[width=\linewidth]{mae_width.pdf}
+    \includegraphics[width=\linewidth]{../figures/mae_width.pdf}
     \caption{\label{fig:mae_width}The impact of the actual ensemble width on the mean absolute prediction error, averaged across all five iterations, with indicated standard deviation.}
@@
   \begin{minipage}[t]{0.45\linewidth}
-    \includegraphics[width=\linewidth]{actual_vs_predicted_location.pdf}
+    \includegraphics[width=\linewidth]{../figures/actual_vs_predicted_location.pdf}
     \caption{\label{fig:actual_vs_predicted_location}A comparison between the actual and the predicted ensemble location for a single iteration (of the total five) }
@@
   \begin{minipage}[t]{0.45\linewidth}
-    \includegraphics[width=\linewidth]{mae_location.pdf}
+    \includegraphics[width=\linewidth]{../figures/mae_location.pdf}
     \caption{\label{fig:mae_location}The impact of the actual ensemble width on the mean absolute prediction error, averaged across all five iterations, with indicated standard deviation.}
@@
 
-In contrast to the correlation between ensemble width and its prediction error, there is no relationship between the actual location and its prediction error, as illustrated in Figures \ref{fig:actual_vs_predicted_location} and \ref{fig:mae_location}.
+In contrast to the correlation between ensemble width and its prediction error, there is no significant relationship between the actual location and its prediction error, as illustrated in Figures \ref{fig:actual_vs_predicted_location} and \ref{fig:mae_location}.
 This finding indicates that the model's capabilities for localizing the center of the ensemble is robust, unaffected by the actual spatial positioning of the ensemble, including lateral locations.
 
-% \begin{figure}
-% \centering
-% \includegraphics[width=0.5\linewidth]{topology.png}
-% \caption{\label{fig:topology}Topology of model}
-% \end{figure}
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{../figures/map_mae_location.pdf}
+    \caption{\label{fig:map_mae_location}The heatmap that illustrates the mean absolute error (MAE) of ensemble location distribution across different ensemble locations (x-axis) and ensemble widths (y-axis).
+The color intensity corresponds to the MAE values, with lighter areas indicating higher errors.}
+\end{figure}
 
-% \begin{figure}
-% \centering
-% \includegraphics[width=\linewidth]{train-history.png}
-% \caption{\label{fig:train-history}History of model training}
-% \end{figure}
+Figure \ref{fig:map_mae_location} shows influence of both the ensemble width and location on mean absolute error for ensemble location.
+This shows more detailed view of data previously presented on Figure \ref{fig:mae_location}.
+The figure shows that there is a slight asymmetry.
 
-% \begin{figure}
-% \centering
-% \includegraphics[width=\linewidth]{width-err.png}
-% \caption{\label{fig:width-err}Ensemble width vs. error}
-% \end{figure}
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{../figures/map_mae_width.pdf}
+    \caption{\label{fig:map_mae_width}The heatmap that illustrates the mean absolute error (MAE) of ensemble width distribution across different ensemble locations (x-axis) and ensemble widths (y-axis).
+The color intensity corresponds to the MAE values, with lighter areas indicating higher errors.
+Notably, the values between $30\degree$ and $60\degree$ on the y-axis exhibit unexpectedly higher MAE values region --- please see Figure \ref{fig:mae_width} for comparison.
+}
+\end{figure}
+
+Similarly, Figure \ref{fig:map_mae_width} shows influence of both the ensemble width and location on mean absolute error for ensemble width.
 
@@
 
+\cite{greenwade93}
+
 \bibliographystyle{alpha}
@@
 \end{document}
-
-% Your introduction goes here!
-Simply start writing your document and use the Recompile button to view the updated PDF preview.
-Examples of commonly used commands and features are listed below, to help you get started.
-
-% Once you're familiar with the editor, you can find various project settings in the Overleaf menu, accessed via the button in the very top left of the editor.
-To view tutorials, user guides, and further documentation, please visit our \href{https://www.overleaf.com/learn}{help library}, or head to our plans page to \href{https://www.overleaf.com/user/subscription/plans}{choose your plan}.
-
-% \section{Some examples to get started}
-
-% \subsection{How to create Sections and Subsections}
-
-% Simply use the section and subsection commands, as in this example document!
-With Overleaf, all the formatting and numbering is handled automatically according to the template you've chosen.
-If you're using the Visual Editor, you can also create new section and subsections via the buttons in the editor toolbar.
-
-% \subsection{How to include Figures}
-
-% First you have to upload the image file from your computer using the upload link in the file-tree menu.
-Then use the includegraphics command to include it in your document.
-Use the figure environment and the caption command to add a number and a caption to your figure.
-See the code for Figure \ref{fig:frog} in this section for an example.
-
-% Note that your figure will automatically be placed in the most appropriate place for it, given the surrounding text and taking into account other figures or tables that may be close by.
-You can find out more about adding images to your documents in this help article on \href{https://www.overleaf.com/learn/how-to/Including_images_on_Overleaf}{including images on Overleaf}.
-
-% \begin{figure}
-% \centering
-% \includegraphics[width=0.25\linewidth]{frog.jpg}
-% \caption{\label{fig:frog}This frog was uploaded via the file-tree menu.}
-% \end{figure}
-
-% \subsection{How to add Tables}
-
-% Use the table and tabular environments for basic tables --- see Table~\ref{tab:widgets}, for example.
-For more information, please see this help article on \href{https://www.overleaf.com/learn/latex/tables}{tables}.
-
-% \begin{table}
-% \centering
-% \begin{tabular}{l|r}
-% Item & Quantity \\\hline
-% Widgets & 42 \\
-% Gadgets & 13
-% \end{tabular}
-% \caption{\label{tab:widgets}An example table.}
-% \end{table}
-
-% \subsection{How to add Comments and Track Changes}
-
-% Comments can be added to your project by highlighting some text and clicking ``Add comment'' in the top right of the editor pane.
-To view existing comments, click on the Review menu in the toolbar above.
-To reply to a comment, click on the Reply button in the lower right corner of the comment.
-You can close the Review pane by clicking its name on the toolbar when you're done reviewing for the time being.
-
-% Track changes are available on all our \href{https://www.overleaf.com/user/subscription/plans}{premium plans}, and can be toggled on or off using the option at the top of the Review pane.
-Track changes allow you to keep track of every change made to the document, along with the person making the change.
-
-% \subsection{How to add Lists}
-
-% You can make lists with automatic numbering \dots
-
-% \begin{enumerate}
-% \item Like this,
-% \item and like this.
-% \end{enumerate}
-% \dots or bullet points \dots
-% \begin{itemize}
-% \item Like this,
-% \item and like this.
-% \end{itemize}
-
-% \subsection{How to write Mathematics}
-
-% \LaTeX{} is great at typesetting mathematics.
-Let $X_1, X_2, \ldots, X_n$ be a sequence of independent and identically distributed random variables with $\text{E}[X_i] = \mu$ and $\text{Var}[X_i] = \sigma^2 < \infty$, and let
-% \[S_n = \frac{X_1 + X_2 + \cdots + X_n}{n}
-%       = \frac{1}{n}\sum_{i}^{n} X_i\]
-% denote their mean. Then as $n$ approaches infinity, the random variables $\sqrt{n}(S_n - \mu)$ converge in distribution to a normal $\mathcal{N}(0, \sigma^2)$.
-
-% \subsection{How to change the margins and paper size}
-
-% Usually the template you're using will have the page margins and paper size set correctly for that use-case.
-For example, if you're using a journal article template provided by the journal publisher, that template will be formatted according to their requirements.
-In these cases, it's best not to alter the margins directly.
-
-% If however you're using a more general template, such as this one, and would like to alter the margins, a common way to do so is via the geometry package.
-You can find the geometry package loaded in the preamble at the top of this example file, and if you'd like to learn more about how to adjust the settings, please visit this help article on \href{https://www.overleaf.com/learn/latex/page_size_and_margins}{page size and margins}.
-
-% \subsection{How to change the document language and spell check settings}
-
-% Overleaf supports many different languages, including multiple different languages within one document.
-
-% To configure the document language, simply edit the option provided to the babel package in the preamble at the top of this example project.
-To learn more about the different options, please visit this help article on \href{https://www.overleaf.com/learn/latex/International_language_support}{international language support}.
-
-% To change the spell check language, simply open the Overleaf menu at the top left of the editor window, scroll down to the spell check setting, and adjust accordingly.
-
-% \subsection{How to add Citations and a References List}
-
-% You can simply upload a \verb|.bib| file containing your BibTeX entries, created with a tool such as JabRef.
-You can then cite entries from it, like this: \cite{greenwade93}.
-Just remember to specify a bibliography style, as well as the filename of the \verb|.bib|.
-You can find a \href{https://www.overleaf.com/help/97-how-to-include-a-bibliography-using-bibtex}{video tutorial here} to learn more about BibTeX.
-
-% If you have an \href{https://www.overleaf.com/user/subscription/plans}{upgraded account}, you can also import your Mendeley or Zotero library directly as a \verb|.bib| file, via the upload menu in the file-tree.
-
-% \subsection{Good luck!}
-
-% We hope you find Overleaf useful, and do take a look at our \href{https://www.overleaf.com/learn}{help library} for more tutorials and user guides!
-Please also let us know if you have any feedback using the Contact Us link at the bottom of the Overleaf menu --- or use the contact form at \url{https://www.overleaf.com/contact}.
-
-% \bibliographystyle{alpha}
-% \bibliography{sample}
-
-% \end{document}

### 2024-06-04 src/analyze/analyze_plot.py

@@ -1,4 +1,5 @@
 import os
+import sys
 from pathlib import Path
 
 import joblib
@@ -8,18 +9,19 @@ from dotmap import DotMap
 from matplotlib import pyplot as plt
 from matplotlib.ticker import FuncFormatter
 import tensorflow as tf
+import colormaps
+import matplotlib.colors as mcolors
 
 from model import Model
 
-model_name = "multi_inout_1_model_final_2"
-model_path = Path("models") / model_name
+model_name = "multi_inout_1_model_final_3_nogcc"
+model_dir = Path("/data/train/out_models") / model_name
 out_img_format = "pdf"
 
 
 def load_results():
     prediction_results = []
-    src_model_path = model_path
-    for path_rep_it in src_model_path.glob("*"):
+    for path_rep_it in model_dir.glob("*"):
         request = joblib.load(path_rep_it / "request.pkl")
         model = Model(request).try_load_or_compute_input_data()
         prediction_result = model.predict_test_data()
@@ -38,20 +40,24 @@ def load_results():
         score_file_test_width = score_df["test_score_width"]
         score_file_test_location = score_df["test_score_location"]
         assert np.all(np.abs(score_file_test_width - score_test_width) < 1e-6)
-        assert np.all(np.abs(score_file_test_location - score_test_location) < 1e-6)
+        assert np.all(np.abs(score_file_test_location -
+                      score_test_location) < 1e-6)
 
         test_filenames = prediction_result.filename
         actual_widths_test = [
             float(x.split("_")[-4].replace("width", "")) for x in test_filenames
         ]
-        stored_actual_widths_test = np.array(model.store_split.y_test_width * 45.0)
-        assert np.all(abs(stored_actual_widths_test - actual_widths_test) < 1e-5)
+        stored_actual_widths_test = np.array(
+            model.store_split.y_test_width * 45.0)
+        assert np.all(abs(stored_actual_widths_test -
+                      actual_widths_test) < 1e-5)
 
         actual_recordings = [
             os.path.basename(x.split("_")[-7].replace("width", ""))
             for x in test_filenames
         ]
-        assert np.all(model.store_split.actual_recordings_test == actual_recordings)
[…]

### 2024-06-04 src/train/architecture/multi_inout_1.py

@@ -15,25 +15,18 @@ def cnn_block(input_layer):
     return x
 
 
-def gcc_phat_block(input_layer):
-    x = l.Dense(32, activation='relu')(input_layer)
-    return x
-
-
 def architecture(input_shape):
+    # input
     mag_input_layer = l.Input(shape=input_shape, name='in_mag')
-    # phase_input_layer = l.Input(shape=input_shape, name='in_phase')
-    # gcc_input_layer = l.Input(shape=(66,), name='in_gcc')
-
-    mag_model = cnn_block(mag_input_layer)
-    # phase_model = cnn_block(phase_input_layer)
-    # gcc_model = gcc_phat_block(gcc_input_layer)
-    merged = l.concatenate([mag_model])
-
-    x = l.Dense(64, activation='relu')(merged)
+    
+    # inner layers
+    mag_model = cnn_block(mag_input_layer)    
+    x = l.Dense(128, activation='relu')(mag_model)
+    x = l.Dense(64, activation='relu')(x)
     x = l.Dense(32, activation='relu')(x)
     x = l.Dense(16, activation='relu')(x)
-
+    
+    # output
     output_width = l.Dense(1, name='out_width')(x)
     output_location = l.Dense(1, name='out_location')(x)
 

### 2024-06-04 src/train/gcc_phat.py

@@ -55,18 +55,3 @@ def plot_gcc_phat_feature(filename: str, base_path: str, len: float = 0.0007):
     plt.savefig(f'gcc_phat/gcc_phat_{filename}.png')
     plt.close()
 
-
-if __name__ == "__main__":
-    base_path = '/run/media/pawel/alpha/spatresults/spat'
-
-    # List all files in the directory
-    all_files = os.listdir(base_path)
-
-    # Filter to get only .wav files
-    wav_files = [f for f in all_files if f.endswith('.wav')]
-
-    # Select 5 random files
-    random_files = random.sample(wav_files, 10)
-
-    for filename in random_files:
-        plot_gcc_phat_feature(filename, base_path)

### 2024-06-04 src/train/model.py

@@ -33,6 +33,7 @@ from gcc_phat import gcc_phat_feature, gcc_phat_feature_nsamples
 class Request:
     name: str = "unknown_model"
     input_dir: str = "input"
+    memmap_dir: str = "memmap"
     num_samples: int = 23040
     sample_rate: int = 48000
     target_bands: int = 300
@@ -87,9 +88,11 @@ class Model:
         self.request: Request = request
         self.name = request.name
         self.input_dir = Path(request.input_dir)
+        self.memmap_dir = Path(request.memmap_dir)
         self.num_samples = request.num_samples
         self.sample_rate = request.sample_rate
-        self.filepaths = [p for p in self.input_dir.iterdir()][0 : self.num_samples]
+        self.filepaths = [p for p in self.input_dir.iterdir()
+                          ][0: self.num_samples]
         self.target_bands = request.target_bands
         self.time_window_len = request.time_window_len
         self.time_window_overlap = request.time_window_overlap
@@ -106,6 +109,10 @@ class Model:
         self.train_epochs = request.train_epochs
         self.train_patience = request.train_patience
         self.train_loss = request.train_loss
+        # self.train_learn_schedule = keras.optimizers.schedules.ExponentialDecay(
+        #     initial_learning_rate=request.learn_rate,
+        #     decay_steps=10000,
+        #     decay_rate=request.learn_decay)
         self.train_optimizer = keras.optimizers.Adam(
             learning_rate=request.learn_rate,
             beta_1=request.learn_b1,
@@ -154,12 +161,17 @@ class Model:
 
         print("[extract] Initializing memory")
         self.__initialize_extract_memory()
-        self.store_extract.actual_widths = np.zeros(self.num_samples, dtype=self.dtype)
-        self.store_extract.actual_locations = np.zeros(self.num_samples, dtype=self.dtype)
-        self.store_extract.actual_recordings = ["" for _ in range(self.num_samples)]
-        self.store_extract.actual_filenames = ["" for _ in range(self.num_samples)]
+        self.store_extract.actual_widths = np.zeros(
+            self.num_samples, dtype=self.dtype)
+        self.store_extract.actual_locations = np.zeros(
+            self.num_samples, dtype=self.dtype)
+        self.store_extract.actual_recordings = [
+            "" for _ in range(self.num_samples)]
[…]

### 2024-06-04 src/train/run.py

@@ -7,7 +7,7 @@ from dotmap import DotMap
 
 from model import Model, Request
 
-with open('requests.yaml') as f:
+with open('/src/train/requests.yaml') as f:
     data = yaml.safe_load(f)
     settings = DotMap(data, _dynamic=False)
 

### 2024-06-04 test.py

@@ -1,3 +1,0 @@
-from model import Model
-m = Model()
-m.load_split().train()

### 2024-06-07 paper/main.tex
Update repository structure

@@
 \usepackage{textcomp}
+\usepackage[backend=biber, style=nature]{biblatex}
+\usepackage{csquotes}
+
+\addbibresource{references.bib}
 
 \title{Predicting Ensemble Width and Location in Binaural Recordings of Music with Convolutional Neural Networks}
-\author[1]{Paweł Antoniuk*}
+\author[1,*]{Paweł Antoniuk}
 \author[1]{Sławomir Zieliński}
-\affil[1]{Faculty of Computer Science, Białystok University of Technology, 15-351 Białystok, Poland; pawel.antoniuk@sd.pb.edu.pl}
+\affil[1]{Faculty of Computer Science, Białystok University of Technology}
 \affil[*]{Corresponding author: pawel.antoniuk@sd.pb.edu.pl}
-
 \date{}
@@
 In this study, 30 head-related transfer functions and 192 binaural music recordings from publicly accessible multi-track recording repositories were used to synthesize 23,040 binaural recordings.
-The synthesised recordings were then used to train a convolutional neural network prediction model, the aim of which was to predict the width and location of the ensemble for unseen recordings.
-The results indicate that models of ensemble breadth and position can be successfully constructed with low prediction errors --- $4.62\degree$ ($\pm0.09\degree$) degrees for ensemble location and $8.63\degree$ ($\pm0.29\degree$) degrees for ensemble width.
+The synthesized recordings were then used to train a convolutional neural network prediction model, the aim of which was to predict the width and location of the ensemble for unseen recordings.
+The results indicate that models of ensemble breadth and position can be successfully constructed with low prediction errors --- $4.76\degree$ ($\pm0.10\degree$) for ensemble location and $8.57\degree$ ($\pm0.19\degree$) for ensemble width.
 This approach is the first of its kind to predict both ensemble width and location, offering a more accurate representation of spatial properties.
@@
 \section{Introduction}
+
+The human auditory system demonstrates exceptional proficiency in segregating, localizing, and interpreting diverse auditory signals, despite being limited to two ears.
+This capability arises from the intricate analysis of temporal, amplitude, and spectral disparities, a process termed binaural hearing \cite{blauert_spatial_1996}.
+This enables precise localization of sound sources in complex auditory environments.
+A particularly useful feature of binaural hearing is the 'cocktail party effect', which exemplifies the system's ability to focus on specific sounds while suppressing background noise \cite{cherry_experiments_1953}.
+An understanding of the auditory system is essential for an enhanced comprehension of its limits, the creation of more immersive binaural experiences for entertainment purposes \cite{zhang_surround_2017}, and the improvement of hearing aid devices to enhance auditory signal reception and improve spatial awareness \cite{thiemann_speech_2016}.
+
+The advancement of sophisticated machine learning techniques, particularly deep learning networks, has prompted an intriguing exploration of the extent to which these tools can emulate the human auditory system without relying on advanced spatial audio feature engineering, traditionally employed in audio source localization tasks \cite{yang_deepear_2022, vera-diaz_towards_2018, pang_multitask_2019}.
+To investigate this, the study developed an audio localization method based on a multi-task convolutional neural network (CNN) model.
+
+Inspired by the fact, that humans tend to localize sound sources in groups rather than individually \cite{bregman_auditory_1990, rumsey_spatial_2002}, the objective of the proposed model is to predict ensemble location and width instead of positions of individual sources.
+This study is unique as it not only conceptualized the method but also tested it on a vast, real-life music corpus of 23,040 binaural excerpts that were synthesized using 192 multi-track music recordings and 30 publicly available head-related transfer functions (HRTFs) from various sources.
+The music recordings encompassed various genres, including rock, jazz, pop, and classical music.
+
+The developed method has the potential to be highly beneficial in automated assessment tasks, where a significant number of binaural recordings must be evaluated and labeled in terms of their spatial content information.
+This may assist audio engineers in objectively assessing and segregating binaural audio recordings with regard to their spatial content.
+Furthermore, it could facilitate the development of an autonomous web-crawler bot that will collect binaural recordings from publicly accessible repositories and label them according to the spatial properties of the sound sources, such as the location of the music ensemble or the sparsity of audio source positions.
+
+The findings demonstrate that this method is effective in accurately predicting the spatial characteristics of sound sources in near-real-world scenarios.
+One of the key advantages of the proposed method is that it does not assume the number of audio sources.
+However, a significant limitation of this study is the lack of reverberation in the synthesized recordings.
+
+The remainder of this paper is organized as follows.
+In Section \ref{sec:related-studies}, related studies are presented.
+The description of the method developed as part of this study is provided in Section \ref{sec:methodology}.
+This section also includes a detailed definition of ensemble width and the description of the experiments that was used to evaluate this method.
+Section \ref{sec:results} discusses the results of the method-evaluation experiments conducted in this study.
+Finally, the paper is summarized in Section \ref{sec:conclusions}.
+
 \section{Related studies}
+\label{sec:related-studies}
+
+The majority of existing literature on the subject of sound source localization employs techniques that leverage the advantages of microphone arrays with more than two channels \cite{kaveh_statistical_1986, pavlidi_real-time_2012, pan_multi-tone_2021, hahmann_sound_2022, chung_sound_2022, liu_sound_2022}.
+In the context of sound source localization in binaural signals, the predominant focus of research is on the identification of individual sound sources, rather than groups of sounds \cite{dietz_auditory_2011, may_probabilistic_2011, may_binaural_2012, woodruff_binaural_2012, may_robust_2015, ma16c_interspeech, ma_exploiting_2017, benaroya_binaural_2018}.
+In terms of source direction of arrival (DOA), the majority of research assumes a fixed number of sound sources \cite{pang_multitask_2019, vecchiotti19, ma_exploiting_2017, woodruff_binaural_2012, s_spatiogram_2021}, which limits its practical applications as this information is rarely known in real-life binaural recordings.
+Moreover, the majority of studies have focused on relatively homogeneous signals, namely speech \cite{dietz_auditory_2011, may_probabilistic_2011, may_binaural_2012, woodruff_binaural_2012, may_robust_2015, ma16c_interspeech, ma_exploiting_2017, benaroya_binaural_2018, wang_binaural_2020, liu_multiple_2018, yang_deepear_2022, ma_robust_2018}.
+
+In contrast to the aforementioned studies, the proposed method does not have any limitation to the number of sources.
+Furthermore, the approach is not limited to the speech, and indeed has been employed to work with much more diverse musical datasets, comprising sound sources such as instruments and vocals.
+In contrast to previous studies that have focused on individual sources, the proposed method does not aim to separate them but rather considers them as an ensemble, or in this case, a musical ensemble.
+To the authors' knowledge, this is one of the first methods to localize ensemble width (see \cite{antoniuk2023blind} for the previous study) and the first to localize both ensemble position and width simultaneously using a multi-task model.
 
 \section{Methodology}
+\label{sec:methodology}
 
-Experiments in this study were conducted on 2340 binaural recordings of music.
+Experiments in this study were conducted on 23040 binaural recordings of music.
 The binaural recordings were synthesized semi-automatically using 192 multi-track publicly-available music recordings and 30 HRTF databases.
-For each multi-track recording and HRTF database pair, four binaural recordings were synthesized for different random ensemble parameters --- its location $\phi$ and width $\theta$ --- as defined in Section \ref{subsec:ensemble-definition}.
+For each multi-track recording and HRTF database pair, four binaural recordings were synthesized for different random ensemble parameters --- its location $\theta$ and width $\theta$ --- as defined in Section \ref{subsec:ensemble-definition}.
 Both parameters were drawn from uniform random distribution.
@@
 
-\begin{figure}
+\begin{figure}[ht]
   \centering
@@
 \subsection{Synthesis of binaural music recordings}
+
 \subsection{Spectrogram conversions of binaural music recordings}
@@
 \subsection{Model training and evaluation}
+
 \section{Results}
+\label{sec:results}
 
@@
   \begin{minipage}[t]{0.45\linewidth}
-    \includegraphics[width=\linewidth]{../figures/actual_vs_predicted_width.pdf}
-    \caption{\label{fig:actual_vs_predicted_width}A comparison between the actual and the predicted ensemble width for a single iteration (of the total five) }
+    \includegraphics[width=\linewidth]{../figures/actual_vs_predicted_location.pdf}
+    \caption{\label{fig:actual_vs_predicted_location}A comparison between the actual ensemble location $\theta$ and the predicted ensemble location $\theta'$ for a single iteration (of the total five) }
   \end{minipage}
@@
   \begin{minipage}[t]{0.45\linewidth}
-    \includegraphics[width=\linewidth]{../figures/mae_width.pdf}
-    \caption{\label{fig:mae_width}The impact of the actual ensemble width on the mean absolute prediction error, averaged across all five iterations, with indicated standard deviation.}
+    \includegraphics[width=\linewidth]{../figures/mae_location.pdf}
+    \caption{\label{fig:mae_location}The impact of the actual ensemble width $\omega$ on the mean absolute prediction error, averaged across all five iterations, with indicated standard deviation.}
   \end{minipage}
@@
 Figure \ref{fig:actual_vs_predicted_width} illustrates the comparison between the actual and the predicted ensemble width.
-The results demonstrate that the model exhibits optimal prediction quality for narrower ensemble widths ($X\degree$ for $Y\degree$) and that its performance deteriorates with the increase of the ensemble width ($Z\degree$ for $W\degree$).
+The results demonstrate that the model exhibits better prediction quality for narrower ensemble widths ($12.44\degree$ for $\omega>80\degree$) and that its performance deteriorates with the increase of the ensemble width ($Z\degree$ for $W\degree$).
 Figure \ref{fig:mae_width} further demonstrates that the relationship between the prediction error and the actual width is not linear, exhibiting a depression between 60 and 75 degrees.
@@
   \begin{minipage}[t]{0.45\linewidth}
-    \includegraphics[width=\linewidth]{../figures/actual_vs_predicted_location.pdf}
-    \caption{\label{fig:actual_vs_predicted_location}A comparison between the actual and the predicted ensemble location for a single iteration (of the total five) }
+    \includegraphics[width=\linewidth]{../figures/actual_vs_predicted_width.pdf}
+    \caption{\label{fig:actual_vs_predicted_width}A comparison between the actual ensemble width $\omega$ and the predicted ensemble width $\omega'$ for a single iteration (of the total five) }
   \end{minipage}
@@
   \begin{minipage}[t]{0.45\linewidth}
-    \includegraphics[width=\linewidth]{../figures/mae_location.pdf}
-    \caption{\label{fig:mae_location}The impact of the actual ensemble width on the mean absolute prediction error, averaged across all five iterations, with indicated standard deviation.}
+    \includegraphics[width=\linewidth]{../figures/mae_width.pdf}
+    \caption{\label{fig:mae_width}The impact of the actual ensemble width $\omega$ on the mean absolute prediction error, averaged across all five iterations, with indicated standard deviation.}
   \end{minipage}
@@
 
-Figure \ref{fig:map_mae_location} shows influence of both the ensemble width and location on mean absolute error for ensemble location.
-This shows more detailed view of data previously presented on Figure \ref{fig:mae_location}.
-The figure shows that there is a slight asymmetry.
-
-\begin{figure}
+\begin{figure}[!htb]
   \centering
@@
 
-Similarly, Figure \ref{fig:map_mae_width} shows influence of both the ensemble width and location on mean absolute error for ensemble width.
+Figure \ref{fig:map_mae_location}  illustrates the influence of both ensemble width and location on mean absolute error for ensemble location.
+This figure provides a more detailed view of the data presented in Figure \ref{fig:mae_location}.
+The figure indicates a slight asymmetry.
+Similarly, Figure \ref{fig:map_mae_width} illustrates the influence of both ensemble width and location on mean absolute error for ensemble width.
 
 \section{Conclusions}
-\section{References}
+\label{sec:conclusions}
 
-\cite{greenwade93}
+\clearpage
 
-\bibliographystyle{alpha}
-\bibliography{references}
+\printbibliography
+
 \end{document}

### 2024-06-07 src/analyze/all_scores.py

@@ -1,20 +1,19 @@
 from pathlib import Path
 import pandas as pd
 import numpy as np
-import  joblib
+import joblib
 
-src_models_dir = Path('models')
+src_models_dir = Path("/app/data/train/out_models")
 test_scores = {}
 histories = {}
-for model_dir in src_models_dir.glob('*'):
-    for rep_it_path in model_dir.glob('*'):
-        score_path = next(rep_it_path.glob('score_*.csv'))
+for model_dir in src_models_dir.glob("*"):
+    for rep_it_path in model_dir.glob("*"):
+        score_path = next(rep_it_path.glob("score_*.csv"))
         score_df = pd.read_csv(score_path)
-        test_score = score_df['best_val_mae']
+        test_score = score_df["best_val_mae"]
         test_scores.setdefault(model_dir.name, []).append(float(test_score))
 
-        metadata = joblib.load(rep_it_path / 'metadata.pkl')
+        metadata = joblib.load(rep_it_path / "metadata.pkl")
         histories.setdefault(model_dir.name, []).append(metadata.history)
 
 {key: np.mean(values) * 90 for key, values in test_scores.items()}
-

### 2024-06-07 src/analyze/plot.py

@@ -1,89 +1,21 @@
 import os
-import sys
 from pathlib import Path
-
 import joblib
+
 import numpy as np
-import pandas as pd
 from dotmap import DotMap
 from matplotlib import pyplot as plt
 from matplotlib.ticker import FuncFormatter
 import tensorflow as tf
 import colormaps
 import matplotlib.colors as mcolors
-
 from model import Model
+from results import get_results
 
-model_name = "multi_inout_1_model_final_3_nogcc"
-model_dir = Path("/data/train/out_models") / model_name
+
 out_img_format = "pdf"
-
-
-def load_results():
-    prediction_results = []
-    for path_rep_it in model_dir.glob("*"):
-        request = joblib.load(path_rep_it / "request.pkl")
-        model = Model(request).try_load_or_compute_input_data()
-        prediction_result = model.predict_test_data()
-        errors_width = (
-            prediction_result.predicted_width - prediction_result.actual_width
-        )
-        errors_location = (
-            prediction_result.predicted_location - prediction_result.actual_location
-        )
-        score_test_width = np.mean(np.abs(errors_width))
-        score_test_location = np.mean(np.abs(errors_location))
-
-        # assertions
-        score_path = next(path_rep_it.glob("score_*.csv"))
-        score_df = pd.read_csv(score_path)
-        score_file_test_width = score_df["test_score_width"]
-        score_file_test_location = score_df["test_score_location"]
-        assert np.all(np.abs(score_file_test_width - score_test_width) < 1e-6)
-        assert np.all(np.abs(score_file_test_location -
-                      score_test_location) < 1e-6)
-
-        test_filenames = prediction_result.filename
-        actual_widths_test = [
-            float(x.split("_")[-4].replace("width", "")) for x in test_filenames
-        ]
-        stored_actual_widths_test = np.array(
-            model.store_split.y_test_width * 45.0)
-        assert np.all(abs(stored_actual_widths_test -
-                      actual_widths_test) < 1e-5)
-
-        actual_recordings = [
[…]

### 2024-06-07 src/analyze/results.py

@@ -1,0 +1,72 @@
+import joblib
+import os
+from pathlib import Path
+from model import Model
+import numpy as np
+import pandas as pd
+from dotmap import DotMap
+
+
+def load_results(model_dir: Path):
+    prediction_results = []
+    for path_rep_it in model_dir.glob("*"):
+        request = joblib.load(path_rep_it / "request.pkl")
+        print(f"loading model[{path_rep_it}] for request {request}")
+        model = Model(request).try_load_or_compute_input_data()
+        prediction_result = model.predict_test_data()
+        errors_width = (
+            prediction_result.predicted_width - prediction_result.actual_width
+        )
+        errors_location = (
+            prediction_result.predicted_location - prediction_result.actual_location
+        )
+        score_test_width = np.mean(np.abs(errors_width))
+        score_test_location = np.mean(np.abs(errors_location))
+
+        # assertions
+        score_path = next(path_rep_it.glob("score_*.csv"))
+        score_df = pd.read_csv(score_path)
+        score_file_test_width = score_df["test_score_width"]
+        score_file_test_location = score_df["test_score_location"]
+        assert np.all(np.abs(score_file_test_width - score_test_width) < 1e-6)
+        assert np.all(np.abs(score_file_test_location - score_test_location) < 1e-6)
+
+        test_filenames = prediction_result.filename
+        actual_widths_test = [
+            float(x.split("_")[-4].replace("width", "")) for x in test_filenames
+        ]
+        stored_actual_widths_test = np.array(model.store_split.y_test_width * 45.0)
+        assert np.all(abs(stored_actual_widths_test - actual_widths_test) < 1e-5)
+
+        actual_recordings = [
+            os.path.basename(x.split("_")[-7].replace("width", ""))
+            for x in test_filenames
+        ]
+        assert np.all(model.store_split.actual_recordings_test == actual_recordings)
+
+        prediction_results.append(
+            DotMap(
+                {
+                    "filename": prediction_result.filename,
+                    "actual_width": prediction_result.actual_width * 90,
+                    "predicted_width": prediction_result.predicted_width * 90,
+                    "errors_width": errors_width * 90,
+                    "actual_location": prediction_result.actual_location * 45,
+                    "predicted_location": prediction_result.predicted_location * 45,
+                    "errors_location": errors_location * 45,
+                }
+            )
+        )
+
+    return prediction_results
+
+
+def get_results(model_name: str):
[…]

### 2024-06-07 src/analyze/score_details.py

@@ -1,0 +1,42 @@
+from pathlib import Path
+from results import get_results
+import numpy as np
+
+model_name = "multi_inout_1_model_final_3_nogcc_nosame"
+
+
+def get_stats(results, filter = None) -> tuple[float, float, float, float]:
+    if filter:
+        error_widths = [result.errors_width[filter(result)] for result in results]    
+        error_locations = [result.errors_location[filter(result)] for result in results]    
+    else:
+        error_widths = [result.errors_width for result in results] 
+        error_locations = [result.errors_location for result in results]     
+    
+    error_widths_means = [np.mean(np.abs(err)) for err in error_widths]
+    mean_width = np.mean(error_widths_means)
+    std_width = np.std(error_widths_means)
+    
+    error_locations_means = [np.mean(np.abs(err)) for err in error_locations]
+    mean_location = np.mean(error_locations_means)
+    std_location = np.std(error_locations_means)
+    
+    return (mean_width, std_width, mean_location, std_location)
+
+
+results = get_results(model_name)
+
+mean_width, std_width, mean_location, std_location = get_stats(results)
+print(f"[all] Width MAE: {mean_width:0.2f} (std: {std_width:0.2f})")
+print(f"[all] Location MAE: {mean_location:0.2f} (std: {std_location:0.2f})")
+print()
+
+mean_width, std_width, mean_location, std_location = get_stats(results, lambda r: r.actual_width < 30)
+print(f"[width<30] Width MAE: {mean_width:0.2f} (std: {std_width:0.2f})")
+print(f"[wodth<30] Location MAE: {mean_location:0.2f} (std: {std_location:0.2f})")
+print()
+
+mean_width, std_width, mean_location, std_location = get_stats(results, lambda r: r.actual_width > 80)
+print(f"[width>80] Width MAE: {mean_width:0.2f} (std: {std_width:0.2f})")
+print(f"[wodth>80] Location MAE: {mean_location:0.2f} (std: {std_location:0.2f})")
+print()

### 2024-06-07 src/train/architecture/multi_inout_1.py

@@ -3,13 +3,13 @@ from keras import models as m
 
 
 def cnn_block(input_layer):
-    x = l.Conv2D(32, kernel_size=(2, 2), activation='relu', padding='same')(input_layer)
+    x = l.Conv2D(32, kernel_size=(2, 2), activation='relu')(input_layer)
     x = l.MaxPooling2D(pool_size=(2, 3))(x)
-    x = l.Conv2D(64, kernel_size=(2, 2), activation='relu', padding='same')(x)
+    x = l.Conv2D(64, kernel_size=(2, 2), activation='relu')(x)
     x = l.MaxPooling2D(pool_size=(2, 3))(x)
-    x = l.Conv2D(128, kernel_size=(2, 2), activation='relu', padding='same')(x)
+    x = l.Conv2D(128, kernel_size=(2, 2), activation='relu')(x)
     x = l.MaxPooling2D(pool_size=(2, 2))(x)
-    x = l.Conv2D(256, kernel_size=(2, 2), activation='relu', padding='same')(x)
+    x = l.Conv2D(256, kernel_size=(2, 2), activation='relu')(x)
     x = l.MaxPooling2D(pool_size=(2, 2))(x)
     x = l.GlobalAveragePooling2D()(x)
     return x

### 2024-06-07 src/train/model.py

@@ -261,7 +261,7 @@ class Model:
         start_time = timeit.default_timer()
         data_hash = self.__get_split_hash()
         print(f"[load] Split hash: {data_hash}")
-        self.store_split = joblib.load(f"/data/state/{data_hash}_store_split.pkl")
+        self.store_split = joblib.load(f"/app/data/train/state/{data_hash}_store_split.pkl")
         self.__initialize_split_memory()
         elapsed = timeit.default_timer() - start_time
         print(f"[load] Loaded cached splits in {elapsed:.2f} s")
@@ -271,8 +271,8 @@ class Model:
         start_time = timeit.default_timer()
         data_hash = self.__get_split_hash()
         print(f"[save] Split hash: {data_hash}")
-        os.makedirs("/data/state", exist_ok=True)
-        joblib.dump(self.store_split, f"/data/state/{data_hash}_store_split.pkl")
+        os.makedirs("/app/data/train/state", exist_ok=True)
+        joblib.dump(self.store_split, f"/app/data/train/state/{data_hash}_store_split.pkl")
         elapsed = timeit.default_timer() - start_time
         print(f"[save] Saved cached splits in {elapsed:.2f} s")
         return self
@@ -282,7 +282,7 @@ class Model:
         data_hash = self.__get_extract_hash()
         print(f"[load] Extraction hash: {data_hash}")
         self.store_extract = joblib.load(
-            f"/data/state/{data_hash}_store_extract.pkl")
+            f"/app/data/train/state/{data_hash}_store_extract.pkl")
         self.__initialize_extract_memory()
         elapsed = timeit.default_timer() - start_time
         print(f"[load] Loaded cached spectrograms in {elapsed:.2f} s")
@@ -292,8 +292,8 @@ class Model:
         start_time = timeit.default_timer()
         data_hash = self.__get_extract_hash()
         print(f"[save] Extraction hash: {data_hash}")
-        os.makedirs("/data/state/", exist_ok=True)
-        joblib.dump(self.store_extract, f"/data/state/{data_hash}_store_extract.pkl")
+        os.makedirs("/app/data/train/state/", exist_ok=True)
+        joblib.dump(self.store_extract, f"/app/data/train/state/{data_hash}_store_extract.pkl")
         elapsed = timeit.default_timer() - start_time
         print(f"[save] Saved spectrograms in {elapsed:.2f} s")
         return self
@@ -686,7 +686,7 @@ class Model:
 
         # save architecture source
         architecture_script_path = (
[…]

### 2024-06-07 src/train/run.py

@@ -7,7 +7,7 @@ from dotmap import DotMap
 
 from model import Model, Request
 
-with open('/src/train/requests.yaml') as f:
+with open('/app/src/train/requests.yaml') as f:
     data = yaml.safe_load(f)
     settings = DotMap(data, _dynamic=False)
 

### 2024-06-10 paper/main.tex
Update paper: related studies

@@
 Consequently, there is now a need for automated and objective measurements of spatial content information, with ensemble width and location being the most important.
-In this study, 30 head-related transfer functions and 192 binaural music recordings from publicly accessible multi-track recording repositories were used to synthesize 23,040 binaural recordings.
-The synthesized recordings were then used to train a convolutional neural network prediction model, the aim of which was to predict the width and location of the ensemble for unseen recordings.
+This study presents a novel method for predicting ensemble width and location in binaural recordings.
+To this end, 30 head-related transfer functions and 192 binaural music recordings from publicly accessible multi-track recording repositories were used to synthesize 23,040 binaural recordings.
+The synthesized recordings were then used to train a multi-task convolutional neural network prediction model, the aim of which was to predict the width and location of the ensemble for unseen recordings.
 The results indicate that models of ensemble breadth and position can be successfully constructed with low prediction errors --- $4.76\degree$ ($\pm0.10\degree$) for ensemble location and $8.57\degree$ ($\pm0.19\degree$) for ensemble width.
@@
 The human auditory system demonstrates exceptional proficiency in segregating, localizing, and interpreting diverse auditory signals, despite being limited to two ears.
-This capability arises from the intricate analysis of temporal, amplitude, and spectral disparities, a process termed binaural hearing \cite{blauert_spatial_1996}.
-This enables precise localization of sound sources in complex auditory environments.
-A particularly useful feature of binaural hearing is the 'cocktail party effect', which exemplifies the system's ability to focus on specific sounds while suppressing background noise \cite{cherry_experiments_1953}.
-An understanding of the auditory system is essential for an enhanced comprehension of its limits, the creation of more immersive binaural experiences for entertainment purposes \cite{zhang_surround_2017}, and the improvement of hearing aid devices to enhance auditory signal reception and improve spatial awareness \cite{thiemann_speech_2016}.
+This capability arises from the intricate analysis of temporal, amplitude, and spectral disparities, a process termed binaural hearing \cite{blauert_spatial_1996}, which enables precise localization of sound sources in complex auditory environments.
+An important advantage of binaural hearing is demonstrated by the `cocktail party effect', which showcases the system's capacity to concentrate on individual sounds while suppressing background noise \cite{cherry_experiments_1953}.
+An understanding of the auditory system is essential for comprehending its limits and creating more immersive binaural experiences for entertainment purposes \cite{zhang_surround_2017}.
+It also helps to improve hearing aid devices by enhancing auditory signal reception and improving spatial awareness \cite{thiemann_speech_2016}.
 
@@
 The findings demonstrate that this method is effective in accurately predicting the spatial characteristics of sound sources in near-real-world scenarios.
+Furthermore, this paper presents an experiment framework that allows for the objective measurement of a binaural localization technique in an objective manner, utilizing a large-scale dataset synthesized from real-world examples of music signals (for other use cases for this framework, see \cite{antoniuk2023blind, zielinski_automatic_2022, zielinski_spatial_2022, zielinski_comparison_2020}).
 One of the key advantages of the proposed method is that it does not assume the number of audio sources.
-However, a significant limitation of this study is the lack of reverberation in the synthesized recordings.
+However, a significant limitation of this study is the lack of reverberation in the synthesized recordings, which represents a material for future research.
 
@@
 In the context of sound source localization in binaural signals, the predominant focus of research is on the identification of individual sound sources, rather than groups of sounds \cite{dietz_auditory_2011, may_probabilistic_2011, may_binaural_2012, woodruff_binaural_2012, may_robust_2015, ma16c_interspeech, ma_exploiting_2017, benaroya_binaural_2018}.
-In terms of source direction of arrival (DOA), the majority of research assumes a fixed number of sound sources \cite{pang_multitask_2019, vecchiotti19, ma_exploiting_2017, woodruff_binaural_2012, s_spatiogram_2021}, which limits its practical applications as this information is rarely known in real-life binaural recordings.
+In terms of source direction of arrival (DOA) methods, the majority of research assumes a fixed number of sound sources \cite{pang_multitask_2019, vecchiotti19, ma_exploiting_2017, woodruff_binaural_2012, s_spatiogram_2021}, which limits its practical applications as this information is rarely known in real-life binaural recordings.
 Moreover, the majority of studies have focused on relatively homogeneous signals, namely speech \cite{dietz_auditory_2011, may_probabilistic_2011, may_binaural_2012, woodruff_binaural_2012, may_robust_2015, ma16c_interspeech, ma_exploiting_2017, benaroya_binaural_2018, wang_binaural_2020, liu_multiple_2018, yang_deepear_2022, ma_robust_2018}.
 
-In contrast to the aforementioned studies, the proposed method does not have any limitation to the number of sources.
-Furthermore, the approach is not limited to the speech, and indeed has been employed to work with much more diverse musical datasets, comprising sound sources such as instruments and vocals.
-In contrast to previous studies that have focused on individual sources, the proposed method does not aim to separate them but rather considers them as an ensemble, or in this case, a musical ensemble.
-To the authors' knowledge, this is one of the first methods to localize ensemble width (see \cite{antoniuk2023blind} for the previous study) and the first to localize both ensemble position and width simultaneously using a multi-task model.
+In contrast to the aforementioned studies, the proposed method is not limited by the number of sources.
+Furthermore, the approach is not limited to speech and has been applied to much more diverse musical datasets, including sound sources such as instruments and vocals.
+In contrast to previous studies that have focused on individual sources, the proposed method does not aim to separate them, but rather considers them as an ensemble, or in this case, a musical ensemble.
+This approach is inspired by the way real musical ensembles are arranged on stage.
+To the authors' knowledge, this is one of the first methods to localize ensemble width (see \cite{antoniuk2023blind} for the previous ensemble-width-related study), and the first to localize both ensemble position and width simultaneously using a multi-task model.
+
+Sound localization methods can be classified into two primary categories based on the implementation of their underlying algorithms: glass-box and black-box.
+The glass-box methods, which are more traditional in the literature, rely on manual designing algorithms that mimic the auditory system to explicitly extract key features from the signal for location prediction, such as one of combination of the following: interaural level difference, interaural time difference, interaural coherence or interaural phase difference (see \cite{blauert_spatial_1996} for features description).
+In contrast, black-box methods typically employ only basic feature engineering and rely heavily on modern machine learning techniques to both extract features and make the final prediction.
+Consequently, they are less reliant on manual intervention (see \cite{vecchiotti19} for an example).
+The black-box algorithm extracts features implicitly and internally, without knowledge of their internal purpose or the equivalent of the human auditory system.
+This renders such methods less transparent and less predictable.
+This, in conjunction with the fact that they are typically implemented with deep neural networks with a large number of learning parameters, necessitates the development of them with the use of very large datasets, typically comprising hundreds of thousands of examples.
 

### 2024-06-10 src/analyze/plot.py

@@ -139,6 +139,7 @@ plt.scatter(
     edgecolor="none",
     rasterized=True,
 )
+plt.axline((0, 0), (1, 1), color='r')
 plt.xlabel(r"Actual Width $\omega$")
 plt.ylabel(r"Predicted Width $\omega'$")
 plt.grid()
@@ -148,6 +149,7 @@ plt.xlim([0, 90])
 plt.ylim([0, 90])
 plt.gca().xaxis.set_major_formatter(FuncFormatter(lambda x, pos: f"{x}°"))
 plt.gca().yaxis.set_major_formatter(FuncFormatter(lambda x, pos: f"{x}°"))
+plt.legend(['Sample', 'One-to-one line'])
 plt.tight_layout()
 plt.savefig(f"/app/figures/actual_vs_predicted_width.{out_img_format}")
 # plt.show()
@@ -163,6 +165,7 @@ plt.scatter(
     edgecolor="none",
     rasterized=True,
 )
+plt.axline((0, 0), (1, 1), color='r')
 plt.xlabel(r"Actual Location $\theta$")
 plt.ylabel(r"Predicted Location $\theta'$")
 plt.grid()
@@ -172,6 +175,7 @@ plt.xlim([-45, 45])
 plt.ylim([-45, 45])
 plt.gca().xaxis.set_major_formatter(FuncFormatter(lambda x, pos: f"{x}°"))
 plt.gca().yaxis.set_major_formatter(FuncFormatter(lambda x, pos: f"{x}°"))
+plt.legend(['Sample', 'One-to-one line'])
 plt.tight_layout()
 plt.savefig(f"/app/figures/actual_vs_predicted_location.{out_img_format}")
 # plt.show()

### 2024-06-10 src/train/model.py

@@ -24,7 +24,6 @@ from dotmap import DotMap
 from scipy.signal import spectrogram
 from scipy.signal.windows import hamming
 from sklearn.model_selection import GroupShuffleSplit
-from keras import mixed_precision
 
 from gcc_phat import gcc_phat_feature, gcc_phat_feature_nsamples
 
@@ -91,8 +90,7 @@ class Model:
         self.memmap_dir = Path(request.memmap_dir)
         self.num_samples = request.num_samples
         self.sample_rate = request.sample_rate
-        self.filepaths = [p for p in self.input_dir.iterdir()
-                          ][0: self.num_samples]
+        self.filepaths = [p for p in self.input_dir.iterdir()][0 : self.num_samples]
         self.target_bands = request.target_bands
         self.time_window_len = request.time_window_len
         self.time_window_overlap = request.time_window_overlap
@@ -134,45 +132,26 @@ class Model:
         self.spectrogram_type = request.spectrogram_type
         self.dtype = request.dtype
         self.gcc_phat_len = request.gcc_phat_len
-
-        # self.log_buffer = io.StringIO()
-
         self.__print_properties_trimmed()
-
         keras.config.set_dtype_policy(self.dtype)
-
-        # Configure GPU VRAM
-        # gpus = tf.config.experimental.list_physical_devices('GPU')
-        # if gpus:
-        #     try:
-        #         config = [
-        #             tf.config.experimental.VirtualDeviceConfiguration(
-        #                 memory_limit=20 * 1024
-        #             )
-        #         ]
-        #         tf.config.experimental.set_virtual_device_configuration(
-        #             gpus[0], config)
-        #     except RuntimeError as e:
-        #         print(e)
-        #         traceback.print_exc()
 
     def extract(self) -> Model:
         start_time = timeit.default_timer()
 
         print("[extract] Initializing memory")
         self.__initialize_extract_memory()
-        self.store_extract.actual_widths = np.zeros(
-            self.num_samples, dtype=self.dtype)
+        self.store_extract.actual_widths = np.zeros(self.num_samples, dtype=self.dtype)
         self.store_extract.actual_locations = np.zeros(
-            self.num_samples, dtype=self.dtype)
-        self.store_extract.actual_recordings = [
-            "" for _ in range(self.num_samples)]
[…]

### 2024-06-11 paper/main.tex
Update paper: results

@@
 \usepackage{csquotes}
+\usepackage[toc,page]{appendix}
 
@@
 
-\title{Predicting Ensemble Width and Location in Binaural Recordings of Music with Convolutional Neural Networks}
+\title{Predicting Ensemble Location and Width in Binaural Recordings of Music with Convolutional Neural Networks}
 \author[1,*]{Paweł Antoniuk}
@@
 \begin{abstract}
-  Binaural audio technology has been around for many years, but its popularity has dramatically increased over the past decade due to advancements in virtual reality and streaming technologies.
+  Binaural audio technology has been in existence for many years, but its popularity has significantly increased over the past decade as a consequence of advancements in virtual reality and streaming technologies.
 Along with its growing popularity, the quantity of publicly accessible binaural audio materials has also expanded.
-Consequently, there is now a need for automated and objective measurements of spatial content information, with ensemble width and location being the most important.
-This study presents a novel method for predicting ensemble width and location in binaural recordings.
+Consequently, there is now a need for automated and objective measurements of spatial content information, with ensemble location and width being the most important.
+This study presents a novel method for predicting ensemble location and width in binaural recordings.
 To this end, 30 head-related transfer functions and 192 binaural music recordings from publicly accessible multi-track recording repositories were used to synthesize 23,040 binaural recordings.
-The synthesized recordings were then used to train a multi-task convolutional neural network prediction model, the aim of which was to predict the width and location of the ensemble for unseen recordings.
-The results indicate that models of ensemble breadth and position can be successfully constructed with low prediction errors --- $4.76\degree$ ($\pm0.10\degree$) for ensemble location and $8.57\degree$ ($\pm0.19\degree$) for ensemble width.
-This approach is the first of its kind to predict both ensemble width and location, offering a more accurate representation of spatial properties.
-This suggests significant potential for advancing spatial audio applications in virtual reality and streaming technologies, by providing audio engineers with tools that can leverage these methods to enhance spatial audio experiences.
+The synthesized recordings were then used to train a multi-task convolutional neural network model, the aim of which was to predict the location and width of the ensemble for unseen recordings.
+The results indicate that models of ensemble location and width can be successfully constructed with low prediction errors --- $4.76\degree$ ($\pm0.10\degree$) for ensemble location and $8.57\degree$ ($\pm0.19\degree$) for ensemble width.
+This approach is a novel method for predicting both ensemble location and width, offering a more accurate representation of spatial properties.
+This suggests significant potential for advancing spatial audio applications in virtual reality and streaming technologies, by providing audio engineers with tools that can leverage this method to enhance spatial audio experiences.
 \end{abstract}
@@
 An understanding of the auditory system is essential for comprehending its limits and creating more immersive binaural experiences for entertainment purposes \cite{zhang_surround_2017}.
-It also helps to improve hearing aid devices by enhancing auditory signal reception and improving spatial awareness \cite{thiemann_speech_2016}.
+It also helps to improve hearing aid devices by enhancing auditory signal reception and improving spatial awareness \cite{hirsh_binaural_1950, thiemann_speech_2016}.
 
@@
 
-The developed method has the potential to be highly beneficial in automated assessment tasks, where a significant number of binaural recordings must be evaluated and labeled in terms of their spatial content information.
-This may assist audio engineers in objectively assessing and segregating binaural audio recordings with regard to their spatial content.
-Furthermore, it could facilitate the development of an autonomous web-crawler bot that will collect binaural recordings from publicly accessible repositories and label them according to the spatial properties of the sound sources, such as the location of the music ensemble or the sparsity of audio source positions.
-
 The findings demonstrate that this method is effective in accurately predicting the spatial characteristics of sound sources in near-real-world scenarios.
@@
 
-The remainder of this paper is organized as follows.
-In Section \ref{sec:related-studies}, related studies are presented.
-The description of the method developed as part of this study is provided in Section \ref{sec:methodology}.
-This section also includes a detailed definition of ensemble width and the description of the experiments that was used to evaluate this method.
+The developed method has the potential to be highly beneficial in automated assessment tasks, where a significant number of binaural recordings must be evaluated and labeled in terms of their spatial content information.
+This may assist audio engineers in objectively assessing and segregating binaural audio recordings with regard to their spatial content.
+Furthermore, it could facilitate the development of an autonomous web-crawler bot that will collect binaural recordings from publicly accessible repositories and label them according to the spatial properties of the sound sources, such as the location of the music ensemble or the sparsity of audio source positions.
+
+The remainder of this paper is organized as follows: Section \ref{sec:related-studies} presents related studies.
+The description of the method developed for this study is provided in Section \ref{sec:methodology}, which also includes detailed definitions of ensemble location and width, along with a description of the experiments used to evaluate this method.
 Section \ref{sec:results} discusses the results of the method-evaluation experiments conducted in this study.
-Finally, the paper is summarized in Section \ref{sec:conclusions}.
+Finally, the paper concludes in Section \ref{sec:conclusions}.
 
@@
 
-In contrast to the aforementioned studies, the proposed method is not limited by the number of sources.
-Furthermore, the approach is not limited to speech and has been applied to much more diverse musical datasets, including sound sources such as instruments and vocals.
-In contrast to previous studies that have focused on individual sources, the proposed method does not aim to separate them, but rather considers them as an ensemble, or in this case, a musical ensemble.
-This approach is inspired by the way real musical ensembles are arranged on stage.
+In contrast to the aforementioned studies, the proposed method is not constrained by the number of sources.
+Furthermore, the approach is not limited to speech and has been applied to a wide range of musical datasets, including instruments and vocals.
+In contrast to previous studies that primarily focused on individual sources, the proposed method does not aim to separate them, but rather considers them as an ensemble, or in this case, a musical ensemble.
+This approach is similar to how the real musical ensembles are arranged on stage.
 To the authors' knowledge, this is one of the first methods to localize ensemble width (see \cite{antoniuk2023blind} for the previous ensemble-width-related study), and the first to localize both ensemble position and width simultaneously using a multi-task model.
 
-Sound localization methods can be classified into two primary categories based on the implementation of their underlying algorithms: glass-box and black-box.
-The glass-box methods, which are more traditional in the literature, rely on manual designing algorithms that mimic the auditory system to explicitly extract key features from the signal for location prediction, such as one of combination of the following: interaural level difference, interaural time difference, interaural coherence or interaural phase difference (see \cite{blauert_spatial_1996} for features description).
-In contrast, black-box methods typically employ only basic feature engineering and rely heavily on modern machine learning techniques to both extract features and make the final prediction.
-Consequently, they are less reliant on manual intervention (see \cite{vecchiotti19} for an example).
-The black-box algorithm extracts features implicitly and internally, without knowledge of their internal purpose or the equivalent of the human auditory system.
-This renders such methods less transparent and less predictable.
-This, in conjunction with the fact that they are typically implemented with deep neural networks with a large number of learning parameters, necessitates the development of them with the use of very large datasets, typically comprising hundreds of thousands of examples.
+Sound localization methods can be classified into two categories based on the implementation of their underlying algorithms: glass-box (e.g., \cite{dietz_auditory_2011, may_probabilistic_2011, may_binaural_2012, may_robust_2015, woodruff_binaural_2012, ma16c_interspeech, ma_exploiting_2017, ma_robust_2018}) and black-box (e.g., \cite{vecchiotti19, vera-diaz_towards_2018, yang_deepear_2022}).
+Glass-box methods, more traditional in the literature, rely on manually designed algorithms that mimic the auditory system to explicitly extract key features for location prediction, such as interaural level differences, interaural time differences, interaural coherence, or interaural phase differences (see \cite{blauert_spatial_1996} for feature descriptions).
+An example of one of the most advanced auditory models that is able to extract such features was developed by the Two!Ears project \cite{raake_computational_nodate}.
+
+In contrast, black-box methods typically employ minimal feature engineering and heavily rely on modern machine learning techniques to extract features and make predictions.
+These methods implicitly and internally extract features, without disclosing their purposes or mimicking the human auditory system.
+This lack of transparency and predictability, combined with their reliance on deep neural networks with numerous learning parameters, necessitates the development and evaluation using very large datasets.
+These datasets often comprise thousands of examples (e.g., 6300 examples in the TIMIT corpus \cite{garofolo_john_s_timit_1993} used in \cite{yang_deepear_2022, benaroya_binaural_2018, wang_binaural_2020, vecchiotti19, ma_exploiting_2017, pang_multitask_2019, ma_robust_2018, may_robust_2015}) or custom corpora consisting of hundreds of thousands of recordings used in \cite{antoniuk2023blind, zielinski_automatic_2022, zielinski_spatial_2022, zielinski_comparison_2020}.
 
@@
 
-Experiments in this study were conducted on 23040 binaural recordings of music.
-The binaural recordings were synthesized semi-automatically using 192 multi-track publicly-available music recordings and 30 HRTF databases.
-For each multi-track recording and HRTF database pair, four binaural recordings were synthesized for different random ensemble parameters --- its location $\theta$ and width $\theta$ --- as defined in Section \ref{subsec:ensemble-definition}.
-Both parameters were drawn from uniform random distribution.
+This section presents a detailed description of the main objective of the model developed as part of this study, as outlined in Section \ref{subsec:ensemble_definition}.
+It also describes the audio dataset used for training and evaluating the model, as detailed in Section \ref{subsec:synthesis}.
+In Section \ref{subsec:feature_extraction}, the feature extraction procedure is presented.
+Section \ref{subsec:topology} describes the model topology, whereas Section \ref{subsec:training_evaluation} address model training and evaluation.
 
 \subsection{Ensemble location and width definition}
-\label{subsec:ensemble-definition}
+\label{subsec:ensemble_definition}
 
-The primary objective of the model developed in this study is to predict the ensemble location ($\theta$) and width ($\omega$), as illustrated in Figure \ref{fig:scene}.
-The ensemble is defined as a group of audio point sources.
-The ensemble width $\omega$ is defined as the angular width between two extreme point sources, while the ensemble location $\theta$ represents the middle angle between two extreme sound sources.
+The objective of the model developed in this study is to predict the ensemble location ($\theta$) and width ($\omega$), as illustrated in Figure \ref{fig:scene}.
+An ensemble is defined as a group of audio point sources positioned on a circle around the listener on a virtual acoustic scene with equal distance to the listener.
+The location of source $i$ is denoted by $\theta_i$.
+The ensemble width ($\omega$) is defined as the angular width between two extreme point sources ($max_i(\theta_i)-min_i(\theta_i)$), while the ensemble location, designated by $\theta$, represents the middle angle between two extreme sound sources ($(max_i(\theta_i)+min_i(\theta_i))/2$).
+For the purposes of this study, the locations of the sources were limited to the frontal hemisphere only, i.e.
+$\theta\in[-45\degree, 45\degree]$, $\omega\in[0\degree, 90\degree]$, as this range encompasses the majority of real-world recording scenarios.
+It should be noted that although humans possess some limited abilities to localize sound sources in the vertical plane, in this study all sources are placed in the horizontal plane, at the height of the listener.
+This covers the majority of cases for real-world recordings (see \cite{ma_robust_2018, zielinski_spatial_2022} for similar studies that cover top-down discrimination).
 
@@
   \includegraphics[width=\linewidth]{../pictures/scene.png}
-  \caption{\label{fig:scene}Illustration of ensemble width ($\omega$) and ensemble location ($\theta$) relative to the direction of the head.
-Black dots represent the positions of audio sources.
+  \caption{\label{fig:scene}Illustration of ensemble width ($\omega$) and ensemble location ($\theta$) that are relative to the direction of the head.
+Black dots represent the positions of audio sources $\theta_i$.
 The ensemble location ($\theta$) is defined as the angular position of the center of the ensemble relative to the direction the head is turned.
@@
 \subsection{Synthesis of binaural music recordings}
+\label{subsec:synthesis}
 
-\subsection{Spectrogram conversions of binaural music recordings}
+The experiments conducted in this study involved 23,040 binaural recordings of music.
+The binaural recordings were synthesized using 192 multi-track publicly-available music recordings \cite{noauthor_mixing_nodate} and 30 HRTF databases (see \nameref{appendix:a_hrtf} for a detailed list of HRTF databases used in this study).
+The number of tracks in multi-track recording ranged from 5 to 62, with median of 9.
+For each pair of a multi-track recording and an HRTF database, four binaural recordings were synthesized with different random ensemble parameters, namely location $\theta$ and width $\omega$, as defined in Section \ref{subsec:ensemble_definition}.
+Both parameters were drawn from a uniform random distribution.
+Furthermore, the tracks of the input multi-track recordings were randomly assigned to sound source positions ($\theta_i$) to enhance the diversity of the final binaural corpora.
 
-\begin{figure}
+The binaural recordings were obtained in this study using the binaural synthesis procedure, known as binauralization, whose aim was to simulate the positions of sound sources within a virtual acoustic environment \cite{blauert_spatial_1996}.
+This was achieved by convolving multi-track signals with head-related impulse responses from a specified head-related transfer function (HRTF) database.
+The resulting binaural output signal $y_c[n]$ for each stereo channel $c$ (left or right) at sample $n$ is given by the following equation:
+\begin{equation}
+  y_c[n] = \sum_{i=1}^{N} \sum_{k=0}^{K-1} x_i[k] \times h_{c,\theta_i}[n-k] ,
+\end{equation}
+where $x_i$ represents the signal of an individual sound source $i$ from the input music recording and $h_{c,\theta_i}$ denotes the head-related impulse response for channel $c$ at location $\theta_i$ of source track $i$.
+
+Due to copyright restrictions, the music corpus utilized in this study was not published and can be provided upon request to the authors of this paper.
+
+\subsection{Feature extraction}
+\label{subsec:feature_extraction}
+
+% The procedure proposed in this study did not
+
+\begin{figure}[ht]
   \centering
   \includegraphics[width=\linewidth]{../figures/samples.pdf}
-  \caption{\label{fig:samples}An illustrative example of the magnitude spectrograms generated from synthesized binaural recordings of music.
-Each spectrogram was derived from a different randomly selected binaural music recording.
-These spectrograms were subsequently employed as input data for training and assessing the convolutional neural network (CNN) in this study.
+  \caption{\label{fig:samples}Example magnitude spectrograms from randomly selected synthesized binaural music recordings, used as input data for training and evaluating the convolutional neural network (CNN) in this study.
 }
@@
 
-Prior to being fed into the model, the binaural recordings of music were transformed into magnitude spectrograms.
-Figure \ref{fig:samples} illustrates example spectrograms for four distinct binaural recordings of music.
-It should be noted that the original spectrograms were stored in a floating-point-precision matrix format.
+Prior to input into the model, the binaural recordings of music were transformed into magnitude spectrograms.
+Figure \ref{fig:samples} displays example spectrograms from four distinct binaural recordings, exhibiting a broad bandwidth and the diversity of the signals used in this study.
+It is important to note that the original spectrograms were preserved in a floating-point precision matrix format, which prevented any loss of information due to precision conversion.
+The spectrograms were calculated using the Fast Fourier Transform (FFT) algorithm, with a limitation of 150 bands, spaced linearly from 100 Hz to 16 kHz.
+Additionally, a Hamming window of 40 ms was applied to each frame of the signal, resulting in a total of 349 time frames.
+This procedure was conducted for both the left and right channels, yielding two spectrograms for each binaural sample.
+Each sample was represented by a matrix of dimensions $2 \times 349 \times 150$.
+This method parallels the procedure presented in \cite{zielinski_automatic_2022}.
 
-\subsection{Convolutional neural network topology}
+\subsection{Network topology}
+\label{subsec:topology}
+
 \subsection{Model training and evaluation}
+\label{subsec:training_evaluation}
 
@@
 \label{sec:results}
+
+\begin{figure}[ht]
+  \centering
+  \begin{minipage}[t]{0.45\linewidth}
+    \includegraphics[width=\linewidth]{../figures/actual_vs_predicted_width.pdf}
+    \caption{\label{fig:actual_vs_predicted_width}A comparison between the actual ensemble width $\omega$ and the predicted ensemble width $\omega'$ for a single iteration (of the total five) }
+  \end{minipage}
+  \hspace{0.5cm}
+  \begin{minipage}[t]{0.45\linewidth}
+    \includegraphics[width=\linewidth]{../figures/mae_width.pdf}
+    \caption{\label{fig:mae_width}The impact of the actual ensemble width $\omega$ on the mean absolute prediction error, averaged across all five iterations, with indicated standard deviation.}
+  \end{minipage}
+\end{figure}
+
+Figure \ref{fig:actual_vs_predicted_width} illustrates the comparison between the actual and predicted ensemble widths.
+The results demonstrate that the model exhibits better prediction quality for narrower ensemble widths ($5.65^\circ$ for $\omega < 30^\circ$) and that its performance deteriorates with an increase in ensemble width ($12.44^\circ$ for $\omega > 80^\circ$).
+This suggests that the actual ensemble width significantly impacts the accuracy of ensemble width estimation, leading to worse predictions as width increases.
+Figure \ref{fig:mae_width} further demonstrates that the relationship between the error and the width is not linear, exhibiting an interesting depression between $60^\circ$ and $75^\circ$.
 
@@
 
-Figure \ref{fig:actual_vs_predicted_width} illustrates the comparison between the actual and the predicted ensemble width.
-The results demonstrate that the model exhibits better prediction quality for narrower ensemble widths ($12.44\degree$ for $\omega>80\degree$) and that its performance deteriorates with the increase of the ensemble width ($Z\degree$ for $W\degree$).
-Figure \ref{fig:mae_width} further demonstrates that the relationship between the prediction error and the actual width is not linear, exhibiting a depression between 60 and 75 degrees.
-This suggests, that the ensemble width has a significant impact on the ensemble width estimation.
-
-\begin{figure}[ht]
-  \centering
-  \begin{minipage}[t]{0.45\linewidth}
-    \includegraphics[width=\linewidth]{../figures/actual_vs_predicted_width.pdf}
-    \caption{\label{fig:actual_vs_predicted_width}A comparison between the actual ensemble width $\omega$ and the predicted ensemble width $\omega'$ for a single iteration (of the total five) }
-  \end{minipage}
-  \hspace{0.5cm}
-  \begin{minipage}[t]{0.45\linewidth}
-    \includegraphics[width=\linewidth]{../figures/mae_width.pdf}
-    \caption{\label{fig:mae_width}The impact of the actual ensemble width $\omega$ on the mean absolute prediction error, averaged across all five iterations, with indicated standard deviation.}
-  \end{minipage}
-\end{figure}
-
 In contrast to the correlation between ensemble width and its prediction error, there is no significant relationship between the actual location and its prediction error, as illustrated in Figures \ref{fig:actual_vs_predicted_location} and \ref{fig:mae_location}.
 This finding indicates that the model's capabilities for localizing the center of the ensemble is robust, unaffected by the actual spatial positioning of the ensemble, including lateral locations.
+
+The mean absolute prediction error for ensemble width was $8.57^\circ$ ($\pm0.19^\circ$), while the error for ensemble location prediction was $4.76^\circ$ ($\pm0.10^\circ$).
+Although both ensemble parameters were constrained within the same angular span of $90^\circ$, the ensemble location was predicted with significantly better precision --- approximately $80\%$ better.
+This shows that the model prediction quality is much better for ensemble location prediction than width prediction.
+This difference can be attributed to the fact that the model's performance on ensemble width deteriorates significantly for wider ensembles, whereas location prediction remains largely unaffected.
+For a visual comparison, please refer to Figures \ref{fig:mae_location} and \ref{fig:mae_width}.
 
@@
 
-\begin{figure}[!htb]
+\begin{figure}[ht]
   \centering
@@
 
-Figure \ref{fig:map_mae_location}  illustrates the influence of both ensemble width and location on mean absolute error for ensemble location.
-This figure provides a more detailed view of the data presented in Figure \ref{fig:mae_location}.
-The figure indicates a slight asymmetry.
-Similarly, Figure \ref{fig:map_mae_width} illustrates the influence of both ensemble width and location on mean absolute error for ensemble width.
+Figure \ref{fig:map_mae_location} illustrates the influence of both ensemble location and width on the mean absolute error for ensemble location, offering a detailed view complementing the results presented in Figure \ref{fig:mae_location}.
+Some asymmetric anomalies are depicted in this figure, mostly within the $\theta \in [15\degree, 30\degree]$ region, which can be attributed to the sparsity of sample result data across this heatmap.
+Although this figure primarily shows that ensemble location does not significantly influence the model's location prediction accuracy, it does demonstrate that ensemble width has a significant impact.
+Similarly, Figure \ref{fig:map_mae_width} reveals a characteristic depression in $\omega \in [30\degree, 60\degree]$ previously shown from a different perspective in Figure \ref{fig:mae_width}.
+This heatmap highlights another interesting phenomenon in its upper corners --— the error in these areas is considerably higher.
+This indicates that the model's performance for estimating ensemble width is substantially worse at extreme widths and locations, i.e., when both the width and locations are at their maximum.
 
@@
 
+\section*{Appendix A}
+\label{appendix:a_hrtf}
+
 \printbibliography

### 2024-06-11 src/analyze/plot.py

@@ -22,7 +22,7 @@ def load_sample_data():
     src_request_path = model_dir / "0" / "request.pkl"
     request = joblib.load(src_request_path)
     model = Model(request).try_load_or_compute_input_data()
-    np.random.seed(1)
+    np.random.seed(3)
     n_samples = 100
     random_indices = np.random.choice(
         model.store_split.x_train_mag_shape[0], n_samples, replace=False
@@ -47,7 +47,7 @@ def load_sample_data():
 
 
 def get_sample_data():
-    sample_data_filename = "/app/data/analyze/state/analyze_sample_data.pkl"
+    sample_data_filename = f"/app/data/analyze/state/analyze_sample_data{model_name}.pkl"
     try:
         return joblib.load(sample_data_filename)
     except FileNotFoundError:

### 2024-06-11 src/analyze/score_details.py

@@ -29,6 +29,7 @@ results = get_results(model_name)
 mean_width, std_width, mean_location, std_location = get_stats(results)
 print(f"[all] Width MAE: {mean_width:0.2f} (std: {std_width:0.2f})")
 print(f"[all] Location MAE: {mean_location:0.2f} (std: {std_location:0.2f})")
+print(f"[all] Location MAE is better than Width MAE by {(mean_width / mean_location - 1)*100:0.2f}%")
 print()
 
 mean_width, std_width, mean_location, std_location = get_stats(results, lambda r: r.actual_width < 30)

### 2024-06-19 paper/main.tex
Add experiments with noise prediction

@@
 Along with its growing popularity, the quantity of publicly accessible binaural audio materials has also expanded.
-Consequently, there is now a need for automated and objective measurements of spatial content information, with ensemble location and width being the most important.
+Consequently, there is now a need for automated and objective measurements of spatial content information, with ensemble location and width being the most prominent.
 This study presents a novel method for predicting ensemble location and width in binaural recordings.
@@
 The results indicate that models of ensemble location and width can be successfully constructed with low prediction errors --- $4.76\degree$ ($\pm0.10\degree$) for ensemble location and $8.57\degree$ ($\pm0.19\degree$) for ensemble width.
-This approach is a novel method for predicting both ensemble location and width, offering a more accurate representation of spatial properties.
-This suggests significant potential for advancing spatial audio applications in virtual reality and streaming technologies, by providing audio engineers with tools that can leverage this method to enhance spatial audio experiences.
+The method developed in this study outperforms previous spatiogram-based methods recently published in the literature and holds promise for future development as a part of a novel tool for audio engineers to assess binaural audio.
 \end{abstract}
@@
 The advancement of sophisticated machine learning techniques, particularly deep learning networks, has prompted an intriguing exploration of the extent to which these tools can emulate the human auditory system without relying on advanced spatial audio feature engineering, traditionally employed in audio source localization tasks \cite{yang_deepear_2022, vera-diaz_towards_2018, pang_multitask_2019}.
-To investigate this, the study developed an audio localization method based on a multi-task convolutional neural network (CNN) model.
+To investigate this, the study developed an audio localization method based on a multi-task convolutional neural network (CNN) model \cite{lecun_handwritten_1989, lecun_deep_2015}.
 
@@
 
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\linewidth]{../figures/samples.pdf}
-  \caption{\label{fig:samples}Example magnitude spectrograms from randomly selected synthesized binaural music recordings, used as input data for training and evaluating the convolutional neural network (CNN) in this study.
+% Zamiast tego dac obrazek z glowami i spektrogramami, lewy i prawi i roznica, sprawdzic czy to ma sens, potencjalnie usunac
+% \begin{figure}[ht]
+%   \centering
+%   \includegraphics[width=\linewidth]{../figures/samples.pdf}
+%   \caption{\label{fig:samples}Example magnitude spectrograms from randomly selected synthesized binaural music recordings, used as input data for training and evaluating the convolutional neural network (CNN) in this study.
 }
-\end{figure}
+% \end{figure}
+
+% Figure \ref{fig:samples} displays example spectrograms from four distinct binaural recordings, exhibiting a broad bandwidth and the diversity of the signals used in this study
 
 Prior to input into the model, the binaural recordings of music were transformed into magnitude spectrograms.
-Figure \ref{fig:samples} displays example spectrograms from four distinct binaural recordings, exhibiting a broad bandwidth and the diversity of the signals used in this study.
 It is important to note that the original spectrograms were preserved in a floating-point precision matrix format, which prevented any loss of information due to precision conversion.
@@
 \label{subsec:topology}
+
+The network topology used in this study was highly inspired by a AlexNet convolutional neural network \cite{krizhevsky_imagenet_2012}.
+Although AlexNet network was designed for image classification, the binaural recordings in this study were converted into magnitude spectrograms, as described in Section \ref{subsec:feature_extraction}.
+The network is consisting of a series of convolutional units followed by a series of classification units.
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\linewidth]{../pictures/architecture.pdf}
+  \caption{\label{fig:architecture}Topology of the Convolutional Neural Network (CNN) used for identifying ensemble location and width}
+\end{figure}
 
@@
 
+% jak uzyskane wyniki koresponduja z literatura: w jjednym przypadku ejst lepsza niz metoda oparta na spatiogramach, a w drugim przypadku nie ma 
 The mean absolute prediction error for ensemble width was $8.57^\circ$ ($\pm0.19^\circ$), while the error for ensemble location prediction was $4.76^\circ$ ($\pm0.10^\circ$).
@@
 
+% 1. napisac wnioski, potem occulusion. we wnioskach odniesc sie do spatiogramach
+
 \clearpage
-
 \section*{Appendix A}

### 2024-06-19 src/analyze/occlusion.py

@@ -1,0 +1,29 @@
+import os
+from pathlib import Path
+import joblib
+
+import numpy as np
+from dotmap import DotMap
+from matplotlib import pyplot as plt
+from matplotlib.ticker import FuncFormatter
+import tensorflow as tf
+import colormaps
+import matplotlib.colors as mcolors
+from model import Model
+from results import get_results
+
+model_name = "multi_inout_1_model_final_3_nogcc_nosame"
+model_dir = Path("/app/data/train/out_models") / model_name
+src_request_path = model_dir / "0" / "request.pkl"
+request = joblib.load(src_request_path)
+
+model = Model(request).try_load_or_compute_input_data()
+
+for i_sample in range(3):
+    X = model.x_test_mag[i_sample:i_sample+1, :, :, :]
+    y_actual_location = model.store_split.y_test_location[i_sample]
+    y_actual_width = model.store_split.y_test_width[i_sample]
+    y_pred = model.predict(X)
+    err_location = abs(y_actual_location - y_pred['predicted_location']) * 45
+    err_width = abs(y_actual_width - y_pred['predicted_width']) * 90
+    print(f'Score: Location: {err_location}, Width: {err_width}')

### 2024-06-19 src/analyze/plot_spectrograms_comp.py

@@ -1,0 +1,137 @@
+import os
+from pathlib import Path
+import joblib
+
+# import sys
+
+# sys.path.append('"/app/src/train')
+
+import numpy as np
+from dotmap import DotMap
+from matplotlib import pyplot as plt
+from matplotlib.ticker import FuncFormatter
+import tensorflow as tf
+import colormaps
+import matplotlib.colors as mcolors
+from model import Model
+from results import get_results
+import soundfile as sf
+
+
+model_name = "final_width_location"
+model_dir = Path("/app/data/train/out_models") / model_name
+src_request_path = model_dir / "1" / "request.pkl"
+request = joblib.load(src_request_path)
+data_dir = Path("/app/data/spatialize/spatresults/spat")
+model = Model(request).try_load_or_compute_input_data()
+errors_width = []
+errors_location = []
+
+spectrogram_per_actual_width_l = {}
+spectrogram_per_actual_width_r = {}
+
+for sample_path in list(data_dir.glob("*.wav")):
+    print(f"[test] sample: {sample_path}")
+    metadata = model.extract_metadata_from_sample_name(sample_path)
+    audio_data, sample_rate = sf.read(sample_path)
+    features = model.extract_features_from_raw_signal(audio_data, sample_rate)
+    pred = model.predict_by_features(features)
+    error_width = abs(pred.predicted_width[0] - metadata.actual_width * 2)
+    error_location = abs(pred.predicted_location[0] - metadata.actual_location)
+    errors_width.append(error_width)
+    errors_location.append(error_location)
+    actual_width = int(metadata.actual_width) // 5 * 5
+    spectrogram_per_actual_width_l.setdefault(actual_width, []).append(
+        features["in_mag"][0, :, :, 0]
+    )
+    spectrogram_per_actual_width_r.setdefault(actual_width, []).append(
+        features["in_mag"][0, :, :, 1]
+    )
+
+    print(
+        f"[test] actual_width: {metadata.actual_width * 2}, \
+          actual_location: {metadata.actual_location}"
+    )
+    print(
+        f"[test] predicted_width: {pred.predicted_width[0]}, \
+          predicted_location: {pred.predicted_location[0]}"
+    )
[…]

### 2024-06-19 src/analyze/plot_spectrograms_comp_loc.py

@@ -1,0 +1,139 @@
+import os
+from pathlib import Path
+import joblib
+
+# import sys
+
+# sys.path.append('"/app/src/train')
+
+import numpy as np
+from dotmap import DotMap
+from matplotlib import pyplot as plt
+from matplotlib.ticker import FuncFormatter
+import tensorflow as tf
+import colormaps
+import matplotlib.colors as mcolors
+from model import Model
+from results import get_results
+import soundfile as sf
+
+
+model_name = "final_width_location"
+model_dir = Path("/app/data/train/out_models") / model_name
+src_request_path = model_dir / "1" / "request.pkl"
+request = joblib.load(src_request_path)
+data_dir = Path("/app/data/spatialize/spatresults-location/spat")
+model = Model(request).try_load_or_compute_input_data()
+errors_width = []
+errors_location = []
+
+spectrogram_per_actual_location_l = {}
+spectrogram_per_actual_location_r = {}
+
+for sample_path in list(data_dir.glob("*.wav")):
+    print(f"[test] sample: {sample_path}")
+    metadata = model.extract_metadata_from_sample_name(sample_path)
+    audio_data, sample_rate = sf.read(sample_path)
+    features = model.extract_features_from_raw_signal(audio_data, sample_rate)
+    pred = model.predict_by_features(features)
+    error_width = abs(pred.predicted_width[0] - metadata.actual_width * 2)
+    error_location = abs(pred.predicted_location[0] - metadata.actual_location)
+    errors_width.append(error_width)
+    errors_location.append(error_location)
+    actual_location = int(metadata.actual_location) // 5 * 5
+    spectrogram_per_actual_location_l.setdefault(actual_location, []).append(
+        features["in_mag"][0, :, :, 0]
+    )
+    spectrogram_per_actual_location_r.setdefault(actual_location, []).append(
+        features["in_mag"][0, :, :, 1]
+    )
+
+    print(
+        f"[test] actual_width: {metadata.actual_width * 2}, \
+          actual_location: {metadata.actual_location}"
+    )
+    print(
+        f"[test] predicted_width: {pred.predicted_width[0]}, \
+          predicted_location: {pred.predicted_location[0]}"
+    )
[…]

### 2024-06-19 src/analyze/results.py

@@ -13,23 +13,26 @@ def load_results(model_dir: Path):
         request = joblib.load(path_rep_it / "request.pkl")
         print(f"loading model[{path_rep_it}] for request {request}")
         model = Model(request).try_load_or_compute_input_data()
+
         prediction_result = model.predict_test_data()
-        errors_width = (
-            prediction_result.predicted_width - prediction_result.actual_width
-        )
-        errors_location = (
-            prediction_result.predicted_location - prediction_result.actual_location
-        )
+        predicted_width = prediction_result.predicted_width * 90
+        actual_width = prediction_result.actual_width * 90
+        predicted_location = prediction_result.predicted_location * 45
+        actual_location = prediction_result.actual_location * 45
+
+        errors_width = predicted_width - actual_width
+        errors_location = predicted_location - actual_location
         score_test_width = np.mean(np.abs(errors_width))
         score_test_location = np.mean(np.abs(errors_location))
 
         # assertions
         score_path = next(path_rep_it.glob("score_*.csv"))
         score_df = pd.read_csv(score_path)
-        score_file_test_width = score_df["test_score_width"]
-        score_file_test_location = score_df["test_score_location"]
-        assert np.all(np.abs(score_file_test_width - score_test_width) < 1e-6)
-        assert np.all(np.abs(score_file_test_location - score_test_location) < 1e-6)
+        score_file_test_width = score_df["test_score_width"] * 90
+        score_file_test_location = score_df["test_score_location"] * 45
+
+        assert np.all(np.abs(score_file_test_width - score_test_width) < 1e-4)
+        assert np.all(np.abs(score_file_test_location - score_test_location) < 1e-4)
 
         test_filenames = prediction_result.filename
         actual_widths_test = [
@@ -48,12 +51,12 @@ def load_results(model_dir: Path):
             DotMap(
                 {
                     "filename": prediction_result.filename,
-                    "actual_width": prediction_result.actual_width * 90,
-                    "predicted_width": prediction_result.predicted_width * 90,
-                    "errors_width": errors_width * 90,
-                    "actual_location": prediction_result.actual_location * 45,
-                    "predicted_location": prediction_result.predicted_location * 45,
-                    "errors_location": errors_location * 45,
+                    "actual_width": actual_width,
+                    "predicted_width": predicted_width,
+                    "errors_width": errors_width,
+                    "actual_location": actual_location,
+                    "predicted_location": predicted_location,
+                    "errors_location": errors_location,
                 }
             )
         )

### 2024-06-19 src/spatialize/SOFAresample.m

@@ -1,0 +1,27 @@
+% Resample SOFA routine
+function Obj = SOFAresample(Obj, targetFs)
+    currentFs = Obj.Data.SamplingRate;
+
+    if currentFs == targetFs
+        return
+    end
+
+    % Based on HRTFsamplingRateConverter10.m (S. Zieliński)
+    M = size(Obj.Data.IR,1); % Number of measurements
+    N = size(Obj.Data.IR,3); % Length of measurements
+    IR = Obj.Data.IR;
+    IR2 = zeros(M, 2, round(ceil(targetFs / currentFs * N*3) / 3));
+
+    for ii = 1:M
+        ir = squeeze(IR(ii, :, :))';
+        irx3 = [ir; ir; ir];
+        irx3 = resample(irx3, targetFs, currentFs);
+        N2 = round(length(irx3)/3);
+        ir2 = irx3(N2+1:2*N2, :);
+        IR2(ii, :, :) = ir2';
+    end
+
+    Obj.Data.IR = IR2;
+    Obj.Data.SamplingRate = targetFs;
+    Obj=SOFAupdateDimensions(Obj);
+end

### 2024-06-19 src/spatialize/copyFrontRecordings.m

@@ -1,0 +1,30 @@
+inputDir = 'rec_front_small_uniform/spat';
+outputDir = 'frontNewMIT3';
+
+filenames = dir(fullfile(inputDir, "*.wav"));
+fullFilenames = fullfile(inputDir, {filenames.name});
+[~, filenames, ~] = fileparts({filenames.name});
+filenameParts = split(filenames, "_");
+realWidths = erase(filenameParts(1, :, 4), 'width');
+realWidths = str2double(convertCharsToStrings(realWidths));
+realLocs = erase(filenameParts(1, :, 6), 'azoffset');
+realLocs = str2double(convertCharsToStrings(realLocs));
+HRTFIDs = filenameParts(1, :, 2);
+HRTFGroupNames = filenameParts(1, :, 3);
+
+if ~isfolder(outputDir)
+    mkdir(outputDir)
+end
+
+delete(fullfile(outputDir, '*.wav'))
+
+for i = 1:length(fullFilenames)
+    HRTFID = HRTFIDs(i);
+    HRTFGroupName = HRTFGroupNames(i);
+    width = realWidths(i); % 0-90
+    loc = realLocs(i); % -180-180
+    if HRTFGroupName == "mit" && width + abs(loc) <= 90
+        copyfile(fullFilenames{i}, outputDir)
+    end
+end
+

### 2024-06-19 src/spatialize/dirWithoutDots.m

@@ -1,0 +1,3 @@
+function items = dirWithoutDots(path)
+items = dir(path);
+items = items(~startsWith({items.name}, '.'));

### 2024-06-19 src/spatialize/generateNoiseRecordings.m

@@ -1,0 +1,10 @@
+T = 10;
+fs = 48000;
+nch = 15;
+s = wgn(T * fs, nch, 1);
+outputDir = ['../recordings-all/Noise' num2str(nch)];
+
+for iChannel = 1:size(s, 2)
+    name = sprintf("%s/%d.wav", outputDir, iChannel);
+    audiowrite(name, s(:, iChannel), fs, BitsPerSample=32)
+end

### 2024-06-19 src/spatialize/geninputs.m

@@ -1,0 +1,14 @@
+function geninputs(params)
+
+fs = 48000;
+duration = 10;
+outFolderPath = fullfile(params.RecordingsBaseDir, "whitenoise");
+
+delete(fullfile(outFolderPath, '*'))
+    
+for i = 1:10
+    y = wgn(1, duration * fs, 1);
+    outFilenamePath = fullfile(outFolderPath, "whitenoise" + i + ".wav");
+    mkdir(outFolderPath);
+    audiowrite(outFilenamePath, y, fs);
+end

### 2024-06-19 src/spatialize/getSceneMetaresult.m

@@ -1,0 +1,24 @@
+function metaResults = getSceneMetaresult(HRTF, audioName, trackNames, ...
+    elevation, azimuthLocation, params)
+
+    width = rand * params.MaxWidth;
+%     azimuthOffset = (params.AzimuthOffsetRange(2) ...
+%         - params.AzimuthOffsetRange(1)) .* rand ...
+%         + params.AzimuthOffsetRange(1);
+%     azimuthOffset = wrapTo180(azimuthOffset);
+    azimuthOffset = azimuthLocation;
+
+%     randTrackAngles = rand(length(trackNames), 1);
+%     randTrackAngles = rescale(randTrackAngles, -width, width);
+%     randTrackAngles = randTrackAngles + azimuthOffset;
+%     randTrackAngles = wrapTo180(randTrackAngles);
+%     randTrackAngles(:, 2) = elevation;
+
+    metaResults.AudioName = audioName;
+    metaResults.TrackNames = trackNames;
+    metaResults.HRTFId = HRTF.Id;
+    metaResults.RandTrackAngles = [azimuthLocation' zeros(size(azimuthLocation,2),1)];
+    metaResults.Elevation = elevation;
+    metaResults.SceneWidth = width;
+    metaResults.AzimuthEnsembleOffset = azimuthOffset;
+end

### 2024-06-19 src/spatialize/loadAudioTracks.m

@@ -1,0 +1,23 @@
+% load tracks routine
+function [tracks,trackNames] = loadAudioTracks(audioFilename, params)
+    songName = fullfile(audioFilename.folder, audioFilename.name);
+    trackFilenames = dir(fullfile(songName, '*.wav'));
+    audioInfo = audioinfo(fullfile(trackFilenames(1).folder, ...
+        trackFilenames(1).name));
+    totalSamples = audioInfo.TotalSamples;
+    tracks = zeros(totalSamples, length(trackFilenames));
+
+    for iTrackFilename = 1:length(trackFilenames)
+        trackPath = fullfile(trackFilenames(iTrackFilename).folder, ...
+            trackFilenames(iTrackFilename).name);
+        [track,Fs] = audioread(trackPath);
+
+        if Fs ~= params.RecordingsExpectedFs
+            error('Track frequency is not expected frequency');
+        end
+
+        tracks(:, iTrackFilename) = track;
+    end
+
+    trackNames = {trackFilenames.name};
+end

### 2024-06-19 src/spatialize/loadHRTFs.m

@@ -1,0 +1,114 @@
+% Load HRTFs routine
+function HRTFs = loadHRTFs(params)
+    HRTFFilenames = dir(fullfile(params.HRTFBaseDir, '*', '*.sofa'));
+
+    % HRTF struct definition
+    HRTFs = struct('Id', [], ...
+        'Name', [], ...
+        'Folder', [], ...
+        'HRTFGroup', [], ...
+        'SOFA', [], ...
+        'Position', [], ...
+        'Distance', []);
+    HRTFGroupData = containers.Map;
+
+    for iHRTF = 1:length(HRTFFilenames)
+        filename = HRTFFilenames(iHRTF);
+        fullFilename = fullfile(filename.folder, filename.name);
+
+        HRTFs(iHRTF) = loadHRTF(iHRTF, fullFilename, params);
+
+        if HRTFs(iHRTF).SOFA.Data.SamplingRate ~= params.RecordingsExpectedFs
+            [loadStatus,HRTFs(iHRTF)] = tryLoadResampledHRTF(iHRTF, ...
+                HRTFs(iHRTF), params);
+            if ~loadStatus
+                resampleAndSave(HRTFs(iHRTF), params);
+                [loadStatus,HRTFs(iHRTF)] = tryLoadResampledHRTF(iHRTF, ...
+                    HRTFs(iHRTF), params);
+
+                if ~loadStatus
+                    error('Cannot find previously resampled HRTF');
+                end
+            end
+        end
+
+        ir =  HRTFs(iHRTF).SOFA.Data.IR;
+        if size(ir, 3) > params.IRmax
+            Nfadeout = params.FadeDuration*params.RecordingsExpectedFs;
+            fade = [repelem(1,params.IRmax-Nfadeout), ...
+                (cos(linspace(0,pi,Nfadeout))+1)/2];
+            fade = reshape(fade,1,1,[]);
+            ir = ir(:, :, 1:params.IRmax);
+            HRTFs(iHRTF).SOFA.Data.IR = ir .* fade;
+        end
+
+        if ~isKey(HRTFGroupData, HRTFs(iHRTF).HRTFGroup)
+            HRTFGroupData(HRTFs(iHRTF).HRTFGroup) = [];
+        end
+
+        HRTFGroupData(HRTFs(iHRTF).HRTFGroup) = [...
[…]

### 2024-06-19 src/spatialize/normalizeAudioTracks.m

@@ -1,0 +1,7 @@
+function tracks = normalizeAudioTracks(tracks, params)
+    for iTrack = 1:size(tracks,2)
+        tracks(:, iTrack) = normalizeLoudness(tracks(:, iTrack), ...
+             params.RecordingsExpectedFs, ...
+             params.TargetTrackLoudness);
+    end
+end

### 2024-06-19 src/spatialize/normalizeLoudness.m

@@ -1,0 +1,8 @@
+function signal = normalizeLoudness(signal, fs, targetLoudness)
+    loudness = integratedLoudness(signal, fs);
+    while abs(loudness - targetLoudness) > 0.001
+        gain = 10^((targetLoudness - loudness)/20);
+        signal = signal .* gain;
+        loudness = integratedLoudness(signal, fs);
+    end
+end

### 2024-06-19 src/spatialize/resampleAndSave.m

@@ -1,0 +1,22 @@
+% Resample and save routine
+function HRTF = resampleAndSave(HRTF, params)
+    fprintf('[%s][%s] Resampling from %d Hz to %d Hz\n', ...
+        HRTF.HRTFGroup, HRTF.Name, ...
+        HRTF.SOFA.Data.SamplingRate, ...
+        params.RecordingsExpectedFs);
+
+    HRTF.SOFA = SOFAresample(HRTF.SOFA, params.RecordingsExpectedFs);
+
+    resampledSOFAdir = fullfile(params.HRTFBaseDir, ...
+        ['_resampled_' num2str(params.RecordingsExpectedFs)], ...
+        HRTF.HRTFGroup);
+    resampledSOFAfilename = ['_resampled_' ...
+        num2str(params.RecordingsExpectedFs) '_' HRTF.Name];
+
+    if ~exist(resampledSOFAdir, 'dir')
+        mkdir(resampledSOFAdir);
+    end
+
+    fullSOFAfilename = fullfile(resampledSOFAdir, resampledSOFAfilename);
+    HRTF.SOFA = SOFAsave(fullSOFAfilename, HRTF.SOFA, 0);
+end

### 2024-06-19 src/spatialize/spatialize_front.m

@@ -1,0 +1,466 @@
+% Paweł Antoniuk 2021
+% Bialystok University of Technology
+% with the help of S. Zieliński
+
+%% Initialize
+clearvars; close all; clc;
+addpath('SOFA/API_MO');
+SOFAstart;
+SOFAgetVersion()
+% parpool(8)
+
+%% Params
+params.HRTFBaseDir = '/run/media/pawel/data/eurosip-width-features/spatialize/medHRTFs';
+params.RecordingsBaseDir = '/run/media/pawel/data/eurosip-width-features/recordings-all';
+params.FinalResultsOutputDir = '../spatresults';
+params.SpatOutputDir = [params.FinalResultsOutputDir filesep 'spat'];
+params.RecordingsExpectedFs = 48000;
+params.RecordingLoadRange = [2.5 inf];
+params.RecordingSpatRange = [0.5 7];
+params.RecordingFadeTime = [0.01 0.01];
+params.RecordingLevelScale = 0.9;
+params.NChannels = 2;
+params.Elevations = [0];
+params.IRmax = 3 * 512;
+params.FadeDuration = 2*10^-3;
+params.TargetTrackLoudness = -23; % db
+params.MaxWidth = 45; % from ensemble center
+params.NRepetitions = 4;
+params.AzimuthOffsetRange = [-45 45];
+params.InverseAzimuthHRTFGroups = ["cipic"];
+
+%% Load HRTFs and get audio filenames
+HRTFs = loadHRTFs(params);
+audioFilenames = dirWithoutDots(params.RecordingsBaseDir);
+
+%% Spatialize songs
+nAudioFilenames = length(audioFilenames);
+allSpatMetaresults = cell(nAudioFilenames, params.NRepetitions);
+
+start = tic;
+for iRep = 1:params.NRepetitions
+    parfor iAudioFilename = 1:nAudioFilenames       
+        %% Load audio tracks and spatialize them 
+        audioFilename = audioFilenames(iAudioFilename);
+        [tracks,trackNames] = loadAudioTracks(audioFilename, params);
+        tracks = normalizeAudioTracks(tracks, params);
+        [spatResults,spatMetaresults] = spatializeSong(HRTFs, tracks, ...
+            trackNames, audioFilename.name, params);
+        allSpatMetaresults{iAudioFilename,iRep} = spatMetaresults;
+        
+        %% Postprocess results
[…]

### 2024-06-19 src/spatialize/spatialize_front_noise.m

@@ -1,0 +1,464 @@
+% Paweł Antoniuk 2021
+% Bialystok University of Technology
+% with the help of S. Zieliński
+
+%% Initialize
+clearvars; close all; clc;
+addpath('SOFA/API_MO');
+SOFAstart;
+SOFAgetVersion()
+% parpool(8)
+
+%% Params
+params.HRTFBaseDir = '/run/media/pawel/data/eurosip-width-features/spatialize/medHRTFs';
+params.RecordingsBaseDir = '../../data/spatialize/recordings'
+params.FinalResultsOutputDir = '../../data/spatialize/spatresults';
+params.SpatOutputDir = [params.FinalResultsOutputDir filesep 'spat'];
+params.RecordingsExpectedFs = 48000;
+params.RecordingLoadRange = [2.5 inf];
+params.RecordingSpatRange = [0.5 7];
+params.RecordingFadeTime = [0.01 0.01];
+params.RecordingLevelScale = 0.9;
+params.NChannels = 2;
+params.Elevations = [0];
+params.IRmax = 3 * 512;
+params.FadeDuration = 2*10^-3;
+params.TargetTrackLoudness = -23; % db
+params.MaxWidth = 45; % from ensemble center
+params.NRepetitions = 4;
+params.AzimuthOffsetRange = [-45 45];
+params.InverseAzimuthHRTFGroups = ["cipic"];
+
+%% Load HRTFs and get audio filenames
+HRTFs = loadHRTFs(params);
+audioFilenames = dirWithoutDots(params.RecordingsBaseDir);
+
+%% Spatialize songs
+nAudioFilenames = length(audioFilenames);
+allSpatMetaresults = cell(nAudioFilenames, params.NRepetitions);
+
+start = tic;
+for iRep = 1:params.NRepetitions
+    parfor iAudioFilename = 1:nAudioFilenames       
+        %% Load audio tracks and spatialize them 
+        audioFilename = audioFilenames(iAudioFilename);
+        [tracks,trackNames] = loadAudioTracks(audioFilename, params);
+        tracks = normalizeAudioTracks(tracks, params);
+        [spatResults,spatMetaresults] = spatializeSong(HRTFs, tracks, ...
+            trackNames, audioFilename.name, params);
+        allSpatMetaresults{iAudioFilename,iRep} = spatMetaresults;
+        
+        %% Postprocess results
+        spatResults = posprocessSpatResults(spatResults, params);
[…]

### 2024-06-19 src/spatialize/spatialize_front_noise_sym.m

@@ -1,0 +1,436 @@
+% Paweł Antoniuk 2021
+% Bialystok University of Technology
+% with the help of S. Zieliński
+
+%% Initialize
+clearvars; close all; clc;
+addpath('SOFA/API_MO');
+SOFAstart;
+SOFAgetVersion()
+% parpool(8)
+
+%% Params
+params.HRTFBaseDir = '/run/media/pawel/data/eurosip-width-features/spatialize/medHRTFs';
+params.RecordingsBaseDir = '../../data/spatialize/recordings';
+params.FinalResultsOutputDir = '../../data/spatialize/spatresults';
+params.SpatOutputDir = [params.FinalResultsOutputDir filesep 'spat'];
+params.RecordingsExpectedFs = 48000;
+params.RecordingLoadRange = [2.5 inf];
+params.RecordingSpatRange = [0.5 7];
+params.RecordingFadeTime = [0.01 0.01];
+params.RecordingLevelScale = 0.9;
+params.NChannels = 2;
+params.Elevations = [0];
+params.IRmax = 3 * 512;
+params.FadeDuration = 2*10^-3;
+params.TargetTrackLoudness = -23; % db
+params.MaxWidth = 45; % from ensemble center
+params.NRepetitions = 10;
+params.NTracks = 10;
+params.AzimuthOffsetRange = [-45 45];
+params.InverseAzimuthHRTFGroups = ["cipic"];
+params.SymmetricalWidths = [0 5 10 15 20 25 30 35 40 45];
+
+%% Load HRTFs and get audio filenames
+HRTFs = loadHRTFs(params);
+
+%% Spatialize songs
+allSpatMetaresults = cell(1, params.NRepetitions);
+
+start = tic;
+parfor iRep = 1:params.NRepetitions     
+    %% Load audio tracks and spatialize them 
+    [tracks,trackNames] = loadAudioTracks(params);
+    tracks = normalizeAudioTracks(tracks, params);
+    audioFilename = 'Noise';
+    [spatResults,spatMetaresults] = spatializeSong(HRTFs, tracks, ...
+        trackNames, audioFilename, params);
+    allSpatMetaresults{iRep} = spatMetaresults;
+    
+    %% Postprocess results
+    spatResults = posprocessSpatResults(spatResults, params);
+    
+    %% Save results
+    saveSpatResults(spatResults, spatMetaresults, HRTFs, ...
+        audioFilename, iRep, params);
+    
[…]

### 2024-06-19 src/spatialize/spatialize_front_noise_sym_loc.m

@@ -1,0 +1,446 @@
+% Paweł Antoniuk 2021
+% Bialystok University of Technology
+% with the help of S. Zieliński
+
+%% Initialize
+clearvars; close all; clc;
+addpath('SOFA/API_MO');
+SOFAstart;
+SOFAgetVersion()
+% parpool(8)
+
+%% Params
+params.HRTFBaseDir = '/run/media/pawel/data/eurosip-width-features/spatialize/medHRTFs';
+params.RecordingsBaseDir = '../../data/spatialize/recordings';
+params.FinalResultsOutputDir = '../../data/spatialize/spatresults-location';
+params.SpatOutputDir = [params.FinalResultsOutputDir filesep 'spat'];
+params.RecordingsExpectedFs = 48000;
+params.RecordingLoadRange = [2.5 inf];
+params.RecordingSpatRange = [0.5 7];
+params.RecordingFadeTime = [0.01 0.01];
+params.RecordingLevelScale = 0.9;
+params.NChannels = 2;
+params.Elevations = [0];
+params.IRmax = 3 * 512;
+params.FadeDuration = 2*10^-3;
+params.TargetTrackLoudness = -23; % db
+params.MaxWidth = 45; % from ensemble center
+params.NRepetitions = 10;
+params.NTracks = 10;
+% params.AzimuthOffsetRange = [-45 45];
+params.InverseAzimuthHRTFGroups = ["cipic"];
+params.SymmetricalWidths = [30];
+params.Locations = [-30 -15 0 15];
+
+%% Load HRTFs and get audio filenames
+HRTFs = loadHRTFs(params);
+
+%% Spatialize songs
+allSpatMetaresults = cell(1, params.NRepetitions);
+
+start = tic;
+for iRep = 1:params.NRepetitions     
+    %% Load audio tracks and spatialize them 
+    [tracks,trackNames] = loadAudioTracks(params);
+    tracks = normalizeAudioTracks(tracks, params);
+    audioFilename = 'Noise';
+    [spatResults,spatMetaresults] = spatializeSong(HRTFs, tracks, ...
+        trackNames, audioFilename, params);
+    allSpatMetaresults{iRep} = spatMetaresults;
+    
+    %% Postprocess results
+    spatResults = posprocessSpatResults(spatResults, params);
+    
+    %% Save results
+    saveSpatResults(spatResults, spatMetaresults, HRTFs, ...
+        audioFilename, iRep, params);
+    
[…]

### 2024-06-19 src/train/model.py

@@ -81,6 +81,7 @@ class Model:
         self.x_val_phase: Optional[np.memmap] = None
         self.x_test_mag: Optional[np.memmap] = None
         self.x_test_phase: Optional[np.memmap] = None
+        self.__model: Optional[keras.Model] = None
 
         self.store_split: DotMap = DotMap()
         self.store_extract: DotMap = DotMap()
@@ -195,7 +196,7 @@ class Model:
             with ThreadPoolExecutor(max_workers=2) as executor:
                 try:
                     futures = [
-                        executor.submit(self._extract_features, filepath)
+                        executor.submit(self.__extract_features_from_filepath, filepath)
                         for filepath in self.filepaths
                     ]
                     for i, future in enumerate(futures):
@@ -212,7 +213,7 @@ class Model:
                     raise e
         else:
             for i, filepath in enumerate(self.filepaths):
-                result = self._extract_features(filepath)
+                result = self.__extract_features_from_filepath(filepath)
                 to_execute(i, result)
 
     def try_load_or_compute_input_data(self) -> Model:
@@ -428,48 +429,7 @@ class Model:
         print("[split] [5/5] Normalizing data")
         start_time = timeit.default_timer()
 
-        x_train_mag_min = np.min(self.x_train_mag)
-        x_train_mag_max = np.max(self.x_train_mag)
-        x_train_mag_mean = np.mean(self.x_train_mag)
-        x_train_mag_std = np.std(self.x_train_mag)
-        x_train_phase_min = np.min(self.x_train_phase)
-        x_train_phase_max = np.max(self.x_train_phase)
-        x_train_gcc_phat_min = np.min(self.store_split.x_train_gcc_phat)
-        x_train_gcc_phat_max = np.max(self.store_split.x_train_gcc_phat)
-        x_train_gcc_phat_mean = np.mean(self.store_split.x_train_gcc_phat)
-        x_train_gcc_phat_std = np.std(self.store_split.x_train_gcc_phat)
-
-        self.x_train_mag[:] = (self.x_train_mag - x_train_mag_mean) / x_train_mag_std
-        self.x_val_mag[:] = (self.x_val_mag - x_train_mag_mean) / x_train_mag_std
-        self.x_test_mag[:] = (self.x_test_mag - x_train_mag_mean) / x_train_mag_std
-
-        self.x_train_phase[:] = (self.x_train_phase - x_train_phase_min) / (
-            x_train_phase_max - x_train_phase_min
-        )
[…]

### 2024-06-19 paper/main.tex
Update paper

@@
 \usepackage{graphicx}
-\usepackage[colorlinks=true, allcolors=blue]{hyperref}
 \usepackage{gensymb}
@@
 \usepackage[toc,page]{appendix}
+\usepackage[colorlinks=true, allcolors=blue]{hyperref}
+\usepackage{multirow}
+\usepackage{tabularx}
 
@@
 
-The advancement of sophisticated machine learning techniques, particularly deep learning networks, has prompted an intriguing exploration of the extent to which these tools can emulate the human auditory system without relying on advanced spatial audio feature engineering, traditionally employed in audio source localization tasks \cite{yang_deepear_2022, vera-diaz_towards_2018, pang_multitask_2019}.
-To investigate this, the study developed an audio localization method based on a multi-task convolutional neural network (CNN) model \cite{lecun_handwritten_1989, lecun_deep_2015}.
+The advancement of sophisticated machine learning techniques, particularly deep learning networks, has spurred an intriguing exploration into the extent to which these tools can emulate the human auditory system.
+This exploration proceeds without reliance on the advanced spatial audio feature engineering traditionally used in audio source localization, as evidenced by references \cite{yang_deepear_2022, vera-diaz_towards_2018, pang_multitask_2019}.
+While the application of convolutional neural networks (CNNs) \cite{lecun_handwritten_1989} to audio signals --- especially in conjunction with spectrograms \cite{thomas_analyzing_2014, espi_exploiting_2015} or other techniques \cite{abdel-hamid_applying_2012, sainath_deep_2013} --- is not novel, it has been successfully employed in prior studies.
+Building on these foundations, this study developed an audio localization method by creating a multi-task CNN model.
 
@@
 Furthermore, the approach is not limited to speech and has been applied to a wide range of musical datasets, including instruments and vocals.
-In contrast to previous studies that primarily focused on individual sources, the proposed method does not aim to separate them, but rather considers them as an ensemble, or in this case, a musical ensemble.
-This approach is similar to how the real musical ensembles are arranged on stage.
+In contrast to previous studies that primarily focused on individual sources, the proposed method does not aim to separate them, but rather considers them as a group, or in this case, a musical ensemble.
+This approach is similar to how the real musical ensembles are arranged on stage, thereby emphasizing practical applications in live settings.
 To the authors' knowledge, this is one of the first methods to localize ensemble width (see \cite{antoniuk2023blind} for the previous ensemble-width-related study), and the first to localize both ensemble position and width simultaneously using a multi-task model.
@@
 
-In contrast, black-box methods typically employ minimal feature engineering and heavily rely on modern machine learning techniques to extract features and make predictions.
-These methods implicitly and internally extract features, without disclosing their purposes or mimicking the human auditory system.
-This lack of transparency and predictability, combined with their reliance on deep neural networks with numerous learning parameters, necessitates the development and evaluation using very large datasets.
-These datasets often comprise thousands of examples (e.g., 6300 examples in the TIMIT corpus \cite{garofolo_john_s_timit_1993} used in \cite{yang_deepear_2022, benaroya_binaural_2018, wang_binaural_2020, vecchiotti19, ma_exploiting_2017, pang_multitask_2019, ma_robust_2018, may_robust_2015}) or custom corpora consisting of hundreds of thousands of recordings used in \cite{antoniuk2023blind, zielinski_automatic_2022, zielinski_spatial_2022, zielinski_comparison_2020}.
+In contrast, black-box methods generally employ minimal feature engineering, relying instead on advanced machine learning techniques to autonomously extract features and make predictions.
+These methods, while effective, do not explicitly reveal the features they extract nor necessarily mimic the human auditory system.
+This opacity and the unpredictable nature of the outcomes, coupled with their dependency on deep neural networks that have extensive learning parameters, require the use of very large datasets for development and evaluation.
+Such datasets often include thousands of examples, as seen in the TIMIT corpus with 6300 examples \cite{garofolo_john_s_timit_1993} used in multiple studies \cite{yang_deepear_2022, benaroya_binaural_2018, wang_binaural_2020, vecchiotti19, ma_exploiting_2017, pang_multitask_2019, ma_robust_2018, may_robust_2015}, or custom corpora comprising hundreds of thousands of recordings \cite{antoniuk2023blind, zielinski_automatic_2022, zielinski_spatial_2022, zielinski_comparison_2020}.
+This poses a significant challenge in assembling a sufficiently large and diverse collection of labeled binaural recordings.
+However, this challenge can be addressed through the synthesis of binaural sounds, as demonstrated in various studies \cite{antoniuk2023blind, zielinski_automatic_2022, zielinski_spatial_2022, zielinski_comparison_2020,yang_deepear_2022, ma_robust_2018} and further elaborated upon in Section \ref{subsec:synthesis} of this paper.
 
@@
 The experiments conducted in this study involved 23,040 binaural recordings of music.
-The binaural recordings were synthesized using 192 multi-track publicly-available music recordings \cite{noauthor_mixing_nodate} and 30 HRTF databases (see \nameref{appendix:a_hrtf} for a detailed list of HRTF databases used in this study).
+The binaural recordings were synthesized using 192 multi-track publicly-available music recordings \cite{noauthor_mixing_nodate} and 30 HRTF databases (see Table \ref{table:hrtfs} in \nameref{appendix:a_hrtf} for a detailed list of HRTF databases used in this study).
 The number of tracks in multi-track recording ranged from 5 to 62, with median of 9.
@@
 Furthermore, the tracks of the input multi-track recordings were randomly assigned to sound source positions ($\theta_i$) to enhance the diversity of the final binaural corpora.
+Before the synthesis, the tracks were equalized to $-23$ LUFS, in accordance with the ITU-R BS.1770 recommendation \cite{noauthor_recommendation_nodate}.
 
@@
 
-The network topology used in this study was highly inspired by a AlexNet convolutional neural network \cite{krizhevsky_imagenet_2012}.
-Although AlexNet network was designed for image classification, the binaural recordings in this study were converted into magnitude spectrograms, as described in Section \ref{subsec:feature_extraction}.
-The network is consisting of a series of convolutional units followed by a series of classification units.
-
 \begin{figure}[ht]
@@
 
+The network topology employed in this study, depicted in Figure \ref{fig:architecture}, was significantly influenced by the AlexNet convolutional neural network \cite{krizhevsky_imagenet_2012}.
+Although AlexNet was originally designed for image classification, this study adapted it for the audio prediction task by converting binaural recordings into magnitude spectrograms, as described in Section \ref{subsec:feature_extraction}.
+This conversion allowed the spectrograms to be treated as images, enabling their use in a standard image-prediction-like task.
+
+The network accepts two spectrograms as inputs, and its architecture consists of a series of convolutional units followed by classification units, culminating in two output heads for predicting ensemble width and location.
+The architecture was designed in a multi-task fashion, enabling a single network to predict both ensemble parameters simultaneously.
+This multi-task capability is achieved by connecting the output of the final classification unit to two separate outputs, one for each ensemble parameter.
+
+Despite the availability of widely used techniques for addressing overfitting, such as the Dropout Layer \cite{srivastava_dropout_nodate}, and for accelerating training, such as Batch Normalization \cite{ioffe_batch_2015}, neither technique was employed in this study due to their observed ineffectiveness for the specific prediction task at hand.
+Instead, a Global Average Pooling Layer \cite{lin_network_2014} was utilized, known for its proven capabilities in reducing overfitting.
+In this particular task, the inclusion of this layer significantly reduced overfitting, lowering the final mean absolute error score by $0.83\degree$ (average across 10 trials) compared to configurations using a Flattening Layer.
+
 \subsection{Model training and evaluation}
 \label{subsec:training_evaluation}
+
+Monte-Carlo cross validation (without replacement).
 
@@
 
+{\renewcommand{\arraystretch}{1.1}
+\begin{table}[!h]
+  \begin{center}
+    \caption{\label{table:hrtfs}List of HRTF sets used to synthesize binaural audio excerpts}
+    \begin{tabularx}{\linewidth}{
+      |>{\hsize=0.3\hsize}X|
+      >{\hsize=0.7\hsize}X|
+      >{\raggedright\hsize=1.3\hsize}X|
+      >{\raggedright\hsize=0.6\hsize}X|
+      >{\raggedright\arraybackslash\hsize=2.1\hsize}X|
+      >{\hsize=1\hsize}X|
+      }
+      \cline{1-6}
+      \textbf{No.} & \textbf{Type} & \textbf{Head}                             & \textbf{Radius {[}m{]}} & \textbf{Source}                                                                                                                                             & \textbf{Acronym}                \\
+      \cline{1-6}
+      1.
+& Human         & Human subject                             & 1.2                     & \multirow{2}{\hsize}{RWTH Aachen University \cite{braren_high-resolution_2020}}                                                                             & \multirow{2}{\hsize}{AACHEN}    \\
+      \cline{1-4}
+      2.
+& Artificial    & GRAS 45BB-4 KEMAR                         & 1                       &                                                                                                                                                             &                                 \\
+      \cline{1-6}
+      3.
+& Human         & Subject 2                                 & 1.2                     & \multirow{4}{\hsize}{Austrian Academy of Sciences \cite{noauthor_hrtf-database_nodate}}                                                                     & \multirow{4}{\hsize}{ARI}       \\
+      \cline{1-4}
+      4.
+& Human         & Subject 4                                 & 1.2                     &                                                                                                                                                             &                                 \\
+      \cline{1-4}
+      5.
+& Human         & Subject 10                                & 1.2                     &                                                                                                                                                             &                                 \\
+      \cline{1-4}
+      6.
+& Artificial    & ARI Printed Head                          & 1.2                     &                                                                                                                                                             &                                 \\
+      \cline{1-6}
+      7.
+& Human         & Subject 012                               & 1                       & \multirow{3}{\hsize}{CIPIC Interface Laboratory, University of California \cite{algazi_cipic_2001}}                                                         & \multirow{3}{\hsize}{CIPIC}     \\
+      \cline{1-4}
+      8.
+& Human         & Subject 015                               & 1                       &                                                                                                                                                             &                                 \\
+      \cline{1-4}
+      9.
+& Human         & Subject 020                               & 1                       &                                                                                                                                                             &                                 \\
+      \cline{1-6}
+      10.
+& Artificial    & Neumann KU 100                            & 0.9                     & NASA (2007) \cite{andreopoulou_inter-laboratory_2015}                                                                                                       & \multirow{2}{\hsize}{CLUBFRITZ} \\
+      \cline{1-6}
+      11.
+& Artificial    & Neumann KU 100                            & 1.5                     & Helsinki University of Technology (2009) \cite{andreopoulou_inter-laboratory_2015}                                                                          &                                 \\
+      \cline{1-6}
+      12.
+& Artificial    & FABIAN                                    & 1.47                    & \multirow{4}{\hsize}{Technical University Berlin, Huawei Technologies, Munich Research Centre, Sennheiser Electronic \cite{brinkmann_cross-evaluated_2019}} & \multirow{4}{\hsize}{HUTUBS}    \\
+      \cline{1-4}
+      13.
+& Human         & Subject pp2                               & 1.47                    &                                                                                                                                                             &                                 \\
+      \cline{1-4}
+      14.
+& Human         & Subject pp3                               & 1.47                    &                                                                                                                                                             &                                 \\
+                   &               &                                           &                         &                                                                                                                                                             &                                 \\
+      \cline{1-6}
+      15.
+& Human         & Subject 1003                              & 1.95                    & \multirow{2}{\hsize}{IRCAM, AKG \cite{noauthor_listen_2023}}                                                                                                & \multirow{2}{\hsize}{LISTEN}    \\
+      \cline{1-4}
+      16.
+& Human         & Subject 1002                              & 1.95                    &                                                                                                                                                             &                                 \\
+      \cline{1-6}
+      17.
+& Artificial    & KEMAR DB-4004 (DB-061)                    & 1.4                     & \multirow{2}{\hsize}{MIT \cite{gardne_hrtf_1994}}                                                                                                           & \multirow{2}{\hsize}{MIT}       \\
+      \cline{1-4}
+      18.
+& Artificial    & KEMAR DB-4004 (DB-065)                    & 1.4                     &                                                                                                                                                             &                                 \\
+      \cline{1-6}
+      19.
+& Human         & Subject 001                               & 1.5                     & \multirow{3}{\hsize}{Tohoku University \cite{watanabe_dataset_2014}}                                                                                        & \multirow{3}{\hsize}{RIEC}      \\
+      \cline{1-4}
+      20.
+& Human         & Subject 002                               & 1.5                     &                                                                                                                                                             &                                 \\
+      \cline{1-4}
+      21.
+& Artificial    & Koken SAMRAI                              & 1.5                     &                                                                                                                                                             &                                 \\
+      \cline{1-6}
+      22.
+& Artificial    & Neumann KU 100                            & 1.2                     & \multirow{3}{\hsize}{University of York \cite{armstrong_perceptual_2018}}                                                                                   & \multirow{3}{\hsize}{SADIE II}  \\
+      \cline{1-4}
+      23.
+& Human         & Subject H3                                & 1.2                     &                                                                                                                                                             &                                 \\
+      \cline{1-4}
+      24.
+& Human         & Subject H4                                & 1.2                     &                                                                                                                                                             &                                 \\
+      \cline{1-6}
+      25.
+& Artificial    & KEMAR                                     & 1                       & South China University of Technology \cite{yu_near-field_2018}                                                                                              & SCUT                            \\
+      \cline{1-6}
+      26.
+& Artificial    & Neumann KU 100                            & 1                       & TH Köln  \cite{porschmann_spherical_2017}                                                                                                                   & TH Köln                         \\
+      \cline{1-6}
+      27.
+& Artificial    & FABIAN                                    & 1.7                     & \multirow{2}{\hsize}{TU Berlin \cite{brinkmann_high_2017, wierstorf_free_2011}}                                                                             & \multirow{2}{\hsize}{TU Berlin} \\
+      \cline{1-4}
+      28.
+& Artificial    & GRAS 45BA KEMAR                           & 1                       &                                                                                                                                                             &                                 \\
+      \cline{1-6}
+      29.
+& Artificial    & GRAS 45BB-4 KEMAR - subject A attachment  & 1                       & \multirow{3}{\hsize}{Aalborg University; University of Iceland \newline \cite{spagnol_viking_2019,spagnol_viking_2020}}                                     & \multirow{3}{\hsize}{VIKING}    \\
+      \cline{1-4}
+      30.
+& Artificial    & GRAS 45BB-4 KEMAR - subject B attachments & 1                       &                                                                                                                                                             &                                 \\
+                   &               &                                           &                         &                                                                                                                                                             &                                 \\
+      \cline{1-6}
+    \end{tabularx}
+  \end{center}
+\end{table}
+
 \printbibliography

### 2024-06-19 src/analyze/results.py

@@ -31,6 +31,7 @@ def load_results(model_dir: Path):
         score_file_test_width = score_df["test_score_width"] * 90
         score_file_test_location = score_df["test_score_location"] * 45
 
+        print(score_file_test_width)
         assert np.all(np.abs(score_file_test_width - score_test_width) < 1e-4)
         assert np.all(np.abs(score_file_test_location - score_test_location) < 1e-4)
 

### 2024-06-19 src/analyze/score_details.py

@@ -2,7 +2,7 @@ from pathlib import Path
 from results import get_results
 import numpy as np
 
-model_name = "multi_inout_1_model_final_3_nogcc_nosame"
+model_name = "final_width_location"
 
 
 def get_stats(results, filter = None) -> tuple[float, float, float, float]:

### 2024-06-19 src/train/architecture/multi_inout_1_flat.py

@@ -1,0 +1,33 @@
+from keras import layers as l
+from keras import models as m
+
+
+def cnn_block(input_layer):
+    x = l.Conv2D(32, kernel_size=(2, 2), activation='relu')(input_layer)
+    x = l.MaxPooling2D(pool_size=(2, 3))(x)
+    x = l.Conv2D(64, kernel_size=(2, 2), activation='relu')(x)
+    x = l.MaxPooling2D(pool_size=(2, 3))(x)
+    x = l.Conv2D(128, kernel_size=(2, 2), activation='relu')(x)
+    x = l.MaxPooling2D(pool_size=(2, 2))(x)
+    x = l.Conv2D(256, kernel_size=(2, 2), activation='relu')(x)
+    x = l.MaxPooling2D(pool_size=(2, 2))(x)
+    x = l.Flatten()(x)
+    return x
+
+
+def architecture(input_shape):
+    # input
+    mag_input_layer = l.Input(shape=input_shape, name='in_mag')
+    
+    # inner layers
+    mag_model = cnn_block(mag_input_layer)    
+    x = l.Dense(128, activation='relu')(mag_model)
+    x = l.Dense(64, activation='relu')(x)
+    x = l.Dense(32, activation='relu')(x)
+    x = l.Dense(16, activation='relu')(x)
+    
+    # output
+    output_width = l.Dense(1, name='out_width')(x)
+    output_location = l.Dense(1, name='out_location')(x)
+
+    return m.Model(inputs=[mag_input_layer], outputs=[output_width, output_location])

### 2024-06-23 paper/main.tex
Update reference styles, table format and software details in the paper

@@
 \usepackage{textcomp}
-\usepackage[backend=biber, style=nature]{biblatex}
+\usepackage[backend=biber, style=apa, maxcitenames=1, uniquelist=false, url=false]{biblatex}
 \usepackage{csquotes}
 \usepackage[toc,page]{appendix}
-\usepackage[colorlinks=true, allcolors=blue]{hyperref}
+\usepackage[colorlinks=true, allcolors=blue, breaklinks=true]{hyperref}
+\usepackage{xurl}
 \usepackage{multirow}
 \usepackage{tabularx}
+\usepackage{ltablex}
+\usepackage{color}
+\usepackage{tabularray}
+
+\defbibenvironment{bibliography}
+  {\begin{enumerate}}
+  {\end{enumerate}}
+  {\item}
+\AtEveryBibitem{%
+  \clearfield{note}%
+}
+
+% \AtEveryCitekey{\iffootnote{\csxdef{blx@footnotemark}{}\csxdef{blx@shortfootnotemark}{}}{}}
+
+% \DeclareCiteCommand{\cite}
+%   {\usebibmacro{prenote}}
+%   {\usebibmacro{citeindex}%
+%    \printtext[bibhyperref]{\usebibmacro{cite}}
+%   }
+%   {\multicitedelim}
+%   {\usebibmacro{postnote}}
+
+%   \DeclareCiteCommand{\parencite}
+%   {\usebibmacro{prenote}}
+%   {\usebibmacro{citeindex}%
+%    \printtext[bibhyperref]{\bibopenparen\usebibmacro{cite}\bibcloseparen}
+%   }
+%   {\multicitedelim}
+%   {\usebibmacro{postnote}}
+
+% % Redefine \parencites to handle multiple citations in one set of parentheses
+% \DeclareCiteCommand{\parencites}
+%   {\usebibmacro{prenote}}
+%   {\usebibmacro{citeindex}%
+%    \printtext[bibhyperref]{\bibopenparen\usebibmacro{cite}\bibcloseparen}
+%   }
+%   {\multicitedelim}
+%   {\usebibmacro{postnote}}
 
@@
 The human auditory system demonstrates exceptional proficiency in segregating, localizing, and interpreting diverse auditory signals, despite being limited to two ears.
-This capability arises from the intricate analysis of temporal, amplitude, and spectral disparities, a process termed binaural hearing \cite{blauert_spatial_1996}, which enables precise localization of sound sources in complex auditory environments.
-An important advantage of binaural hearing is demonstrated by the `cocktail party effect', which showcases the system's capacity to concentrate on individual sounds while suppressing background noise \cite{cherry_experiments_1953}.
-An understanding of the auditory system is essential for comprehending its limits and creating more immersive binaural experiences for entertainment purposes \cite{zhang_surround_2017}.
-It also helps to improve hearing aid devices by enhancing auditory signal reception and improving spatial awareness \cite{hirsh_binaural_1950, thiemann_speech_2016}.
+This capability arises from the intricate analysis of temporal, amplitude, and spectral disparities, a process termed binaural hearing \parencite{blauert_spatial_1996}, which enables precise localization of sound sources in complex auditory environments.
+An important advantage of binaural hearing is demonstrated by the `cocktail party effect', which showcases the system's capacity to concentrate on individual sounds while suppressing background noise \parencite{cherry_experiments_1953}.
+An understanding of the auditory system is essential for comprehending its limits and creating more immersive binaural experiences for entertainment purposes \parencite{zhang_surround_2017}.
+It also helps to improve hearing aid devices by enhancing auditory signal reception and improving spatial awareness \parencite{hirsh_binaural_1950, thiemann_speech_2016}.
 
-The advancement of sophisticated machine learning techniques, particularly deep learning networks, has spurred an intriguing exploration into the extent to which these tools can emulate the human auditory system.
-This exploration proceeds without reliance on the advanced spatial audio feature engineering traditionally used in audio source localization, as evidenced by references \cite{yang_deepear_2022, vera-diaz_towards_2018, pang_multitask_2019}.
-While the application of convolutional neural networks (CNNs) \cite{lecun_handwritten_1989} to audio signals --- especially in conjunction with spectrograms \cite{thomas_analyzing_2014, espi_exploiting_2015} or other techniques \cite{abdel-hamid_applying_2012, sainath_deep_2013} --- is not novel, it has been successfully employed in prior studies.
-Building on these foundations, this study developed an audio localization method by creating a multi-task CNN model.
+The advancement of sophisticated machine learning techniques, particularly deep learning networks, has prompted an intriguing exploration into the extent to which these tools can emulate the human auditory system.
+This exploration proceeds without reliance on the advanced spatial audio feature engineering traditionally used in audio source localization, as evidenced by some prominent recent examples \parencite{yang_deepear_2022, vera-diaz_towards_2018, pang_multitask_2019}.
+While the application of convolutional neural networks (CNNs) \parencite{lecun_handwritten_1989} --- traditionally employed in the visual domain --- to audio signals, particularly in conjunction with spectrograms \parencite{thomas_analyzing_2014, espi_exploiting_2015} or other feature engineering techniques \parencite{abdel-hamid_applying_2012, sainath_deep_2013}, is not novel.
+However, these approaches are continually being refined and adapted to the audio realm.
+Building on these foundations, this study developed an audio localization method by constructing a spectrogram-based multi-task CNN model.
 
-Inspired by the fact, that humans tend to localize sound sources in groups rather than individually \cite{bregman_auditory_1990, rumsey_spatial_2002}, the objective of the proposed model is to predict ensemble location and width instead of positions of individual sources.
-This study is unique as it not only conceptualized the method but also tested it on a vast, real-life music corpus of 23,040 binaural excerpts that were synthesized using 192 multi-track music recordings and 30 publicly available head-related transfer functions (HRTFs) from various sources.
+Inspired by the fact, that humans tend to localize sound sources in groups rather than individually \parencite{bregman_auditory_1990, rumsey_spatial_2002}, the objective of the proposed model is to predict ensemble location and width instead of positions of individual sources.
+This study is unique as it not only conceptualized the method but also tested it on a vast, real-life music corpus of 23,040 binaural excerpts that were synthesized using 192 multi-track music recordings \parencite{senior_mixing_2023} and 30 publicly available head-related transfer functions (HRTFs) from various sources (see Table \ref{table:hrtfs} in \ref{appendix:a_hrtf}).
 The music recordings encompassed various genres, including rock, jazz, pop, and classical music.
@@
 The findings demonstrate that this method is effective in accurately predicting the spatial characteristics of sound sources in near-real-world scenarios.
-Furthermore, this paper presents an experiment framework that allows for the objective measurement of a binaural localization technique in an objective manner, utilizing a large-scale dataset synthesized from real-world examples of music signals (for other use cases for this framework, see \cite{antoniuk2023blind, zielinski_automatic_2022, zielinski_spatial_2022, zielinski_comparison_2020}).
+Furthermore, this paper demonstrates an experimental framework that facilitates the objective measurement of a binaural localization technique, employing a large-scale dataset synthesized from real-world music signal (for applications of similar frameworks, see: \cite{antoniuk_blind_2023, zielinski_automatic_2022, zielinski_spatial_2022, zielinski_comparison_2020}).
 One of the key advantages of the proposed method is that it does not assume the number of audio sources.
-However, a significant limitation of this study is the lack of reverberation in the synthesized recordings, which represents a material for future research.
+However, significant limitations of this study include the absence of reverberation in the synthesized recordings and the method's inapplicability to real-time scenarios --- both are critical areas for future research.
 
@@
 
-The majority of existing literature on the subject of sound source localization employs techniques that leverage the advantages of microphone arrays with more than two channels \cite{kaveh_statistical_1986, pavlidi_real-time_2012, pan_multi-tone_2021, hahmann_sound_2022, chung_sound_2022, liu_sound_2022}.
-In the context of sound source localization in binaural signals, the predominant focus of research is on the identification of individual sound sources, rather than groups of sounds \cite{dietz_auditory_2011, may_probabilistic_2011, may_binaural_2012, woodruff_binaural_2012, may_robust_2015, ma16c_interspeech, ma_exploiting_2017, benaroya_binaural_2018}.
-In terms of source direction of arrival (DOA) methods, the majority of research assumes a fixed number of sound sources \cite{pang_multitask_2019, vecchiotti19, ma_exploiting_2017, woodruff_binaural_2012, s_spatiogram_2021}, which limits its practical applications as this information is rarely known in real-life binaural recordings.
-Moreover, the majority of studies have focused on relatively homogeneous signals, namely speech \cite{dietz_auditory_2011, may_probabilistic_2011, may_binaural_2012, woodruff_binaural_2012, may_robust_2015, ma16c_interspeech, ma_exploiting_2017, benaroya_binaural_2018, wang_binaural_2020, liu_multiple_2018, yang_deepear_2022, ma_robust_2018}.
+The majority of existing literature on the subject of sound source localization employs techniques that leverage the advantages of microphone arrays with more than two channels \parencite{kaveh_statistical_1986, pavlidi_real-time_2012, pan_multi-tone_2021, hahmann_sound_2022, chung_sound_2022, liu_sound_2022}.
+In the context of sound source localization in binaural signals, the predominant focus of research is on the identification of individual sound sources, rather than groups of sounds \parencite{dietz_auditory_2011, may_probabilistic_2011, may_binaural_2012, woodruff_binaural_2012, may_robust_2015, ma16c_interspeech, ma_exploiting_2017, benaroya_binaural_2018}.
+In terms of source direction of arrival (DOA) methods, the majority of research assumes a fixed number of sound sources \parencite{pang_multitask_2019, vera-diaz_towards_2018, ma_exploiting_2017, woodruff_binaural_2012, arthi_spatiogram_2021}, which limits its practical applications as this information is rarely known in real-life binaural recordings.
+Moreover, the majority of studies have focused on relatively homogeneous signals, namely speech \parencite{dietz_auditory_2011, may_probabilistic_2011, may_binaural_2012, woodruff_binaural_2012, may_robust_2015, ma16c_interspeech, ma_exploiting_2017, benaroya_binaural_2018, wang_binaural_2020, liu_multiple_2018, yang_deepear_2022, ma_robust_2018}.
 
@@
 This approach is similar to how the real musical ensembles are arranged on stage, thereby emphasizing practical applications in live settings.
-To the authors' knowledge, this is one of the first methods to localize ensemble width (see \cite{antoniuk2023blind} for the previous ensemble-width-related study), and the first to localize both ensemble position and width simultaneously using a multi-task model.
+To the authors' knowledge, this is one of the first methods to localize ensemble width (see \cite{antoniuk_blind_2023} for the previous ensemble-width-related study), and the first to localize both ensemble position and width simultaneously using a multi-task model.
 
-Sound localization methods can be classified into two categories based on the implementation of their underlying algorithms: glass-box (e.g., \cite{dietz_auditory_2011, may_probabilistic_2011, may_binaural_2012, may_robust_2015, woodruff_binaural_2012, ma16c_interspeech, ma_exploiting_2017, ma_robust_2018}) and black-box (e.g., \cite{vecchiotti19, vera-diaz_towards_2018, yang_deepear_2022}).
+Sound localization methods can be classified into two categories based on the implementation of their underlying algorithms: glass-box (e.g., \parencite{dietz_auditory_2011, may_probabilistic_2011, may_binaural_2012, may_robust_2015, woodruff_binaural_2012, ma16c_interspeech, ma_exploiting_2017, ma_robust_2018}) and black-box (e.g., \parencite{vera-diaz_towards_2018, vera-diaz_towards_2018, yang_deepear_2022}).
 Glass-box methods, more traditional in the literature, rely on manually designed algorithms that mimic the auditory system to explicitly extract key features for location prediction, such as interaural level differences, interaural time differences, interaural coherence, or interaural phase differences (see \cite{blauert_spatial_1996} for feature descriptions).
-An example of one of the most advanced auditory models that is able to extract such features was developed by the Two!Ears project \cite{raake_computational_nodate}.
+An example of one of the most advanced auditory models that is able to extract such features was developed by the Two!Ears project \parencite{raake_computational_nodate}.
 
@@
 This opacity and the unpredictable nature of the outcomes, coupled with their dependency on deep neural networks that have extensive learning parameters, require the use of very large datasets for development and evaluation.
-Such datasets often include thousands of examples, as seen in the TIMIT corpus with 6300 examples \cite{garofolo_john_s_timit_1993} used in multiple studies \cite{yang_deepear_2022, benaroya_binaural_2018, wang_binaural_2020, vecchiotti19, ma_exploiting_2017, pang_multitask_2019, ma_robust_2018, may_robust_2015}, or custom corpora comprising hundreds of thousands of recordings \cite{antoniuk2023blind, zielinski_automatic_2022, zielinski_spatial_2022, zielinski_comparison_2020}.
+Such datasets often include thousands of examples, as seen in the TIMIT corpus with 6300 examples \parencite{garofolo_darpa_1993} used in multiple studies \parencite{yang_deepear_2022, benaroya_binaural_2018, wang_binaural_2020, vera-diaz_towards_2018, ma_exploiting_2017, pang_multitask_2019, ma_robust_2018, may_robust_2015}, or custom corpora comprising hundreds of thousands of recordings \parencite{antoniuk_blind_2023, zielinski_automatic_2022, zielinski_spatial_2022, zielinski_comparison_2020}.
 This poses a significant challenge in assembling a sufficiently large and diverse collection of labeled binaural recordings.
-However, this challenge can be addressed through the synthesis of binaural sounds, as demonstrated in various studies \cite{antoniuk2023blind, zielinski_automatic_2022, zielinski_spatial_2022, zielinski_comparison_2020,yang_deepear_2022, ma_robust_2018} and further elaborated upon in Section \ref{subsec:synthesis} of this paper.
+However, this challenge can be addressed through the synthesis of binaural sounds, as demonstrated in various studies \parencite{antoniuk_blind_2023, zielinski_automatic_2022, zielinski_spatial_2022, zielinski_comparison_2020,yang_deepear_2022, ma_robust_2018} and further elaborated upon in Section \ref{subsec:synthesis} of this paper.
 
@@
 It should be noted that although humans possess some limited abilities to localize sound sources in the vertical plane, in this study all sources are placed in the horizontal plane, at the height of the listener.
-This covers the majority of cases for real-world recordings (see \cite{ma_robust_2018, zielinski_spatial_2022} for similar studies that cover top-down discrimination).
+This covers the majority of cases for real-world recordings (see \cite{ma_robust_2018, zielinski_spatial_2022} for related studies that cover top-down discrimination).
 
@@
 The experiments conducted in this study involved 23,040 binaural recordings of music.
-The binaural recordings were synthesized using 192 multi-track publicly-available music recordings \cite{noauthor_mixing_nodate} and 30 HRTF databases (see Table \ref{table:hrtfs} in \nameref{appendix:a_hrtf} for a detailed list of HRTF databases used in this study).
+The binaural recordings were synthesized using 192 multi-track publicly-available music recordings \parencite{senior_mixing_2023} and 30 HRTF databases (see Table \ref{table:hrtfs} in \nameref{appendix:a_hrtf} for a detailed list of HRTF databases used in this study).
 The number of tracks in multi-track recording ranged from 5 to 62, with median of 9.
@@
 Furthermore, the tracks of the input multi-track recordings were randomly assigned to sound source positions ($\theta_i$) to enhance the diversity of the final binaural corpora.
-Before the synthesis, the tracks were equalized to $-23$ LUFS, in accordance with the ITU-R BS.1770 recommendation \cite{noauthor_recommendation_nodate}.
+Before the synthesis, the tracks were equalized to $-23$ LUFS, in accordance with the \parencite{noauthor_itu-r_2023} recommendation.
 
-The binaural recordings were obtained in this study using the binaural synthesis procedure, known as binauralization, whose aim was to simulate the positions of sound sources within a virtual acoustic environment \cite{blauert_spatial_1996}.
+The binaural recordings were obtained in this study using the binaural synthesis procedure, known as binauralization, whose aim was to simulate the positions of sound sources within a virtual acoustic environment \parencite{blauert_spatial_1996}.
 This was achieved by convolving multi-track signals with head-related impulse responses from a specified head-related transfer function (HRTF) database.
@@
 % The procedure proposed in this study did not 
-
 % Zamiast tego dac obrazek z glowami i spektrogramami, lewy i prawi i roznica, sprawdzic czy to ma sens, potencjalnie usunac
@@
 % \end{figure}
-
 % Figure \ref{fig:samples} displays example spectrograms from four distinct binaural recordings, exhibiting a broad bandwidth and the diversity of the signals used in this study
@@
 Prior to input into the model, the binaural recordings of music were transformed into magnitude spectrograms.
-It is important to note that the original spectrograms were preserved in a floating-point precision matrix format, which prevented any loss of information due to precision conversion.
-The spectrograms were calculated using the Fast Fourier Transform (FFT) algorithm, with a limitation of 150 bands, spaced linearly from 100 Hz to 16 kHz.
+Although spectrograms do not directly provide information that can be translated into ensemble features, especially ensemble width, the goal of this task was to reduce the number of independent variables compared to the raw audio signal by extracting more compressed and informative data in the frequency domain.
+This step was necessary to decrease the likelihood of overfitting, reduce the number of examples needed to train the model, and thereby lower the overall computational power requirements.
+It is worth mentioning, however, that recent methods have shown that CNNs can be suitable for end-to-end audio localization without the spectrogram extraction step \parencite{vera-diaz_towards_2018, vecchiotti_end--end_2019}.
+
+On the computational side, the spectrograms were treated as three-dimensional floating-point precision feature matrices.
+The spectrograms were calculated using the Fast Fourier Transform (FFT) algorithm, with a limitation of 150 bands spaced linearly from 100 Hz to 16 kHz.
 Additionally, a Hamming window of 40 ms was applied to each frame of the signal, resulting in a total of 349 time frames.
@@
 Each sample was represented by a matrix of dimensions $2 \times 349 \times 150$.
-This method parallels the procedure presented in \cite{zielinski_automatic_2022}.
+This method parallels the procedure presented by \cite{zielinski_automatic_2022}.
 
@@
 
-The network topology employed in this study, depicted in Figure \ref{fig:architecture}, was significantly influenced by the AlexNet convolutional neural network \cite{krizhevsky_imagenet_2012}.
+The network topology employed in this study, depicted in Figure \ref{fig:architecture}, was significantly influenced by the AlexNet convolutional neural network demonstrated by \cite{krizhevsky_imagenet_2012}.
 Although AlexNet was originally designed for image classification, this study adapted it for the audio prediction task by converting binaural recordings into magnitude spectrograms, as described in Section \ref{subsec:feature_extraction}.
@@
 
-Despite the availability of widely used techniques for addressing overfitting, such as the Dropout Layer \cite{srivastava_dropout_nodate}, and for accelerating training, such as Batch Normalization \cite{ioffe_batch_2015}, neither technique was employed in this study due to their observed ineffectiveness for the specific prediction task at hand.
-Instead, a Global Average Pooling Layer \cite{lin_network_2014} was utilized, known for its proven capabilities in reducing overfitting.
+The topology finalized in this study was selected based on the highest prediction quality observed on the validation dataset.
+Despite the existence of numerous algorithms for automatic topology selection \parencite{branke_evolutionary_1995, zhang_finding_2018, miikkulainen_evolving_2017, stanley_evolving_2002, shafiee_deep_2016}, the final topology selected in this study was determined manually, based on its superior prediction quality observed on the validation dataset.
+This approach was necessitated primarily by the high computational demands compared to the available resources.
+
+Various architectural configurations were assessed, varying key parameters such as the number of convolutional layers (from 1 to 5), the number of classification units (from 1 to 5), the inclusion or exclusion of Max Pooling Layers after each Convolution Layer, the number of filters within the Convolution Layers, the dimensions of these filters ($2 \times 2$, $2 \times 3$, or $3 \times 3$), the stride size, and the dimensions of the Max Pooling Layers ($2 \times 2$, $2 \times 3$, or $3 \times 3$).
+Although a detailed account of the selection process for the topology is beyond the primary scope of this study --- particularly considering the task's complexity --- it was concluded that the Convolutional Neural Network (CNN) model is robust against variations in the assessed topologies.
+The differences in mean prediction error among the configurations were minimal, typically less than $1^\circ$ for most configurations.
+Among the many tested topologies that yielded almost identical errors, the simplest one was chosen to optimize both performance efficiency and model simplicity.
+
+Despite the availability of widely used techniques for addressing overfitting, such as the Dropout Layer \parencite{srivastava_dropout_2014}, and for accelerating training, such as Batch Normalization \parencite{ioffe_batch_2015}, neither technique was employed in this study due to their observed ineffectiveness for the specific prediction task at hand.
+Instead, a Global Average Pooling Layer \parencite{lin_network_2013} was utilized, known for its proven capabilities in reducing overfitting.
 In this particular task, the inclusion of this layer significantly reduced overfitting, lowering the final mean absolute error score by $0.83\degree$ (average across 10 trials) compared to configurations using a Flattening Layer.
@@
 
-Monte-Carlo cross validation (without replacement).
+The topology described previously resulted in a model with 216,562 learning parameters.
+This model was trained 10 times from scratch, employing the Monte Carlo cross-validation method \parencite{kuhn_applied_2013}.
+For each iteration, the entire dataset was randomly divided into two parts: a development set containing two-thirds of the dataset (15,360 recordings) for model construction, and a test set comprising the remaining one-third (7,680 recordings) for evaluation.
+The development set was further subdivided into training and validation subsets at a 7:1 ratio, with 13,440 recordings allocated to the training subset and 1,920 recordings to the validation subset.
+
+The training subset was employed to update the model's learning parameters, while the validation subset was utilized for implementing the early stopping technique \parencite{pocock_practical_1989, morgan_generalization_1989} in conjunction with model checkpointing \parencite{eisenman_check-n-run_2020}.
+These methods played a crucial role in selecting the model that demonstrated the best generalization capabilities.
+The test subset, containing data not previously encountered in the training or validation phases, was used exclusively for performance assessment once per iteration.
+This process was repeated to collect 10 mean absolute errors, from which the final model error was determined.
+The quantified mean absolute error and its standard deviation illustrated the model's prediction quality, generalization capabilities, and reproducibility.
+
+For each sample, the model received two spectrograms as input: one for the left channel and one for the right channel.
+The rationale behind the application of Convolutional Neural Networks (CNNs) to this task was to automatically extract local features from the spectrograms and use these features to predict two contiguous ensemble parameters: ensemble location and width, both measured in degrees.
+The prediction errors, calculated as the difference between the actual ensemble parameters (known from the binaural synthesis described in Section \ref{subsec:synthesis}) and the predicted values, were utilized in the Adam learning algorithm \parencite{kingma_adam_2014}.
+During training, the losses for both outputs were combined by summing them, thereby treating both ensemble features with equal importance.
+
+All computations in this study were conducted on a workstation equipped with an RTX Nvidia GeForce 4090 GPU and a 48-core AMD Ryzen ThreadRipper processor.
+The binaural recordings were synthesized using the MATLAB \parencite{MATLAB} environment with the Audio Toolbox \parencite{MATLAB_Audio_Toolbox}.
+Feature extraction and model training were performed in the Python \parencite{python}, utilizing the SciPy \parencite{2020SciPy-NMeth} library for feature extraction and Keras for training the CNN model.
+The complete source code for all stages of the process has been made available by the authors and is accessible to the public at GitHub \parencite{github_repo}.
+The feature extraction phase required 21 minutes, data partitioning took 34 minutes, and the total training time for all iterations was 40 minutes.
+In total, the entire training and evaluation process required 95 minutes.
 
@@
 
-{\renewcommand{\arraystretch}{1.1}
-\begin{table}[!h]
-  \begin{center}
-    \caption{\label{table:hrtfs}List of HRTF sets used to synthesize binaural audio excerpts}
-    \begin{tabularx}{\linewidth}{
-      |>{\hsize=0.3\hsize}X|
-      >{\hsize=0.7\hsize}X|
-      >{\raggedright\hsize=1.3\hsize}X|
-      >{\raggedright\hsize=0.6\hsize}X|
-      >{\raggedright\arraybackslash\hsize=2.1\hsize}X|
-      >{\hsize=1\hsize}X|
+\begin{longtblr}[
+  caption = {List of HRTF sets used to synthesize binaural audio excerpts},
+  label = {table:hrtfs}
+  ]{
+  colspec={c|c|>{\raggedright\arraybackslash}X|c|>{\raggedright\arraybackslash}X|c},
+  hlines = {},
+  vlines = {},
+  cell{2}{5} = {r=2}{m},
+  cell{2}{6} = {r=2}{m},
+  cell{4}{5} = {r=4}{m},
+  cell{4}{6} = {r=4}{m},
+  cell{8}{5} = {r=3}{m},
+  cell{8}{6} = {r=3}{m},
+  cell{11}{6} = {r=2}{m},
+  cell{13}{5} = {r=3}{m},
+  cell{13}{6} = {r=3}{m},
+  cell{16}{5} = {r=2}{m},
+  cell{16}{6} = {r=2}{m},
+  cell{18}{5} = {r=2}{m},
+  cell{18}{6} = {r=2}{m},
+  cell{20}{5} = {r=3}{m},
+  cell{20}{6} = {r=3}{m},
+  cell{23}{5} = {r=3}{m},
+  cell{23}{6} = {r=3}{m},
+  cell{28}{5} = {r=2}{m},
+  cell{28}{6} = {r=2}{m},
+  cell{30}{5} = {r=2}{m},
+  cell{30}{6} = {r=2}{m},
   }
-      \cline{1-6}
   \textbf{No.} & \textbf{Type} & \textbf{Head}                             & \textbf{Radius {[}m{]}} & \textbf{Source}                                                                                                                            & \textbf{Acronym} \\
-      \cline{1-6}
   1.
-& Human         & Human subject                             & 1.2                     & \multirow{2}{\hsize}{RWTH Aachen University \cite{braren_high-resolution_2020}}                                                                             & \multirow{2}{\hsize}{AACHEN}    \\
-      \cline{1-4}
+& Human         & Human subject                             & 1.2                     & RWTH Aachen University \parencite{braren_high-resolution_2020}                                                                             & AACHEN           \\
   2.
 & Artificial    & GRAS 45BB-4 KEMAR                         & 1                       &                                                                                                                                            &                  \\
-      \cline{1-6}
   3.
-& Human         & Subject 2                                 & 1.2                     & \multirow{4}{\hsize}{Austrian Academy of Sciences \cite{noauthor_hrtf-database_nodate}}                                                                     & \multirow{4}{\hsize}{ARI}       \\
-      \cline{1-4}
+& Human         & Subject 2                                 & 1.2                     & Austrian Academy of Sciences \parencite{noauthor_hrtf-database_nodate}                                                                     & ARI              \\
   4.
 & Human         & Subject 4                                 & 1.2                     &                                                                                                                                            &                  \\
-      \cline{1-4}
   5.
 & Human         & Subject 10                                & 1.2                     &                                                                                                                                            &                  \\
-      \cline{1-4}
   6.
 & Artificial    & ARI Printed Head                          & 1.2                     &                                                                                                                                            &                  \\
-      \cline{1-6}
   7.
-& Human         & Subject 012                               & 1                       & \multirow{3}{\hsize}{CIPIC Interface Laboratory, University of California \cite{algazi_cipic_2001}}                                                         & \multirow{3}{\hsize}{CIPIC}     \\
-      \cline{1-4}
+& Human         & Subject 012                               & 1                       & CIPIC Interface Laboratory, University of California \parencite{algazi_cipic_2001}                                                         & CIPIC            \\
   8.
 & Human         & Subject 015                               & 1                       &                                                                                                                                            &                  \\
-      \cline{1-4}
   9.
 & Human         & Subject 020                               & 1                       &                                                                                                                                            &                  \\
-      \cline{1-6}
   10.
-& Artificial    & Neumann KU 100                            & 0.9                     & NASA (2007) \cite{andreopoulou_inter-laboratory_2015}                                                                                                       & \multirow{2}{\hsize}{CLUBFRITZ} \\
-      \cline{1-6}
+& Artificial    & Neumann KU 100                            & 0.9                     & NASA (2007) \parencite{andreopoulou_inter-laboratory_2015}                                                                                 & CLUBFRITZ        \\
   11.
-& Artificial    & Neumann KU 100                            & 1.5                     & Helsinki University of Technology (2009) \cite{andreopoulou_inter-laboratory_2015}                                                                          &                                 \\
-      \cline{1-6}
+& Artificial    & Neumann KU 100                            & 1.5                     & Helsinki University of Technology (2009) \parencite{andreopoulou_inter-laboratory_2015}                                                    &                  \\
   12.
-& Artificial    & FABIAN                                    & 1.47                    & \multirow{4}{\hsize}{Technical University Berlin, Huawei Technologies, Munich Research Centre, Sennheiser Electronic \cite{brinkmann_cross-evaluated_2019}} & \multirow{4}{\hsize}{HUTUBS}    \\
-      \cline{1-4}
+& Artificial    & FABIAN                                    & 1.47                    & Technical University Berlin, Huawei Technologies, Munich Research Centre, Sennheiser Electronic \parencite{brinkmann_cross-evaluated_2019} & HUTUBS           \\
   13.
 & Human         & Subject pp2                               & 1.47                    &                                                                                                                                            &                  \\
-      \cline{1-4}
   14.
 & Human         & Subject pp3                               & 1.47                    &                                                                                                                                            &                  \\
-                   &               &                                           &                         &                                                                                                                                                             &                                 \\
-      \cline{1-6}
   15.
-& Human         & Subject 1003                              & 1.95                    & \multirow{2}{\hsize}{IRCAM, AKG \cite{noauthor_listen_2023}}                                                                                                & \multirow{2}{\hsize}{LISTEN}    \\
-      \cline{1-4}
+& Human         & Subject 1003                              & 1.95                    & IRCAM, AKG \parencite{noauthor_listen_2023}                                                                                                & LISTEN           \\
   16.
 & Human         & Subject 1002                              & 1.95                    &                                                                                                                                            &                  \\
-      \cline{1-6}
   17.
-& Artificial    & KEMAR DB-4004 (DB-061)                    & 1.4                     & \multirow{2}{\hsize}{MIT \cite{gardne_hrtf_1994}}                                                                                                           & \multirow{2}{\hsize}{MIT}       \\
-      \cline{1-4}
+& Artificial    & KEMAR DB-4004 (DB-061)                    & 1.4                     & MIT \parencite{gardne_hrtf_1994}                                                                                                           & MIT              \\
   18.
 & Artificial    & KEMAR DB-4004 (DB-065)                    & 1.4                     &                                                                                                                                            &                  \\
-      \cline{1-6}
   19.
-& Human         & Subject 001                               & 1.5                     & \multirow{3}{\hsize}{Tohoku University \cite{watanabe_dataset_2014}}                                                                                        & \multirow{3}{\hsize}{RIEC}      \\
-      \cline{1-4}
+& Human         & Subject 001                               & 1.5                     & Tohoku University \parencite{watanabe_dataset_2014}                                                                                        & RIEC             \\
   20.
 & Human         & Subject 002                               & 1.5                     &                                                                                                                                            &                  \\
-      \cline{1-4}
   21.
 & Artificial    & Koken SAMRAI                              & 1.5                     &                                                                                                                                            &                  \\
-      \cline{1-6}
   22.
-& Artificial    & Neumann KU 100                            & 1.2                     & \multirow{3}{\hsize}{University of York \cite{armstrong_perceptual_2018}}                                                                                   & \multirow{3}{\hsize}{SADIE II}  \\
-      \cline{1-4}
+& Artificial    & Neumann KU 100                            & 1.2                     & University of York \parencite{armstrong_perceptual_2018}                                                                                   & SADIE II         \\
   23.
 & Human         & Subject H3                                & 1.2                     &                                                                                                                                            &                  \\
-      \cline{1-4}
   24.
 & Human         & Subject H4                                & 1.2                     &                                                                                                                                            &                  \\
-      \cline{1-6}
   25.
-& Artificial    & KEMAR                                     & 1                       & South China University of Technology \cite{yu_near-field_2018}                                                                                              & SCUT                            \\
-      \cline{1-6}
+& Artificial    & KEMAR                                     & 1                       & South China University of Technology \parencite{yu_near-field_2018}                                                                        & SSCUT            \\
   26.
-& Artificial    & Neumann KU 100                            & 1                       & TH Köln  \cite{porschmann_spherical_2017}                                                                                                                   & TH Köln                         \\
-      \cline{1-6}
+& Artificial    & Neumann KU 100                            & 1                       & TH Köln  \parencite{porschmann_spherical_2017}                                                                                             & STH Köln         \\
   27.
-& Artificial    & FABIAN                                    & 1.7                     & \multirow{2}{\hsize}{TU Berlin \cite{brinkmann_high_2017, wierstorf_free_2011}}                                                                             & \multirow{2}{\hsize}{TU Berlin} \\
-      \cline{1-4}
+& Artificial    & FABIAN                                    & 1.7                     & TU Berlin \parencite{brinkmann_high_2017, wierstorf_free_2011}                                                                             & TU Berlin        \\
   28.
 & Artificial    & GRAS 45BA KEMAR                           & 1                       &                                                                                                                                            &                  \\
-      \cline{1-6}
   29.
-& Artificial    & GRAS 45BB-4 KEMAR - subject A attachment  & 1                       & \multirow{3}{\hsize}{Aalborg University; University of Iceland \newline \cite{spagnol_viking_2019,spagnol_viking_2020}}                                     & \multirow{3}{\hsize}{VIKING}    \\
-      \cline{1-4}
+& Artificial    & GRAS 45BB-4 KEMAR - subject A attachment  & 1                       & Aalborg University; University of Iceland \newline \parencite{spagnol_viking_2019,spagnol_viking_2020}                                     & VIKING           \\
   30.
 & Artificial    & GRAS 45BB-4 KEMAR - subject B attachments & 1                       &                                                                                                                                            &                  \\
-                   &               &                                           &                         &                                                                                                                                                             &                                 \\
-      \cline{1-6}
-    \end{tabularx}
-  \end{center}
-\end{table}
+\end{longtblr}
 

### 2024-06-24 paper/main.tex
Update paper

@@
 Consequently, there is now a need for automated and objective measurements of spatial content information, with ensemble location and width being the most prominent.
-This study presents a novel method for predicting ensemble location and width in binaural recordings.
+This study presents a novel method for predicting ensemble location and width in binaural recordings of music.
 To this end, 30 head-related transfer functions and 192 binaural music recordings from publicly accessible multi-track recording repositories were used to synthesize 23,040 binaural recordings.
@@
 The human auditory system demonstrates exceptional proficiency in segregating, localizing, and interpreting diverse auditory signals, despite being limited to two ears.
-This capability arises from the intricate analysis of temporal, amplitude, and spectral disparities, a process termed binaural hearing \parencite{blauert_spatial_1996}, which enables precise localization of sound sources in complex auditory environments.
-An important advantage of binaural hearing is demonstrated by the `cocktail party effect', which showcases the system's capacity to concentrate on individual sounds while suppressing background noise \parencite{cherry_experiments_1953}.
-An understanding of the auditory system is essential for comprehending its limits and creating more immersive binaural experiences for entertainment purposes \parencite{zhang_surround_2017}.
-It also helps to improve hearing aid devices by enhancing auditory signal reception and improving spatial awareness \parencite{hirsh_binaural_1950, thiemann_speech_2016}.
+This is possible by internal examination of interaural differences in time, loudness, and frequency, known as binaural hearing \parencite{blauert_spatial_1996}, which enables precise localization of sound sources in complex auditory environments.
+A notable advantage of binaural hearing is exemplified by the `cocktail party effect', highlighting the system's capacity to concentrate on foreground sound sources while suppressing background noise \parencite{cherry_experiments_1953}.
+
+Understanding of the auditory system is essential for comprehending its limits and creating more immersive binaural experiences for entertainment purposes \parencite{zhang_surround_2017}.
+It is also important for enhancing auditory signal reception in hearing aid devices \parencite{hirsh_binaural_1950, thiemann_speech_2016}.
 
@@
 This exploration proceeds without reliance on the advanced spatial audio feature engineering traditionally used in audio source localization, as evidenced by some prominent recent examples \parencite{yang_deepear_2022, vera-diaz_towards_2018, pang_multitask_2019}.
-While the application of convolutional neural networks (CNNs) \parencite{lecun_handwritten_1989} --- traditionally employed in the visual domain --- to audio signals, particularly in conjunction with spectrograms \parencite{thomas_analyzing_2014, espi_exploiting_2015} or other feature engineering techniques \parencite{abdel-hamid_applying_2012, sainath_deep_2013}, is not novel.
+While the application of convolutional neural networks (CNNs) \parencite{lecun_handwritten_1989} --- traditionally employed in the visual domain --- to audio signals, particularly in conjunction with spectrograms \parencite{thomas_analyzing_2014, espi_exploiting_2015, han_convolutional_2017} or other feature engineering techniques \parencite{abdel-hamid_applying_2012, sainath_deep_2013}, is not novel.
 However, these approaches are continually being refined and adapted to the audio realm.
@@
 This study is unique as it not only conceptualized the method but also tested it on a vast, real-life music corpus of 23,040 binaural excerpts that were synthesized using 192 multi-track music recordings \parencite{senior_mixing_2023} and 30 publicly available head-related transfer functions (HRTFs) from various sources (see Table \ref{table:hrtfs} in \ref{appendix:a_hrtf}).
-The music recordings encompassed various genres, including rock, jazz, pop, and classical music.
+The music recordings covered many different types of genres, including rock, jazz, pop, and classical music.
 
@@
 
-The developed method has the potential to be highly beneficial in automated assessment tasks, where a significant number of binaural recordings must be evaluated and labeled in terms of their spatial content information.
-This may assist audio engineers in objectively assessing and segregating binaural audio recordings with regard to their spatial content.
-Furthermore, it could facilitate the development of an autonomous web-crawler bot that will collect binaural recordings from publicly accessible repositories and label them according to the spatial properties of the sound sources, such as the location of the music ensemble or the sparsity of audio source positions.
+The developed method has the potential to be highly beneficial in automated assessment tasks, where a significant number of binaural recordings must be measured or labeled in terms of their spatial content information.
+This could be utilized in the development of a hypothetical autonomous web-crawler bot that will collect binaural recordings from publicly accessible repositories and label them according to the spatial properties of the sound sources, such as the location of the music ensemble or the sparsity of audio source positions.
+This method may also assist audio engineers in objectively assessing and segregating binaural audio recordings with regard to their spatial content.
 
-The remainder of this paper is organized as follows: Section \ref{sec:related-studies} presents related studies.
+The structure of this paper is organized as follows.
+Section \ref{sec:related-studies} presents related studies.
 The description of the method developed for this study is provided in Section \ref{sec:methodology}, which also includes detailed definitions of ensemble location and width, along with a description of the experiments used to evaluate this method.
-Section \ref{sec:results} discusses the results of the method-evaluation experiments conducted in this study.
-Finally, the paper concludes in Section \ref{sec:conclusions}.
+Section \ref{sec:results} presents and discusses the performance of the presented method and the results of the experiments conducted in this study.
+The paper concludes in Section \ref{sec:conclusions}.
 
@@
 The majority of existing literature on the subject of sound source localization employs techniques that leverage the advantages of microphone arrays with more than two channels \parencite{kaveh_statistical_1986, pavlidi_real-time_2012, pan_multi-tone_2021, hahmann_sound_2022, chung_sound_2022, liu_sound_2022}.
-In the context of sound source localization in binaural signals, the predominant focus of research is on the identification of individual sound sources, rather than groups of sounds \parencite{dietz_auditory_2011, may_probabilistic_2011, may_binaural_2012, woodruff_binaural_2012, may_robust_2015, ma16c_interspeech, ma_exploiting_2017, benaroya_binaural_2018}.
+These method simplify the task and can improve the localization performance; however, they do not utilize binaural hearing, rendering them ineffective for binaural recordings.
+
+In the context of sound source localization in binaural signals, the  focus of research is on the identification of individual sound sources, rather than groups of sounds \parencite{dietz_auditory_2011, may_probabilistic_2011, may_binaural_2012, woodruff_binaural_2012, may_robust_2015, ma16c_interspeech, ma_exploiting_2017, benaroya_binaural_2018}.
 In terms of source direction of arrival (DOA) methods, the majority of research assumes a fixed number of sound sources \parencite{pang_multitask_2019, vera-diaz_towards_2018, ma_exploiting_2017, woodruff_binaural_2012, arthi_spatiogram_2021}, which limits its practical applications as this information is rarely known in real-life binaural recordings.
@@
 
-On the computational side, the spectrograms were treated as three-dimensional floating-point precision feature matrices.
-The spectrograms were calculated using the Fast Fourier Transform (FFT) algorithm, with a limitation of 150 bands spaced linearly from 100 Hz to 16 kHz.
-Additionally, a Hamming window of 40 ms was applied to each frame of the signal, resulting in a total of 349 time frames.
+To prepare the input for the model, a Hamming window of 40 ms was applied to each frame of the signal, resulting in a total of 349 time frames.
+From each frame, spectrograms were extracted using the Fast Fourier Transform (FFT) algorithm, with 150 frequency bands spaced linearly from 100 Hz to 16 kHz.
 This procedure was conducted for both the left and right channels, yielding two spectrograms for each binaural sample.
-Each sample was represented by a matrix of dimensions $2 \times 349 \times 150$.
+Consequently, each sample was represented by a floating-point precision matrix of dimensions $2 \times 349 \times 150$.
 This method parallels the procedure presented by \cite{zielinski_automatic_2022}.
@@
 
-The network topology employed in this study, depicted in Figure \ref{fig:architecture}, was significantly influenced by the AlexNet convolutional neural network demonstrated by \cite{krizhevsky_imagenet_2012}.
-Although AlexNet was originally designed for image classification, this study adapted it for the audio prediction task by converting binaural recordings into magnitude spectrograms, as described in Section \ref{subsec:feature_extraction}.
-This conversion allowed the spectrograms to be treated as images, enabling their use in a standard image-prediction-like task.
+The network topology employed in this study was heavily influenced by the AlexNet convolutional neural network introduced by \cite{krizhevsky_imagenet_2012}.
+While AlexNet was originally designed for image classification, it was adapted for the audio prediction task in this study by converting binaural recordings into magnitude spectrograms, as described in Section \ref{subsec:feature_extraction}.
+This conversion allowed the spectrograms to be treated like images, allowing them to be used in a standard image-recognition-like task.
 
-The network accepts two spectrograms as inputs, and its architecture consists of a series of convolutional units followed by classification units, culminating in two output heads for predicting ensemble width and location.
-The architecture was designed in a multi-task fashion, enabling a single network to predict both ensemble parameters simultaneously.
-This multi-task capability is achieved by connecting the output of the final classification unit to two separate outputs, one for each ensemble parameter.
+As illustrated in Figure \ref{fig:architecture}, the network architecture consists of an input layer accepting a pair of spectrograms, followed by a series of convolutional units and classification units, culminating in two output heads for predicting ensemble width and location.
+This design employs a multi-task approach, enabling a single network to predict both ensemble parameters simultaneously.
 
-The topology finalized in this study was selected based on the highest prediction quality observed on the validation dataset.
-Despite the existence of numerous algorithms for automatic topology selection \parencite{branke_evolutionary_1995, zhang_finding_2018, miikkulainen_evolving_2017, stanley_evolving_2002, shafiee_deep_2016}, the final topology selected in this study was determined manually, based on its superior prediction quality observed on the validation dataset.
-This approach was necessitated primarily by the high computational demands compared to the available resources.
+The topology finalized in this study was constructed by selecting from many alternative architectures, choosing the one with the highest prediction quality observed on the validation dataset.
+Despite the existence of numerous algorithms for automatic topology selection \parencite{branke_evolutionary_1995, zhang_finding_2018, miikkulainen_evolving_2017, stanley_evolving_2002, shafiee_deep_2016}, the final topology was determined manually, primarily due to the high computational demands relative to the available resources.
 
-Various architectural configurations were assessed, varying key parameters such as the number of convolutional layers (from 1 to 5), the number of classification units (from 1 to 5), the inclusion or exclusion of Max Pooling Layers after each Convolution Layer, the number of filters within the Convolution Layers, the dimensions of these filters ($2 \times 2$, $2 \times 3$, or $3 \times 3$), the stride size, and the dimensions of the Max Pooling Layers ($2 \times 2$, $2 \times 3$, or $3 \times 3$).
-Although a detailed account of the selection process for the topology is beyond the primary scope of this study --- particularly considering the task's complexity --- it was concluded that the Convolutional Neural Network (CNN) model is robust against variations in the assessed topologies.
-The differences in mean prediction error among the configurations were minimal, typically less than $1^\circ$ for most configurations.
-Among the many tested topologies that yielded almost identical errors, the simplest one was chosen to optimize both performance efficiency and model simplicity.
+Various architectural configurations were assessed, with key parameters being varied such as the number of convolutional layers (from 1 to 5), the number of classification units (from 1 to 5), the inclusion or exclusion of Max Pooling Layers after each Convolution Layer, the number of filters within the Convolution Layers, the dimensions of these filters ($2 \times 2$, $2 \times 3$, or $3 \times 3$), the stride size, and the dimensions of the Max Pooling Layers ($2 \times 2$, $2 \times 3$, or $3 \times 3$).
+Based on this, it was concluded that the model is robust against variations in the assessed topologies.
+The differences in mean prediction error among the configurations were minimal, typically less than $1\degree$ for most configurations.
+Among the many tested topologies that yielded similar errors, the simplest one was selected to optimize both performance efficiency and model simplicity.
 
-Despite the availability of widely used techniques for addressing overfitting, such as the Dropout Layer \parencite{srivastava_dropout_2014}, and for accelerating training, such as Batch Normalization \parencite{ioffe_batch_2015}, neither technique was employed in this study due to their observed ineffectiveness for the specific prediction task at hand.
-Instead, a Global Average Pooling Layer \parencite{lin_network_2013} was utilized, known for its proven capabilities in reducing overfitting.
-In this particular task, the inclusion of this layer significantly reduced overfitting, lowering the final mean absolute error score by $0.83\degree$ (average across 10 trials) compared to configurations using a Flattening Layer.
+Despite the availability of widely used techniques for addressing overfitting, such as the Dropout Layer \parencite{srivastava_dropout_2014}, and for accelerating training, such as Batch Normalization \parencite{ioffe_batch_2015}, neither technique was employed in this study as they were observed to be ineffective for the specific prediction task being undertaken.
+Instead, a Global Average Pooling Layer was utilized, known for its capabilities in reducing overfitting \parencite{lin_network_2013}.
+This was confirmed in this particular task, as the inclusion of this layer significantly reduced overfitting, lowering the final mean absolute error score by $0.83\degree$ (average across 10 trials) compared to configurations using a Flattening Layer.
 
@@
 
-The topology described previously resulted in a model with 216,562 learning parameters.
-This model was trained 10 times from scratch, employing the Monte Carlo cross-validation method \parencite{kuhn_applied_2013}.
-For each iteration, the entire dataset was randomly divided into two parts: a development set containing two-thirds of the dataset (15,360 recordings) for model construction, and a test set comprising the remaining one-third (7,680 recordings) for evaluation.
-The development set was further subdivided into training and validation subsets at a 7:1 ratio, with 13,440 recordings allocated to the training subset and 1,920 recordings to the validation subset.
+The topology described in the previous section resulted in a model with 216,562 learning parameters.
+This model training procedure was repeated 10 times, employing the Monte Carlo cross-validation method \parencite{kuhn_applied_2013}.
+For each repetition, the entire dataset was randomly divided into two parts: a development set containing two-thirds of the dataset (15,360 recordings) for model construction, and a test set consisting of the remaining one-third (7,680 recordings) for evaluation.
 
-The training subset was employed to update the model's learning parameters, while the validation subset was utilized for implementing the early stopping technique \parencite{pocock_practical_1989, morgan_generalization_1989} in conjunction with model checkpointing \parencite{eisenman_check-n-run_2020}.
-These methods played a crucial role in selecting the model that demonstrated the best generalization capabilities.
-The test subset, containing data not previously encountered in the training or validation phases, was used exclusively for performance assessment once per iteration.
-This process was repeated to collect 10 mean absolute errors, from which the final model error was determined.
-The quantified mean absolute error and its standard deviation illustrated the model's prediction quality, generalization capabilities, and reproducibility.
+To ensure that the evaluation process was unbiased, this division was done in such a way that no original multi-track recordings used for synthesis were included in both the development and test sets simultaneously.
+However, this rule was not applied to HRTFs databases, allowing for the possibility of HRTF information leaking between the development and test sets.
+This could be seen as a significant limitation of the study.
+However, it is known that a human auditory system uses a single HRTF represented by ears and head, only slightly changing throughout the entire life, mainly during infancy \parencite{clifton_growth_1988, king_how_2001}, so this limitation could be considered in pair how the human auditory system behaves in the real life.
+Conversely, there are studies that do not implement such a limitation and test binaural localization models in an HRTF-independent manner \parencite{zielinski_spatial_2022, zielinski_automatic_2022,antoniuk_blind_2023}.
+
+The development set was divided into training and validation subsets at a 7:1 ratio, with 13,440 recordings in the training subset and 1,920 recordings in the validation subset.
+The training subset was used to update the model's learning parameters, while the validation subset was solely used for early stopping \parencite{pocock_practical_1989, morgan_generalization_1989} and model checkpointing \parencite{eisenman_check-n-run_2020}, which were used for selecting the model with the best generalization capabilities and preventing overfitting.
+The test subset, which included data not seen during the training or validation phases, was used solely for performance assessment once per a repetition.
+This divide-train-and-evaluate process was repeated to collect 10 mean absolute errors, from which the final model error was determined.
 
@@
 The rationale behind the application of Convolutional Neural Networks (CNNs) to this task was to automatically extract local features from the spectrograms and use these features to predict two contiguous ensemble parameters: ensemble location and width, both measured in degrees.
-The prediction errors, calculated as the difference between the actual ensemble parameters (known from the binaural synthesis described in Section \ref{subsec:synthesis}) and the predicted values, were utilized in the Adam learning algorithm \parencite{kingma_adam_2014}.
+The prediction errors, calculated as the difference between the actual ensemble parameters (known a priori from the binaural synthesis described in Section \ref{subsec:synthesis}) and the predicted values, were used in the Adam optimization algorithm \parencite{kingma_adam_2014}.
 During training, the losses for both outputs were combined by summing them, thereby treating both ensemble features with equal importance.
 
-All computations in this study were conducted on a workstation equipped with an RTX Nvidia GeForce 4090 GPU and a 48-core AMD Ryzen ThreadRipper processor.
-The binaural recordings were synthesized using the MATLAB \parencite{MATLAB} environment with the Audio Toolbox \parencite{MATLAB_Audio_Toolbox}.
-Feature extraction and model training were performed in the Python \parencite{python}, utilizing the SciPy \parencite{2020SciPy-NMeth} library for feature extraction and Keras for training the CNN model.
-The complete source code for all stages of the process has been made available by the authors and is accessible to the public at GitHub \parencite{github_repo}.
-The feature extraction phase required 21 minutes, data partitioning took 34 minutes, and the total training time for all iterations was 40 minutes.
-In total, the entire training and evaluation process required 95 minutes.
+The computational work for this study was conducted on a workstation equipped with an RTX Nvidia GeForce 4090 GPU and a 48-core AMD Ryzen ThreadRipper processor.
+On the software side, MATLAB \parencite{MATLAB} with the Audio Toolbox \parencite{MATLAB_Audio_Toolbox} was used for the binaural recording synthesis, while Python \parencite{python} with the SciPy package \parencite{2020SciPy-NMeth} was used for feature extraction and Keras \parencite{chollet2015keras} for training the CNN model.
+The complete source code for all stages is publicly available on the GitHub repository \parencite{github_repo}.
+The feature extraction phase required 21 minutes, data partitioning took 34 minutes, and the total training time for all iterations was 40 minutes, making the entire training and evaluation process 95 minutes long.
 
-\section{Results}
+\section{Results \& discussion}
 \label{sec:results}
+
+% jak uzyskane wyniki koresponduja z literatura: w jjednym przypadku ejst lepsza niz metoda oparta na spatiogramach, a w drugim przypadku nie ma
+
+The overall model performance measured across 10 experiment iterations, expressed as mean absolute error (MAE), was equal to $8.57\degree$ ($\pm0.19\degree$) for ensemble width and $4.76\degree$ ($\pm0.10\degree$) ensemble location.
+As both ensemble parameters were constrained within the same range of $90\degree$, the results demonstrate that the model exhibits a $44\%$ higher accuracy for ensemble location estimation than for ensemble width.
+This outcome is not unexpected, given that ensemble location is a less complex parameter.
+Essentially, it represents the average location of all sources.
+Therefore, it is more resistant to temporal fluctuations in individual audio sources than ensemble width, which is dependent on the two most extreme sound sources that vary over time.
+Furthermore, predicting ensemble width necessitates the identification of these two extreme sources, a process that is inherently more complex than estimating a single average location.
 
@@
 
-Figure \ref{fig:actual_vs_predicted_width} illustrates the comparison between the actual and predicted ensemble widths.
-The results demonstrate that the model exhibits better prediction quality for narrower ensemble widths ($5.65^\circ$ for $\omega < 30^\circ$) and that its performance deteriorates with an increase in ensemble width ($12.44^\circ$ for $\omega > 80^\circ$).
-This suggests that the actual ensemble width significantly impacts the accuracy of ensemble width estimation, leading to worse predictions as width increases.
-Figure \ref{fig:mae_width} further demonstrates that the relationship between the error and the width is not linear, exhibiting an interesting depression between $60^\circ$ and $75^\circ$.
+Figures \ref{fig:actual_vs_predicted_width} and \ref{fig:mae_width} compare the actual and predicted ensemble widths.
+The results indicate that the model provides more accurate predictions for narrower ensemble widths, with an average MAE of $5.65\degree$ for $\omega < 30\degree$.
+However, performance deteriorates as the ensemble width increases, resulting in a MAE of $12.44\degree$ for $\omega > 80\degree$.
+This deterioration highlights the significant impact of ensemble width on the accuracy of width estimation, particularly for wider ranges.
+In contrast, when comparing predicted locations to actual ones, the error remains relatively stable, as shown in Figures \ref{fig:actual_vs_predicted_location} and \ref{fig:mae_location}.
+The  deterioration in the width prediction can be attributed to the sparse distribution of audio sources in wider ensembles, which amplifies the influence of extreme sound sources on prediction errors, resulting in greater inaccuracies as the ensemble width increases.
+Moreover, Figure \ref{fig:mae_width} reveals that the relationship between the ensemble width and the error is nonlinear, displaying a notable decrease in error between $60\degree$ and $75\degree$.
+The reason for this nonlinearity is currently unclear and requires further investigation.
+
+% This observed depression is in pair with other related studies, like \cite{antoniuk_blind_2023,zielinski_comparison_2020}
 
@@
 In contrast to the correlation between ensemble width and its prediction error, there is no significant relationship between the actual location and its prediction error, as illustrated in Figures \ref{fig:actual_vs_predicted_location} and \ref{fig:mae_location}.
-This finding indicates that the model's capabilities for localizing the center of the ensemble is robust, unaffected by the actual spatial positioning of the ensemble, including lateral locations.
-
-% jak uzyskane wyniki koresponduja z literatura: w jjednym przypadku ejst lepsza niz metoda oparta na spatiogramach, a w drugim przypadku nie ma 
-The mean absolute prediction error for ensemble width was $8.57^\circ$ ($\pm0.19^\circ$), while the error for ensemble location prediction was $4.76^\circ$ ($\pm0.10^\circ$).
-Although both ensemble parameters were constrained within the same angular span of $90^\circ$, the ensemble location was predicted with significantly better precision --- approximately $80\%$ better.
-This shows that the model prediction quality is much better for ensemble location prediction than width prediction.
-This difference can be attributed to the fact that the model's performance on ensemble width deteriorates significantly for wider ensembles, whereas location prediction remains largely unaffected.
-For a visual comparison, please refer to Figures \ref{fig:mae_location} and \ref{fig:mae_width}.
+This finding indicates that the model's capabilities for localizing the center of the ensemble is more robust, unaffected by the actual spatial positioning of the ensemble, including lateral locations.
 
@@
 
-Figure \ref{fig:map_mae_location} illustrates the influence of both ensemble location and width on the mean absolute error for ensemble location, offering a detailed view complementing the results presented in Figure \ref{fig:mae_location}.
-Some asymmetric anomalies are depicted in this figure, mostly within the $\theta \in [15\degree, 30\degree]$ region, which can be attributed to the sparsity of sample result data across this heatmap.
-Although this figure primarily shows that ensemble location does not significantly influence the model's location prediction accuracy, it does demonstrate that ensemble width has a significant impact.
+Figure \ref{fig:map_mae_location} illustrates the influence of both ensemble location and width on the mean absolute error for ensemble location, providing a detailed perspective complementing the results presented in Figure \ref{fig:mae_location}.
+Notably, the figure highlights some asymmetric anomalies, especially within the $\theta \in [15\degree, 30\degree]$ range compared to the $\theta \in [-30\degree, -15\degree]$ range.
+These anomalies can be attributed to the sparsity of sample result data across this heatmap.
+While the figure indicates that ensemble location does not significantly affect the model's location prediction accuracy, it clearly shows that ensemble width has a substantial impact.
 Similarly, Figure \ref{fig:map_mae_width} reveals a characteristic depression in $\omega \in [30\degree, 60\degree]$ previously shown from a different perspective in Figure \ref{fig:mae_width}.
@@
 This indicates that the model's performance for estimating ensemble width is substantially worse at extreme widths and locations, i.e., when both the width and locations are at their maximum.
+
+The achieved model's performance was better to the model based on Spatiograms \parencite{arthi_spatiogram_2021}, which performance was further investigated by \cite{antoniuk_blind_2023} within the same evaluation scenario, using the same dataset, thus making it possible to compare these two results directly.
+The Spatiogram-based model achieved result of $13.62\degree$, making the model presented in this study better by $5.05\degree$, suggesting a significant improvement in the prediction of this ensemble parameter.
+Furthermore, the model presented in the previous study did not predict the ensemble location, which further enhances the effectiveness of this method.
 

### 2024-06-24 src/analyze/plot.py

@@ -328,7 +328,7 @@ cbar.set_ticks(cbar.get_ticks())
 cbar.set_ticklabels(tick_labels)
 
 plt.xlabel(r'Ensemble Location $\theta$')
-plt.ylabel('Ensemble Width')
+plt.ylabel(r'Ensemble Width $\omega$')
 xticks = np.linspace(0, mae_map_location.shape[1] - 1, 7)
 yticks = np.linspace(0, mae_map_location.shape[0] - 1, 4)
 xlabels = [f'{int(label)}°' for label in np.linspace(-45, 45, 7)]
@@ -357,8 +357,8 @@ tick_labels[-1] = f'>{tick_labels[-1]}'
 cbar.set_ticks(cbar.get_ticks())
 cbar.set_ticklabels(tick_labels)
 
-plt.xlabel('Ensemble Location')
-plt.ylabel('Ensemble Width')
+plt.xlabel(r'Ensemble Location $\theta$')
+plt.ylabel(r'Ensemble Width $\omega$')
 xticks = np.linspace(0, mae_map_width.shape[1] - 1, 7)
 yticks = np.linspace(0, mae_map_width.shape[0] - 1, 4)
 xlabels = [f'{int(label)}°' for label in np.linspace(-45, 45, 7)]

### 2024-07-02 paper/coverletter.tex
Fixes after SZ review

\documentclass[a4paper,12pt]{article}
\usepackage[a4paper,top=2cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage[UKenglish]{isodate}

\linespread{1.5}
\setlength{\parindent}{0pt}

\begin{document}

% # Speech processing and communication systems

\begin{flushright}
  Paweł Antoniuk \\
  Faculty of Computer Science \\
  Białystok University of Science \\
  Wiejska 45A, 15-351 Białystok, Poland \\
  \href{mailto:pawel.antoniuk@sd.pb.edu.pl}{pawel.antoniuk@sd.pb.edu.pl} \\
  \vspace{0.5cm}
  \today
\end{flushright}

\vspace{1cm}

\noindent
Editorial Board \\
Archives of Acoustics

\vspace{1cm}

\noindent
Dear Editor,

We submit our manuscript ``Estimating Ensemble Location and Width in Binaural Recordings of Music with Convolutional Neural Networks'' for publication in Archives of Acoustics. Authored by Paweł Antoniuk (corresponding author) and Sławomir Krzysztof Zieliński, this paper presents a method for estimating spatial parameters in binaural audio recordings using a novel ensemble-based approach employing a spectrogram-based deep learning network technique.

\begin{itemize}
  \item We submit this paper to Archives of Acoustics as its content aligns with the journal's scope, particularly in \textbf{speech processing and communication systems}.
  \item We confirm this manuscript is original, unpublished, and not under consideration elsewhere.
  \item We recommend the following reviewers:
        \begin{itemize}
          \item \textbf{prof. dr hab. inż. Bożena Kostek} \\*
          Wydział Elektroniki, Telekomunikacji i Informatyki \\*
          Gdańsk University of Technology \\*
          ul. Gabriela Narutowicza 11/12 \\*
          80-233 Gdańsk \\*
          Poland \\*
          Email: \href{mailto:bozenka@sound.eti.pg.gda.pl}{bozenka@sound.eti.pg.gda.pl}
          \item \textbf{prof. dr hab. inż. Piotr Kleczkowski} \\*
          Wydział Inżynierii Mechanicznej i Robotyki \\*
          AGH University of Krakow \\*
          al. Mickiewicza 30 \\*
          30-059 Kraków \\*
          Poland \\*
          Email: \href{mailto:kleczkow@agh.edu.pl}{kleczkow@agh.edu.pl}
          \item \textbf{dr hab. inż. Grzegorz Szwoch} \\*
          Wydział Elektroniki, Telekomunikacji i Informatyki \\*
          Gdańsk University of Technology \\*
          ul. Gabriela Narutowicza 11/12 \\*
          80-233 Gdańsk \\*
          Poland \\*
          Email: \href{mailto:grzegorz.szwoch@pg.edu.pl}{grzegorz.szwoch@pg.edu.pl}
        \end{itemize}
  \item We propose classifying this manuscript as a \textbf{``research paper''} due to its empirical content.
  \item While the main content is within 17 pages, the manuscript exceeds the recommended 20-page limit to provide comprehensive context, detailed methodology, and extensive references. We believe this additional content is crucial for a full understanding of our research and to allow others to replicate our work.
\end{itemize}

We appreciate your consideration of our manuscript.

\vspace{1cm}

Sincerely, \\
Paweł Antoniuk

\end{document}

### 2024-07-02 paper/main.tex
Fixes after SZ review

@@
-\documentclass{article}
+\documentclass[11pt]{article}
 
@@
 \usepackage{authblk}
-\usepackage[a4paper,top=2cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}
+\usepackage[a4paper,top=2cm,bottom=2cm,left=2cm,right=2cm,marginparwidth=1cm]{geometry}
 \usepackage{amsmath}
@@
 \usepackage{textcomp}
-\usepackage[backend=biber, style=apa, maxcitenames=1, uniquelist=false, url=false]{biblatex}
+\usepackage[backend=biber, style=custom, maxcitenames=1, uniquelist=false, uniquename=false, url=false]{biblatex}
 \usepackage{csquotes}
@@
 \usepackage{tabularray}
+\usepackage{indentfirst}
+\usepackage{setspace}
+\usepackage{orcidlink}
 
+\doublespacing
 \defbibenvironment{bibliography}
@@
 }
-
-% \AtEveryCitekey{\iffootnote{\csxdef{blx@footnotemark}{}\csxdef{blx@shortfootnotemark}{}}{}}
-
-% \DeclareCiteCommand{\cite}
-%   {\usebibmacro{prenote}}
-%   {\usebibmacro{citeindex}%
-%    \printtext[bibhyperref]{\usebibmacro{cite}}
-%   }
-%   {\multicitedelim}
-%   {\usebibmacro{postnote}}
-
-%   \DeclareCiteCommand{\parencite}
-%   {\usebibmacro{prenote}}
-%   {\usebibmacro{citeindex}%
-%    \printtext[bibhyperref]{\bibopenparen\usebibmacro{cite}\bibcloseparen}
-%   }
-%   {\multicitedelim}
-%   {\usebibmacro{postnote}}
-
-% % Redefine \parencites to handle multiple citations in one set of parentheses
-% \DeclareCiteCommand{\parencites}
-%   {\usebibmacro{prenote}}
-%   {\usebibmacro{citeindex}%
-%    \printtext[bibhyperref]{\bibopenparen\usebibmacro{cite}\bibcloseparen}
-%   }
-%   {\multicitedelim}
-%   {\usebibmacro{postnote}}
+\renewcommand\Affilfont{\itshape\centering}
+\renewenvironment{abstract}{}{}
+\newcommand{\keywords}[1]{
+  \noindent\begin{center}\begin{minipage}{\textwidth}
+  \small\textbf{Keywords:} #1
+  \end{minipage}\end{center}
+}
+\DefineBibliographyStrings{english}{%
+  andothers = {\textit{et al\adddot}},
+}
+\DeclareFieldFormat{doi}{%
+  \href{https://doi.org/#1}{#1}%
+}
+\renewbibmacro*{doi+eprint+url}{%
+  \iftoggle{bbx:doi}
+    {\printfield{doi}}
+    {}%
+  \newunit
+  \iftoggle{bbx:eprint}
+    {\usebibmacro{eprint}}
+    {}%
+  \newunit
+  \iftoggle{bbx:url}
+    {\usebibmacro{url+urldate}}
+    {}}
 
@@
 
-\title{Predicting Ensemble Location and Width in Binaural Recordings of Music with Convolutional Neural Networks}
-\author[1,*]{Paweł Antoniuk}
-\author[1]{Sławomir Zieliński}
-\affil[1]{Faculty of Computer Science, Białystok University of Technology}
-\affil[*]{Corresponding author: pawel.antoniuk@sd.pb.edu.pl}
+\title{Estimating Ensemble Location and Width in Binaural Recordings of Music with Convolutional Neural Networks}
+\author[1,*]{Paweł Antoniuk\orcidlink{0000-0002-0914-8920}}
+\author[1]{Sławomir Zieliński\orcidlink{0000-0002-3205-974X}}
+\affil[1]{Faculty of Computer Science, Białystok University of Technology \protect\\ Wiejska 45A, 15-351 Białystok, Poland}
+\affil[*]{Corresponding Author Email: pawel.antoniuk@sd.pb.edu.pl}
 \date{}
@@
 \begin{document}
+
 \maketitle
@@
 \begin{abstract}
-  Binaural audio technology has been in existence for many years, but its popularity has significantly increased over the past decade as a consequence of advancements in virtual reality and streaming technologies.
-Along with its growing popularity, the quantity of publicly accessible binaural audio materials has also expanded.
+  Binaural audio technology has been in existence for many years.
+However, its popularity has significantly increased over the past decade as a consequence of advancements in virtual reality and streaming techniques.
+Along with its growing popularity, the quantity of publicly accessible binaural audio recordings has also expanded.
 Consequently, there is now a need for automated and objective measurements of spatial content information, with ensemble location and width being the most prominent.
-This study presents a novel method for predicting ensemble location and width in binaural recordings of music.
-To this end, 30 head-related transfer functions and 192 binaural music recordings from publicly accessible multi-track recording repositories were used to synthesize 23,040 binaural recordings.
-The synthesized recordings were then used to train a multi-task convolutional neural network model, the aim of which was to predict the location and width of the ensemble for unseen recordings.
-The results indicate that models of ensemble location and width can be successfully constructed with low prediction errors --- $4.76\degree$ ($\pm0.10\degree$) for ensemble location and $8.57\degree$ ($\pm0.19\degree$) for ensemble width.
-The method developed in this study outperforms previous spatiogram-based methods recently published in the literature and holds promise for future development as a part of a novel tool for audio engineers to assess binaural audio.
+This study presents a novel method for estimating these ensemble parameters in binaural recordings of music.
+For this purpose, a dataset of 23,040 binaural recordings was synthesized from 192 publicly-available music recordings using 30 head-related transfer functions.
+The synthesized excerpts were then used to train a multi-task spectrogram-based convolutional neural network model, aiming to estimate the ensemble location and width for unseen recordings.
+The results indicate that a model for estimating ensemble parameters can be successfully constructed with low prediction errors: $4.76\degree$ ($\pm0.10\degree$) for ensemble location and $8.57\degree$ ($\pm0.19\degree$) for ensemble width.
+The method developed in this study outperforms previous spatiogram-based techniques recently published in the literature and shows promise for future development as part of a novel tool for binaural audio recordings analysis.
 \end{abstract}
+
+\keywords{ensemble width, ensemble location, binaural, spatial audio, localization, convolutional neural network, head-related transfer function, angle of arrival}
 
@@
 The human auditory system demonstrates exceptional proficiency in segregating, localizing, and interpreting diverse auditory signals, despite being limited to two ears.
-This is possible by internal examination of interaural differences in time, loudness, and frequency, known as binaural hearing \parencite{blauert_spatial_1996}, which enables precise localization of sound sources in complex auditory environments.
-A notable advantage of binaural hearing is exemplified by the `cocktail party effect', highlighting the system's capacity to concentrate on foreground sound sources while suppressing background noise \parencite{cherry_experiments_1953}.
-
-Understanding of the auditory system is essential for comprehending its limits and creating more immersive binaural experiences for entertainment purposes \parencite{zhang_surround_2017}.
+This is possible, among other factors, by internal examination of interaural differences in time, loudness, and frequency, known as binaural hearing \parencite{blauert_spatial_1996}, which enables precise localization of sound sources in complex auditory environments.
+A notable advantage of binaural hearing is exemplified by the ``cocktail party effect'', highlighting humans' capability to concentrate on foreground sound sources while suppressing background noise \parencite{cherry_experiments_1953}.
+Understanding of the auditory system is essential for comprehending its limits but also for leveraging these insights to create more immersive binaural experiences for entertainment purposes \parencite{zhang_surround_2017}.
 It is also important for enhancing auditory signal reception in hearing aid devices \parencite{hirsh_binaural_1950, thiemann_speech_2016}.
 
-The advancement of sophisticated machine learning techniques, particularly deep learning networks, has prompted an intriguing exploration into the extent to which these tools can emulate the human auditory system.
-This exploration proceeds without reliance on the advanced spatial audio feature engineering traditionally used in audio source localization, as evidenced by some prominent recent examples \parencite{yang_deepear_2022, vera-diaz_towards_2018, pang_multitask_2019}.
-While the application of convolutional neural networks (CNNs) \parencite{lecun_handwritten_1989} --- traditionally employed in the visual domain --- to audio signals, particularly in conjunction with spectrograms \parencite{thomas_analyzing_2014, espi_exploiting_2015, han_convolutional_2017} or other feature engineering techniques \parencite{abdel-hamid_applying_2012, sainath_deep_2013}, is not novel.
-However, these approaches are continually being refined and adapted to the audio realm.
-Building on these foundations, this study developed an audio localization method by constructing a spectrogram-based multi-task CNN model.
+The advance of sophisticated machine learning techniques, especially deep learning networks, has initiated an interesting exploration of their potential to emulate the human auditory system.
+Recently emerged studies have demonstrated that relying on the advanced spatial audio feature engineering is not necessary in computational audio source localization \parencite{vera-diaz_towards_2018, pang_multitask_2019, yang_deepear_2022}.
+While applying convolutional neural networks (CNNs) \parencite{lecun_handwritten_1989} to audio signals is well-established, often in conjunction with spectrograms \parencite{thomas_analyzing_2014, espi_exploiting_2015, han_convolutional_2017} or other feature engineering techniques \parencite{abdel-hamid_applying_2012, sainath_deep_2013}, these approaches continue to be refined and adapted for audio processing.
+Building on these foundations, this study develops an audio localization method using a spectrogram-based multi-task CNN model.
 
-Inspired by the fact, that humans tend to localize sound sources in groups rather than individually \parencite{bregman_auditory_1990, rumsey_spatial_2002}, the objective of the proposed model is to predict ensemble location and width instead of positions of individual sources.
-This study is unique as it not only conceptualized the method but also tested it on a vast, real-life music corpus of 23,040 binaural excerpts that were synthesized using 192 multi-track music recordings \parencite{senior_mixing_2023} and 30 publicly available head-related transfer functions (HRTFs) from various sources (see Table \ref{table:hrtfs} in \ref{appendix:a_hrtf}).
-The music recordings covered many different types of genres, including rock, jazz, pop, and classical music.
+Humans tend to localize groups of sound sources rather than individual ones \parencite{bregman_auditory_1990, rumsey_spatial_2002}.
+Inspired by this fact, the objective of the proposed model is to estimate the location and width of these groups, termed ensembles, instead of the positions of individual sources.
+This study is unique as it not only developed the method but also tested it on a vast, realistic music corpus.
+The corpus comprised 23,040 binaural excerpts synthesized using 192 multi-track music recordings (from a repository provided by \textcite{senior_mixing_2023}) and 30 publicly available head-related transfer functions (HRTFs) from various sources (see Table \ref{table:hrtfs} in \nameref{appendix:a_hrtf} for a detailed list).
+The music recordings covered many different genres, including rock, jazz, pop, and classical music.
 
-The findings demonstrate that this method is effective in accurately predicting the spatial characteristics of sound sources in near-real-world scenarios.
-Furthermore, this paper demonstrates an experimental framework that facilitates the objective measurement of a binaural localization technique, employing a large-scale dataset synthesized from real-world music signal (for applications of similar frameworks, see: \cite{antoniuk_blind_2023, zielinski_automatic_2022, zielinski_spatial_2022, zielinski_comparison_2020}).
+The findings demonstrate that this method is effective in accurately estimating the spatial characteristics of groups of sound sources in near-real-world scenarios.
+This paper also demonstrates an experimental framework that facilitates the objective measurement of a binaural localization technique, employing a large-scale dataset synthesized from real-world music signals (for applications of similar frameworks, see studies conducted by \textcite{antoniuk_blind_2023, zielinski_automatic_2022, zielinski_spatial_2022, zielinski_comparison_2020}).
 One of the key advantages of the proposed method is that it does not assume the number of audio sources.
@@
 The developed method has the potential to be highly beneficial in automated assessment tasks, where a significant number of binaural recordings must be measured or labeled in terms of their spatial content information.
-This could be utilized in the development of a hypothetical autonomous web-crawler bot that will collect binaural recordings from publicly accessible repositories and label them according to the spatial properties of the sound sources, such as the location of the music ensemble or the sparsity of audio source positions.
+This could be utilized in the development of a hypothetical autonomous ``web-crawler bot'' that will collect binaural recordings from publicly accessible repositories and label them according to the spatial properties of the sound sources, such as the location of the music ensemble or the sparsity of audio source positions.
 This method may also assist audio engineers in objectively assessing and segregating binaural audio recordings with regard to their spatial content.
 
-The structure of this paper is organized as follows.
-Section \ref{sec:related-studies} presents related studies.
+This paper is structured as follows: Section \ref{sec:related-studies} presents related studies.
 The description of the method developed for this study is provided in Section \ref{sec:methodology}, which also includes detailed definitions of ensemble location and width, along with a description of the experiments used to evaluate this method.
-Section \ref{sec:results} presents and discusses the performance of the presented method and the results of the experiments conducted in this study.
-The paper concludes in Section \ref{sec:conclusions}.
+Section \ref{sec:results} presents and discusses the performance of the proposed method as well as the results of the experiments conducted in this study.
+Finally, Section \ref{sec:conclusions} offers concluding remarks and suggestions for future research.
 
@@
 
-The majority of existing literature on the subject of sound source localization employs techniques that leverage the advantages of microphone arrays with more than two channels \parencite{kaveh_statistical_1986, pavlidi_real-time_2012, pan_multi-tone_2021, hahmann_sound_2022, chung_sound_2022, liu_sound_2022}.
-These method simplify the task and can improve the localization performance; however, they do not utilize binaural hearing, rendering them ineffective for binaural recordings.
+Most existing literature on computational sound source localization reports techniques that take advantage of multiple microphone arrays with more than two channels \parencite{kaveh_statistical_1986, pavlidi_real-time_2012, pan_multi-tone_2021, hahmann_sound_2022, chung_sound_2022, liu_sound_2022}.
+Although these methods can improve localization precision by providing additional spatial information, they do not utilize binaural hearing, rendering them ineffective for binaural recordings.
+In the context of sound source localization in binaural signals, the  focus of research is put on the identification of individual sound sources, rather than groups of sounds \parencite{dietz_auditory_2011, may_probabilistic_2011, may_binaural_2012, woodruff_binaural_2012, may_robust_2015, ma16c_interspeech, ma_exploiting_2017, benaroya_binaural_2018}.
 
-In the context of sound source localization in binaural signals, the  focus of research is on the identification of individual sound sources, rather than groups of sounds \parencite{dietz_auditory_2011, may_probabilistic_2011, may_binaural_2012, woodruff_binaural_2012, may_robust_2015, ma16c_interspeech, ma_exploiting_2017, benaroya_binaural_2018}.
-In terms of source direction of arrival (DOA) methods, the majority of research assumes a fixed number of sound sources \parencite{pang_multitask_2019, vera-diaz_towards_2018, ma_exploiting_2017, woodruff_binaural_2012, arthi_spatiogram_2021}, which limits its practical applications as this information is rarely known in real-life binaural recordings.
+Considering source Direction of Arrival (DoA) methods, the majority of research assumes a fixed number of sound sources \parencite{pang_multitask_2019, vera-diaz_towards_2018, ma_exploiting_2017, woodruff_binaural_2012, arthi_spatiogram_2021}, which limits its practical applications as this information is rarely known in real-life binaural recordings.
 Moreover, the majority of studies have focused on relatively homogeneous signals, namely speech \parencite{dietz_auditory_2011, may_probabilistic_2011, may_binaural_2012, woodruff_binaural_2012, may_robust_2015, ma16c_interspeech, ma_exploiting_2017, benaroya_binaural_2018, wang_binaural_2020, liu_multiple_2018, yang_deepear_2022, ma_robust_2018}.
@@
 In contrast to the aforementioned studies, the proposed method is not constrained by the number of sources.
-Furthermore, the approach is not limited to speech and has been applied to a wide range of musical datasets, including instruments and vocals.
-In contrast to previous studies that primarily focused on individual sources, the proposed method does not aim to separate them, but rather considers them as a group, or in this case, a musical ensemble.
-This approach is similar to how the real musical ensembles are arranged on stage, thereby emphasizing practical applications in live settings.
-To the authors' knowledge, this is one of the first methods to localize ensemble width (see \cite{antoniuk_blind_2023} for the previous ensemble-width-related study), and the first to localize both ensemble position and width simultaneously using a multi-task model.
+Moreover, the approach is not narrowed to speech and has been applied to a wide range of musical datasets, including instruments and vocals.
+In contrast to studies that primarily focused on individual sources, the proposed method does not aim to separate them, but rather considers them as a group, or in this case --- a musical ensemble --- similar to how real musical ensembles are arranged on stage.
+To the authors' knowledge, this is one of the first methods to localize ensemble width (see \textcite{antoniuk_blind_2023} for the previous ensemble-width-related study), and the first to localize both ensemble position and width simultaneously using a multi-task model.
 
-Sound localization methods can be classified into two categories based on the implementation of their underlying algorithms: glass-box (e.g., \parencite{dietz_auditory_2011, may_probabilistic_2011, may_binaural_2012, may_robust_2015, woodruff_binaural_2012, ma16c_interspeech, ma_exploiting_2017, ma_robust_2018}) and black-box (e.g., \parencite{vera-diaz_towards_2018, vera-diaz_towards_2018, yang_deepear_2022}).
-Glass-box methods, more traditional in the literature, rely on manually designed algorithms that mimic the auditory system to explicitly extract key features for location prediction, such as interaural level differences, interaural time differences, interaural coherence, or interaural phase differences (see \cite{blauert_spatial_1996} for feature descriptions).
-An example of one of the most advanced auditory models that is able to extract such features was developed by the Two!Ears project \parencite{raake_computational_nodate}.
+Sound localization methods can be classified into two categories based on the implementation of their underlying algorithms, termed as glass-box and black-box techniques.
+Glass-box methods could be considered as more traditional in the literature.
+They rely on manually designed algorithms that mimic the auditory system to explicitly extract key features for the localization estimation, such as interaural level differences, interaural time differences, interaural coherence, or interaural phase differences (\textcite{blauert_spatial_1996} provides detailed descriptions of these features).
+Examples of glass-box methods can be found in numerous studies, including those conducted by \textcite{dietz_auditory_2011, may_probabilistic_2011, may_binaural_2012, may_robust_2015, woodruff_binaural_2012, ma16c_interspeech, ma_exploiting_2017, ma_robust_2018, zielinski_automatic_2022}.
+These features are typically extracted using an auditory model.
+An advanced implementation capable of extracting these features was developed as part of the Two!Ears project \parencite{raake_computational_2016}.
 
-In contrast, black-box methods generally employ minimal feature engineering, relying instead on advanced machine learning techniques to autonomously extract features and make predictions.
-These methods, while effective, do not explicitly reveal the features they extract nor necessarily mimic the human auditory system.
-This opacity and the unpredictable nature of the outcomes, coupled with their dependency on deep neural networks that have extensive learning parameters, require the use of very large datasets for development and evaluation.
-Such datasets often include thousands of examples, as seen in the TIMIT corpus with 6300 examples \parencite{garofolo_darpa_1993} used in multiple studies \parencite{yang_deepear_2022, benaroya_binaural_2018, wang_binaural_2020, vera-diaz_towards_2018, ma_exploiting_2017, pang_multitask_2019, ma_robust_2018, may_robust_2015}, or custom corpora comprising hundreds of thousands of recordings \parencite{antoniuk_blind_2023, zielinski_automatic_2022, zielinski_spatial_2022, zielinski_comparison_2020}.
-This poses a significant challenge in assembling a sufficiently large and diverse collection of labeled binaural recordings.
-However, this challenge can be addressed through the synthesis of binaural sounds, as demonstrated in various studies \parencite{antoniuk_blind_2023, zielinski_automatic_2022, zielinski_spatial_2022, zielinski_comparison_2020,yang_deepear_2022, ma_robust_2018} and further elaborated upon in Section \ref{subsec:synthesis} of this paper.
+Black-box methods use a minimal degree of feature engineering, depending on deep neural networks to both extract features and make estimations.
+While effective, these methods do not necessarily consistently mimic human hearing, rendering them less suitable for objective measurement tasks (e.g., \textcite{vera-diaz_towards_2018, yang_deepear_2022}).
+Additionally, it is challenging to reveal their internally extracted features.
+Due to their opacity, unpredictable results, and numerous learning parameters, these methods should be treated more carefully.
+Moreover, they require large datasets for their development and evaluation.
+These datasets often contain thousands of examples, such as the TIMIT corpus \parencite{garofolo_darpa_1993} used in multiple studies \parencite{yang_deepear_2022, benaroya_binaural_2018, wang_binaural_2020, vera-diaz_towards_2018, ma_exploiting_2017, pang_multitask_2019, ma_robust_2018, may_robust_2015}.
+Some researchers have even created custom corpora with hundreds of thousands of recordings \parencite{antoniuk_blind_2023, zielinski_automatic_2022, zielinski_spatial_2022, zielinski_comparison_2020}.
+
+The necessity of having a large corpus to train deep learning models poses a significant challenge in gathering a sufficiently large and diverse collection of labeled binaural recordings.
+However, this challenge can be addressed through the synthesis of binaural sounds, as demonstrated in various studies \parencite{antoniuk_blind_2023, zielinski_automatic_2022, zielinski_spatial_2022, zielinski_comparison_2020, yang_deepear_2022, ma_robust_2018} and discussed further in Section \ref{subsec:synthesis}.
 
@@
 
-This section presents a detailed description of the main objective of the model developed as part of this study, as outlined in Section \ref{subsec:ensemble_definition}.
+This part of the paper presents a detailed description of the model developed in this study, as outlined in Section \ref{subsec:ensemble_definition}.
 It also describes the audio dataset used for training and evaluating the model, as detailed in Section \ref{subsec:synthesis}.
-In Section \ref{subsec:feature_extraction}, the feature extraction procedure is presented.
-Section \ref{subsec:topology} describes the model topology, whereas Section \ref{subsec:training_evaluation} address model training and evaluation.
+In Section \ref{subsec:feature_extraction}, the spectrograms calculation procedure is presented.
+Section \ref{subsec:topology} describes the model topology, whereas Section \ref{subsec:training_evaluation} addresses model training and evaluation.
 
@@
 
-The objective of the model developed in this study is to predict the ensemble location ($\theta$) and width ($\omega$), as illustrated in Figure \ref{fig:scene}.
+\begin{figure}[ht!]
+  \centering
+  \includegraphics[width=\linewidth]{../pictures/scene2.pdf}
+  \caption{\label{fig:scene}Illustration of ensemble width ($\omega$) and ensemble location ($\theta$) relative to the direction of the head orientation.
+Black dots represent the positions of audio sources $\theta_i$.
+The ensemble location ($\theta$) is the angular position of the center of the ensemble relative to the direction the head is facing.
+The ensemble width ($\omega$) is the angular distance between the two most extreme audio sources in the ensemble.}
+\end{figure}
+
+The objective of the model developed in this study is to estimate the ensemble location ($\theta$) and width ($\omega$), as illustrated in Figure \ref{fig:scene}.
 An ensemble is defined as a group of audio point sources positioned on a circle around the listener on a virtual acoustic scene with equal distance to the listener.
 The location of source $i$ is denoted by $\theta_i$.
-The ensemble width ($\omega$) is defined as the angular width between two extreme point sources ($max_i(\theta_i)-min_i(\theta_i)$), while the ensemble location, designated by $\theta$, represents the middle angle between two extreme sound sources ($(max_i(\theta_i)+min_i(\theta_i))/2$).
+The ensemble width ($\omega$) is defined as the angular distance between two extreme point sources ($max_i(\theta_i)-min_i(\theta_i)$), while the ensemble location, designated by $\theta$, represents the middle angle between two extreme sound sources ($(max_i(\theta_i)+min_i(\theta_i))/2$).
 For the purposes of this study, the locations of the sources were limited to the frontal hemisphere only, i.e.
 $\theta\in[-45\degree, 45\degree]$, $\omega\in[0\degree, 90\degree]$, as this range encompasses the majority of real-world recording scenarios.
-It should be noted that although humans possess some limited abilities to localize sound sources in the vertical plane, in this study all sources are placed in the horizontal plane, at the height of the listener.
-This covers the majority of cases for real-world recordings (see \cite{ma_robust_2018, zielinski_spatial_2022} for related studies that cover top-down discrimination).
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=\linewidth]{../pictures/scene.png}
-  \caption{\label{fig:scene}Illustration of ensemble width ($\omega$) and ensemble location ($\theta$) that are relative to the direction of the head.
-Black dots represent the positions of audio sources $\theta_i$.
-The ensemble location ($\theta$) is defined as the angular position of the center of the ensemble relative to the direction the head is turned.
-The ensemble width ($\omega$) is defined as angular distance between two extreme audio sources.
-}
-\end{figure}
+It should be noted that although humans possess some limited abilities to localize sound sources in the vertical plane, in this study all sources are placed in the horizontal plane, at the ear-level of the listener.
+This covers the majority of cases for real-world recordings (see \textcite{ma_robust_2018, zielinski_spatial_2022} for related studies that cover top-down discrimination).
 
@@
 The experiments conducted in this study involved 23,040 binaural recordings of music.
-The binaural recordings were synthesized using 192 multi-track publicly-available music recordings \parencite{senior_mixing_2023} and 30 HRTF databases (see Table \ref{table:hrtfs} in \nameref{appendix:a_hrtf} for a detailed list of HRTF databases used in this study).
+These recordings were synthesized using 192 publicly-available multi-track music recordings \parencite{senior_mixing_2023} and 30 HRTF databases (see Table \ref{table:hrtfs} in \nameref{appendix:a_hrtf} for a detailed list).
+The large number of HRTF databases was necessary to make the model as generalizable as possible.
+In real-world scenarios, the HRTF used for binaural synthesis is often unknown, so constructing a model for a single HRTF would have limited practical utility.
+The aim was to predict ensemble parameters regardless of the specific HRTF function used.
+Additionally, the large number of HRTF functions increased the amount of data available for model training, which is particularly beneficial in the context of deep neural networks.
+The number of HRTF databases (30) was determined using heuristics from previous study conducted by \textcite{zielinski_automatic_2022}, which suggested this number should be sufficient for the task.
+
 The number of tracks in multi-track recording ranged from 5 to 62, with median of 9.
@@
 Furthermore, the tracks of the input multi-track recordings were randomly assigned to sound source positions ($\theta_i$) to enhance the diversity of the final binaural corpora.
-Before the synthesis, the tracks were equalized to $-23$ LUFS, in accordance with the \parencite{noauthor_itu-r_2023} recommendation.
+Before the synthesis, the signals in each track were equalized to $-23$ LUFS, in accordance with the ITU-R BS.1770-5 (\citeyear{noauthor_itu-r_2023}) recommendation.
 
@@
 
-Due to copyright restrictions, the music corpus utilized in this study was not published and can be provided upon request to the authors of this paper.
+After the binauralization procedure, the synthesized recordings were truncated to a duration of seven seconds, with sine-squared fade-in and fade-out effects of 0.01 seconds applied.
+The recordings were then RMS-normalized, scaled by a factor of 0.9, and DC-offset corrected.
+They were stored as uncompressed files at a 48 kHz sampling rate and with a 32-bit resolution.
 
-\subsection{Feature extraction}
+Due to copyright restrictions, the music corpus utilized in this study was not published. However, the corpus can be provided upon reasonable request from the authors of this paper.
+
+\subsection{Calculation of spectrograms}
 \label{subsec:feature_extraction}
 
-% The procedure proposed in this study did not 
-% Zamiast tego dac obrazek z glowami i spektrogramami, lewy i prawi i roznica, sprawdzic czy to ma sens, potencjalnie usunac
-% \begin{figure}[ht]
-%   \centering
-%   \includegraphics[width=\linewidth]{../figures/samples.pdf}
-%   \caption{\label{fig:samples}Example magnitude spectrograms from randomly selected synthesized binaural music recordings, used as input data for training and evaluating the convolutional neural network (CNN) in this study.
-}
-% \end{figure}
-% Figure \ref{fig:samples} displays example spectrograms from four distinct binaural recordings, exhibiting a broad bandwidth and the diversity of the signals used in this study
-
-Prior to input into the model, the binaural recordings of music were transformed into magnitude spectrograms.
+Prior being input into the model, the binaural recordings of music were transformed into magnitude spectrograms.
 Although spectrograms do not directly provide information that can be translated into ensemble features, especially ensemble width, the goal of this task was to reduce the number of independent variables compared to the raw audio signal by extracting more compressed and informative data in the frequency domain.
 This step was necessary to decrease the likelihood of overfitting, reduce the number of examples needed to train the model, and thereby lower the overall computational power requirements.
-It is worth mentioning, however, that recent methods have shown that CNNs can be suitable for end-to-end audio localization without the spectrogram extraction step \parencite{vera-diaz_towards_2018, vecchiotti_end--end_2019}.
+It is worth mentioning, however, that recent methods have shown that CNNs are suitable for end-to-end audio localization without the spectrogram extraction step, as shown by \textcite{vera-diaz_towards_2018, vecchiotti_end--end_2019}.
 
-To prepare the input for the model, a Hamming window of 40 ms was applied to each frame of the signal, resulting in a total of 349 time frames.
+To prepare the input for the model, a Hamming window of 40 ms with an overlap of 20 ms was applied to each frame of the signal, resulting in a total of 349 time frames.
 From each frame, spectrograms were extracted using the Fast Fourier Transform (FFT) algorithm, with 150 frequency bands spaced linearly from 100 Hz to 16 kHz.
 This procedure was conducted for both the left and right channels, yielding two spectrograms for each binaural sample.
-Consequently, each sample was represented by a floating-point precision matrix of dimensions $2 \times 349 \times 150$.
-This method parallels the procedure presented by \cite{zielinski_automatic_2022}.
+Consequently, each sample was represented by a 32-bit floating-point precision matrix of dimensions $2 \times 349 \times 150$.
+This method parallels the procedure presented by \textcite{zielinski_automatic_2022}.
 
@@
 
-\begin{figure}[ht]
+The network topology employed in this study was strongly influenced by the AlexNet convolutional neural network introduced by \textcite{krizhevsky_imagenet_2012}.
+While AlexNet was originally designed for image classification, in this study it was adapted for the audio analysis task by converting binaural recordings into magnitude spectrograms, as described in Section \ref{subsec:feature_extraction}.
+This conversion allowed the spectrograms to be treated like visual data, enabling them to be used in an image-recognition-like task.
+
+As illustrated in Figure \ref{fig:architecture}, the network architecture consists of an input layer accepting a pair of spectrograms, followed by a series of convolutional units and classification units, culminating in two outputs predicting ensemble location and width, respectively.
+This design employs a multi-task approach, enabling a single network to estimate both ensemble parameters simultaneously.
+
+\begin{figure}[ht!]
   \centering
   \includegraphics[width=\linewidth]{../pictures/architecture.pdf}
-  \caption{\label{fig:architecture}Topology of the Convolutional Neural Network (CNN) used for identifying ensemble location and width}
+  \caption{\label{fig:architecture}Topology of the Convolutional Neural Network (CNN) used for estimating ensemble location and width, illustrating the layers (grouped in ``convolutional'' and ``classification'' units) and connections of the network architecture.}
 \end{figure}
 
-The network topology employed in this study was heavily influenced by the AlexNet convolutional neural network introduced by \cite{krizhevsky_imagenet_2012}.
-While AlexNet was originally designed for image classification, it was adapted for the audio prediction task in this study by converting binaural recordings into magnitude spectrograms, as described in Section \ref{subsec:feature_extraction}.
-This conversion allowed the spectrograms to be treated like images, allowing them to be used in a standard image-recognition-like task.
-
-As illustrated in Figure \ref{fig:architecture}, the network architecture consists of an input layer accepting a pair of spectrograms, followed by a series of convolutional units and classification units, culminating in two output heads for predicting ensemble width and location.
-This design employs a multi-task approach, enabling a single network to predict both ensemble parameters simultaneously.
-
-The topology finalized in this study was constructed by selecting from many alternative architectures, choosing the one with the highest prediction quality observed on the validation dataset.
+The topology finalized in this study was chosen, among many alternative architectures, based on the highest prediction quality observed on the validation dataset.
 Despite the existence of numerous algorithms for automatic topology selection \parencite{branke_evolutionary_1995, zhang_finding_2018, miikkulainen_evolving_2017, stanley_evolving_2002, shafiee_deep_2016}, the final topology was determined manually, primarily due to the high computational demands relative to the available resources.
 
-Various architectural configurations were assessed, with key parameters being varied such as the number of convolutional layers (from 1 to 5), the number of classification units (from 1 to 5), the inclusion or exclusion of Max Pooling Layers after each Convolution Layer, the number of filters within the Convolution Layers, the dimensions of these filters ($2 \times 2$, $2 \times 3$, or $3 \times 3$), the stride size, and the dimensions of the Max Pooling Layers ($2 \times 2$, $2 \times 3$, or $3 \times 3$).
-Based on this, it was concluded that the model is robust against variations in the assessed topologies.
+Various architectural configurations were assessed, with key parameters being varied such as the number of convolutional units (from 1 to 5), the number of classification units (from 1 to 5), the inclusion or exclusion of max pooling layers after each convolution layer, the number of filters within the convolutional layers, the dimensions of these filters ($2 \times 2$, $2 \times 3$, or $3 \times 3$), the stride size, and the dimensions of the max pooling layers ($2 \times 2$, $2 \times 3$, or $3 \times 3$).
+Based on this procedure, it was concluded that the model is robust against variations in the assessed topologies.
 The differences in mean prediction error among the configurations were minimal, typically less than $1\degree$ for most configurations.
@@
 
-Despite the availability of widely used techniques for addressing overfitting, such as the Dropout Layer \parencite{srivastava_dropout_2014}, and for accelerating training, such as Batch Normalization \parencite{ioffe_batch_2015}, neither technique was employed in this study as they were observed to be ineffective for the specific prediction task being undertaken.
-Instead, a Global Average Pooling Layer was utilized, known for its capabilities in reducing overfitting \parencite{lin_network_2013}.
-This was confirmed in this particular task, as the inclusion of this layer significantly reduced overfitting, lowering the final mean absolute error score by $0.83\degree$ (average across 10 trials) compared to configurations using a Flattening Layer.
+Despite the availability of widely used techniques for addressing  an overfitting effect, such as the dropout layer \parencite{srivastava_dropout_2014}, and for accelerating training, such as batch normalization \parencite{ioffe_batch_2015}, neither technique was employed in this study as they were observed to be ineffective for the specific estimation task being undertaken.
+Instead, a global average pooling layer was utilized, known for its capabilities in reducing overfitting \parencite{lin_network_2013}.
+This was confirmed in this particular task, as the inclusion of this layer significantly reduced overfitting, lowering the final mean absolute error (MAE) score by $0.83\degree$ (average across 10 trials) compared to configurations where a simple flattening layer was used instead.
 
@@
 The topology described in the previous section resulted in a model with 216,562 learning parameters.
-This model training procedure was repeated 10 times, employing the Monte Carlo cross-validation method \parencite{kuhn_applied_2013}.
-For each repetition, the entire dataset was randomly divided into two parts: a development set containing two-thirds of the dataset (15,360 recordings) for model construction, and a test set consisting of the remaining one-third (7,680 recordings) for evaluation.
+The model training procedure was repeated 10 times, employing the Monte Carlo cross-validation method, as described by \textcite{kuhn_applied_2013}.
+For each repetition, the entire dataset was randomly divided into two parts: a development set containing two-thirds of the dataset (15,360 recordings) for model construction, and a test set consisting of the remaining one-third of the dataset (7,680 recordings) for its evaluation.
+This repetition procedure was employed to ensure more reliable and generalizable results by assessing the model's performance across different subsets of the data.
+Additionally, it helps to account for the inherent variability in neural network training, where slight changes in initial conditions or optimization paths can lead to different model outcomes.
+While a large and diverse dataset could mitigate this issue, the binaural excerpts used in this study were generated from only 196 multi-track music recordings.
+This limited source material raised concerns about potential significant variations between the development and test sets in each repetition.
+These concerns were validated, as the maximum observed difference in MAE between repetitions reached up to $0.85\degree$ for ensemble width.
 
@@
 This could be seen as a significant limitation of the study.
-However, it is known that a human auditory system uses a single HRTF represented by ears and head, only slightly changing throughout the entire life, mainly during infancy \parencite{clifton_growth_1988, king_how_2001}, so this limitation could be considered in pair how the human auditory system behaves in the real life.
-Conversely, there are studies that do not implement such a limitation and test binaural localization models in an HRTF-independent manner \parencite{zielinski_spatial_2022, zielinski_automatic_2022,antoniuk_blind_2023}.
+However, it is known that a human auditory system uses a single HRTF represented by ears, head, and torso, only slightly changing throughout the entire life, mainly during infancy \parencite{clifton_growth_1988, king_how_2001}.
+Therefore, this limitation could be considered in pair how the human auditory system behaves in real life.
+Nevertheless, it is worth noting that some studies implement HRTF-independent testing for binaural localization models, as demonstrated by \textcite{zielinski_spatial_2022, zielinski_automatic_2022,antoniuk_blind_2023}.
 
 The development set was divided into training and validation subsets at a 7:1 ratio, with 13,440 recordings in the training subset and 1,920 recordings in the validation subset.
-The training subset was used to update the model's learning parameters, while the validation subset was solely used for early stopping \parencite{pocock_practical_1989, morgan_generalization_1989} and model checkpointing \parencite{eisenman_check-n-run_2020}, which were used for selecting the model with the best generalization capabilities and preventing overfitting.
+The training subset was used to update the model's learning parameters, while the validation subset was solely used for early stopping \parencite{pocock_practical_1989, morgan_generalization_1989} and model checkpointing \parencite{eisenman_check-n-run_2020}.
+These techniques were employed to select the model with the best generalization capabilities and prevent overfitting.
 The test subset, which included data not seen during the training or validation phases, was used solely for performance assessment once per a repetition.
@@
 For each sample, the model received two spectrograms as input: one for the left channel and one for the right channel.
-The rationale behind the application of Convolutional Neural Networks (CNNs) to this task was to automatically extract local features from the spectrograms and use these features to predict two contiguous ensemble parameters: ensemble location and width, both measured in degrees.
-The prediction errors, calculated as the difference between the actual ensemble parameters (known a priori from the binaural synthesis described in Section \ref{subsec:synthesis}) and the predicted values, were used in the Adam optimization algorithm \parencite{kingma_adam_2014}.
-During training, the losses for both outputs were combined by summing them, thereby treating both ensemble features with equal importance.
+The rationale behind the application of Convolutional Neural Networks (CNNs) to this task was to automatically extract local features from the spectrograms and use these features to estimate two contiguous ensemble parameters: ensemble location and width, both measured in degrees.
+For model training, the Adam algorithm \parencite{kingma_adam_2014} was used.
+The algorithm minimized prediction errors, calculated as the difference between the actual ensemble parameters (known \textit{a priori} from the binaural synthesis described in Section \ref{subsec:synthesis}) and the predicted values.
+
+The optimizer was configured with the following hyperparameters: an initial learning rate of $10^{-3}$, a decay rate of $10^{-6}$, and momentum parameters $\beta_1 = 0.9$ and $\beta_2 = 0.999$.
+Training was conducted using a batch size of 8.
+An early stopping technique was implemented to prevent overfitting, terminating the process if no improvement was observed on the validation set for 20 consecutive epochs.
+Consequently, the maximum number of epochs was never reached; instead, training concluded after 25 to 36 epochs, with a median of 27.5 epochs.
+During the training process, the losses for both outputs were combined additively, ensuring equal weighting of both ensemble features.
 
 The computational work for this study was conducted on a workstation equipped with an RTX Nvidia GeForce 4090 GPU and a 48-core AMD Ryzen ThreadRipper processor.
-On the software side, MATLAB \parencite{MATLAB} with the Audio Toolbox \parencite{MATLAB_Audio_Toolbox} was used for the binaural recording synthesis, while Python \parencite{python} with the SciPy package \parencite{2020SciPy-NMeth} was used for feature extraction and Keras \parencite{chollet2015keras} for training the CNN model.
-The complete source code for all stages is publicly available on the GitHub repository \parencite{github_repo}.
-The feature extraction phase required 21 minutes, data partitioning took 34 minutes, and the total training time for all iterations was 40 minutes, making the entire training and evaluation process 95 minutes long.
+On the software side, MATLAB \parencite{MATLAB} with the Audio Toolbox \parencite{MATLAB_Audio_Toolbox} was used for the binaural recording synthesis, while Python \parencite{python} with the SciPy package \parencite{2020SciPy-NMeth} was used for feature extraction and Keras \parencite{chollet_keras_2015} for training the CNN model.
+The complete source code for all stages is publicly available on the GitHub repository \parencite{antoniuk_software_2024}.
+The spectrogram calculation phase required 21 minutes, data partitioning took 34 minutes, and the total training time for all iterations amounted to 40 minutes, making the entire training and evaluation process 95 minutes long.
 
-\section{Results \& discussion}
+\section{Results and discussion}
 \label{sec:results}
 
-% jak uzyskane wyniki koresponduja z literatura: w jjednym przypadku ejst lepsza niz metoda oparta na spatiogramach, a w drugim przypadku nie ma
-
-The overall model performance measured across 10 experiment iterations, expressed as mean absolute error (MAE), was equal to $8.57\degree$ ($\pm0.19\degree$) for ensemble width and $4.76\degree$ ($\pm0.10\degree$) ensemble location.
-As both ensemble parameters were constrained within the same range of $90\degree$, the results demonstrate that the model exhibits a $44\%$ higher accuracy for ensemble location estimation than for ensemble width.
+The overall model performance measured across 10 experiment iterations, expressed as mean absolute error (MAE), was equal to $8.57\degree$ ($\pm0.19\degree$) for ensemble width and $4.76\degree$ ($\pm0.10\degree$) for ensemble location.
+As both ensemble parameters were constrained within the same range of $90\degree$, the results demonstrate that the model exhibits a $44\%$ lower error for ensemble location compared to ensemble width.
 This outcome is not unexpected, given that ensemble location is a less complex parameter.
@@
 Therefore, it is more resistant to temporal fluctuations in individual audio sources than ensemble width, which is dependent on the two most extreme sound sources that vary over time.
-Furthermore, predicting ensemble width necessitates the identification of these two extreme sources, a process that is inherently more complex than estimating a single average location.
+Furthermore, estimating ensemble width necessitates the identification of these two extreme sources, a process that is inherently more complex than estimating a single average location.
 
-\begin{figure}[ht]
+Figure \ref{fig:actual_vs_predicted_width} compares the actual and predicted ensemble widths for each sample, showing a heteroscedastic relationship between them, with a slight bias towards predicting lower ensemble width values for higher actual widths.
+This relationship exhibits a strong positive correlation, with a Pearson r of 0.9.
+Additionally, the results indicate that the model provides more precise predictions for narrower ensemble widths, with an average Mean Absolute Error (MAE) of $5.65^\circ$ for $\omega < 30^\circ$.
+However, performance deteriorates as the ensemble width increases, resulting in an MAE of $12.44^\circ$ for $\omega > 80^\circ$.
+This effect is more visible in Figure \ref{fig:mae_width}, which highlights the impact of the actual ensemble width on the precision of prediction.
+The correlation between the actual ensemble width and prediction error shows a weak positive relationship, with a Pearson r of 0.27
+
+\begin{figure}[ht!]
   \centering
-  \begin{minipage}[t]{0.45\linewidth}
+  \begin{minipage}[]{0.45\linewidth}
     \includegraphics[width=\linewidth]{../figures/actual_vs_predicted_width.pdf}
-    \caption{\label{fig:actual_vs_predicted_width}A comparison between the actual ensemble width $\omega$ and the predicted ensemble width $\omega'$ for a single iteration (of the total five) }
+    \caption{\label{fig:actual_vs_predicted_width}A comparison between the actual ensemble width $\omega$ and the predicted ensemble width $\omega'$ for a single iteration (of the total ten). }
   \end{minipage}
   \hspace{0.5cm}
-  \begin{minipage}[t]{0.45\linewidth}
+  \begin{minipage}[]{0.45\linewidth}
     \includegraphics[width=\linewidth]{../figures/mae_width.pdf}
-    \caption{\label{fig:mae_width}The impact of the actual ensemble width $\omega$ on the mean absolute prediction error, averaged across all five iterations, with indicated standard deviation.}
+    \caption{\label{fig:mae_width}The impact of the actual ensemble width $\omega$ on the mean absolute prediction error, averaged across all ten iterations, with indicated standard deviation.}
   \end{minipage}
@@
 
-Figures \ref{fig:actual_vs_predicted_width} and \ref{fig:mae_width} compare the actual and predicted ensemble widths.
-The results indicate that the model provides more accurate predictions for narrower ensemble widths, with an average MAE of $5.65\degree$ for $\omega < 30\degree$.
-However, performance deteriorates as the ensemble width increases, resulting in a MAE of $12.44\degree$ for $\omega > 80\degree$.
-This deterioration highlights the significant impact of ensemble width on the accuracy of width estimation, particularly for wider ranges.
-In contrast, when comparing predicted locations to actual ones, the error remains relatively stable, as shown in Figures \ref{fig:actual_vs_predicted_location} and \ref{fig:mae_location}.
-The  deterioration in the width prediction can be attributed to the sparse distribution of audio sources in wider ensembles, which amplifies the influence of extreme sound sources on prediction errors, resulting in greater inaccuracies as the ensemble width increases.
+The reduced accuracy in the width prediction can be attributed to the sparse distribution of audio sources in wider ensembles, which amplifies the influence of extreme sound sources on prediction errors, resulting in lower precision as the ensemble width increases.
 Moreover, Figure \ref{fig:mae_width} reveals that the relationship between the ensemble width and the error is nonlinear, displaying a notable decrease in error between $60\degree$ and $75\degree$.
@@
 
-% This observed depression is in pair with other related studies, like \cite{antoniuk_blind_2023,zielinski_comparison_2020}
+The correlation between actual and predicted ensemble location exhibits a high Pearson r of 0.97, as illustrated in Figure \ref{fig:actual_vs_predicted_location}.
+In contrast to the ensemble width, no significant relationship is observed between actual location and its prediction error.
+This finding suggests that the model's ability to localize the center of the ensemble is robust, unaffected by the actual spatial positioning of the ensemble, including lateral locations.
+Figure \ref{fig:mae_location} corroborates this observation, demonstrating relatively consistent location error across most positions, with minor increases at extreme locations.
+The negligible correlation (Pearson r = -0.03) between absolute location value and prediction error further supports the model's spatial invariance in performance.
 
-\begin{figure}[ht]
+\begin{figure}[ht!]
   \centering
@@
     \includegraphics[width=\linewidth]{../figures/actual_vs_predicted_location.pdf}
-    \caption{\label{fig:actual_vs_predicted_location}A comparison between the actual ensemble location $\theta$ and the predicted ensemble location $\theta'$ for a single iteration (of the total five) }
+    \caption{\label{fig:actual_vs_predicted_location}A comparison between the actual ensemble location $\theta$ and the predicted ensemble location $\theta'$ for a single iteration (of the total ten) }
   \end{minipage}
@@
     \includegraphics[width=\linewidth]{../figures/mae_location.pdf}
-    \caption{\label{fig:mae_location}The impact of the actual ensemble width $\omega$ on the mean absolute prediction error, averaged across all five iterations, with indicated standard deviation.}
+    \caption{\label{fig:mae_location}The impact of the actual ensemble width $\omega$ on the mean absolute prediction error, averaged across all ten iterations, with indicated standard deviation.}
   \end{minipage}
@@
 
-In contrast to the correlation between ensemble width and its prediction error, there is no significant relationship between the actual location and its prediction error, as illustrated in Figures \ref{fig:actual_vs_predicted_location} and \ref{fig:mae_location}.
-This finding indicates that the model's capabilities for localizing the center of the ensemble is more robust, unaffected by the actual spatial positioning of the ensemble, including lateral locations.
+Figure \ref{fig:map_mae_location} illustrates the influence of both ensemble location and width on the mean absolute error for ensemble location, providing a detailed perspective complementing the results presented in Figure \ref{fig:mae_location}.
+Notably, the figure highlights asymmetric anomalies, particularly within the $\theta \in [15\degree, 30\degree]$ range compared to the $\theta \in [-30\degree, -15\degree]$ range, which can be attributed to the sparsity of sample result data across specific regions of this heatmap.
+While the figure suggests that ensemble location does not significantly affect the model's precision in predicting location, it clearly demonstrates that ensemble width has a substantial impact.
+Specifically, there is a positive correlation between the width of the ensemble and the error in its location prediction, with error magnitude increasing as the width expands.
 
-\begin{figure}[ht]
+\begin{figure}[ht!]
   \centering
   \includegraphics[width=\linewidth]{../figures/map_mae_location.pdf}
-  \caption{\label{fig:map_mae_location}The heatmap that illustrates the mean absolute error (MAE) of ensemble location distribution across different ensemble locations (x-axis) and ensemble widths (y-axis).
+  \caption{\label{fig:map_mae_location}The heatmap illustrating the mean absolute error (MAE) of ensemble location distribution across different ensemble locations (x-axis) and ensemble widths (y-axis).
 The color intensity corresponds to the MAE values, with lighter areas indicating higher errors.}
@@
 
-\begin{figure}[ht]
+Figure \ref{fig:map_mae_width} reveals a characteristic depression in $\omega \in [30\degree, 60\degree]$ previously shown from a different perspective in Figure \ref{fig:mae_width}.
+This heatmap highlights another interesting phenomenon in its upper corners as the error in these areas is considerably higher.
+This indicates that the model's performance for estimating ensemble width is substantially worse at extreme widths and locations, i.e., when both the width and locations are near their maximum investigated values ($|\theta| \approx 45\degree, \omega \approx 90\degree$).
+
+\begin{figure}[ht!]
   \centering
   \includegraphics[width=\linewidth]{../figures/map_mae_width.pdf}
-  \caption{\label{fig:map_mae_width}The heatmap that illustrates the mean absolute error (MAE) of ensemble width distribution across different ensemble locations (x-axis) and ensemble widths (y-axis).
-The color intensity corresponds to the MAE values, with lighter areas indicating higher errors.
-Notably, the values between $30\degree$ and $60\degree$ on the y-axis exhibit unexpectedly higher MAE values region --- please see Figure \ref{fig:mae_width} for comparison.
-}
+  \caption{\label{fig:map_mae_width}The heatmap illustrating the mean absolute error (MAE) of ensemble width distribution across different ensemble locations (x-axis) and ensemble widths (y-axis).
+The color intensity corresponds to the MAE values, with lighter areas indicating higher errors.}
 \end{figure}
 
-Figure \ref{fig:map_mae_location} illustrates the influence of both ensemble location and width on the mean absolute error for ensemble location, providing a detailed perspective complementing the results presented in Figure \ref{fig:mae_location}.
-Notably, the figure highlights some asymmetric anomalies, especially within the $\theta \in [15\degree, 30\degree]$ range compared to the $\theta \in [-30\degree, -15\degree]$ range.
-These anomalies can be attributed to the sparsity of sample result data across this heatmap.
-While the figure indicates that ensemble location does not significantly affect the model's location prediction accuracy, it clearly shows that ensemble width has a substantial impact.
-Similarly, Figure \ref{fig:map_mae_width} reveals a characteristic depression in $\omega \in [30\degree, 60\degree]$ previously shown from a different perspective in Figure \ref{fig:mae_width}.
-This heatmap highlights another interesting phenomenon in its upper corners --— the error in these areas is considerably higher.
-This indicates that the model's performance for estimating ensemble width is substantially worse at extreme widths and locations, i.e., when both the width and locations are at their maximum.
+The model presented in this study demonstrates a significant improvement in ensemble-width performance compared to the Spatiogram-based model, first introduced by \textcite{arthi_spatiogram_2021} and further investigated by \textcite{antoniuk_blind_2023}, under similar evaluation conditions.
+While the dataset used in this study was expanded with 40 additional multi-track recordings and 10 HRTF databases, \textcite{antoniuk_blind_2023} showed that the Spatiogram model's performance does not improve with further increases in dataset size.
+This finding enables a direct comparison of results between the two models in terms of the precision of the ensemble width estimation, despite the differences in dataset composition.
+Our model achieved a MAE of $8.57\degree$, outperforming the Spatiogram-based model's result of $13.62\degree$ by $5.05\degree$.
+This substantial improvement is further enhanced by the current model's ability to estimate ensemble location, a feature absent in the previous model.
 
-The achieved model's performance was better to the model based on Spatiograms \parencite{arthi_spatiogram_2021}, which performance was further investigated by \cite{antoniuk_blind_2023} within the same evaluation scenario, using the same dataset, thus making it possible to compare these two results directly.
-The Spatiogram-based model achieved result of $13.62\degree$, making the model presented in this study better by $5.05\degree$, suggesting a significant improvement in the prediction of this ensemble parameter.
-Furthermore, the model presented in the previous study did not predict the ensemble location, which further enhances the effectiveness of this method.
+In terms of ensemble location prediction, the novelty of the proposed method makes direct comparison with existing literature challenging.
+However, its efficacy can be only evaluated indirectly against state-of-the-art individual-source localization techniques.
+The ensemble location prediction precision (MAE $=4.76\degree$) of the proposed method can be contextualized with the leading-edge binaural localization DeepEar model introduced by \textcite{yang_deepear_2022}.
+This model reported MAEs of $7.4\degree$ and $2.3\degree$ for multi-source and single-source Angle of Arrival (AoA) estimation, respectively.
+As another promising example, the WaveLoc-CONV model developed by \textcite{vecchiotti_end--end_2019} demonstrated errors of $0\degree$ in anechoic conditions and $1.7\degree-2.4\degree$ in multi-condition scenarios.
+However, these results are limited to single-source speech localization, a substantially less complex task than the ensemble location prediction addressed by the proposed method.
+These experiments, while differing in objectives and datasets, provide valuable context for the proposed method's performance within current DoA and AoA estimation research.
 
@@
 
-% 1. napisac wnioski, potem occulusion. we wnioskach odniesc sie do spatiogramach
+This paper introduces a novel approach to locating audio sources in binaural recordings.
+Unlike traditional methods that predict the locations of individual audio sources, this study focuses on estimating ``ensemble parameters'' of audio sources, thus allowing the audio scene to be described using two parameters only: ensemble location and width.
+This approach makes it possible to avoid making restrictive assumptions about the number of audio sources, rendering the proposed method more suitable for real-world applications.
+The study also explores the use of convolutional neural networks (CNN) in conjunction with spectrograms applied to their inputs.
+According to the obtained results, the networks show exceptionally good performance, demonstrating their suitability for the investigated scenario.
 
-\clearpage
-\section*{Appendix A}
+The method was developed using 23,040 synthesized binaural excerpts intended to mimic real-world music recordings.
+The results show its outstanding performance, with the model achieving mean absolute error (MAE) of $4.76\degree \pm 0.10\degree$ and $8.57\degree \pm 0.19\degree$ for the estimation of ensemble location and width, respectively.
+While the model is resilient to lateral ensemble locations, it is sensitive to the actual ensemble width, lowering the model accuracy as the width increases.
+The proposed method demonstrates a significant improvement over the previous technique based on spatiograms \parencite{antoniuk_blind_2023}, lowering the mean absolute error by $5.05\degree$.
+
+Despite its high precision, the method exhibits certain limitations.
+Since it has been developed using the binaural excerpts synthesized with the head-related impulse responses being inherently anechoic in their characteristics, the method's performance under reverberant conditions has not been validated.
+Moreover, the proposed method is incapable of operating in real-time scenarios.
+Validating the method under reverberant conditions as well as optimizing its architecture for practical real-time scenarios constitute the topics for future research.
+Other minor limitations include the lack of head-related transfer-function (HRTF) independence between the development and test sets, and the absence of vertical variations in audio source placement, as all sources were positioned on the horizontal plane.
+Additionally, the proposed approach requires substantial computational resources, particularly GPU usage, which was not necessary for the previously used spatiogram-based method.
+
+These limitations, however, present opportunities for future research.
+Despite the current constraints, this study introduces a novel method for describing acoustic scenes in binaural recordings of music, demonstrating substantial potential for advancing binaural audio analysis.
+The method offers promising prospects for developing innovative tools that can objectively measure and analyze large repositories of binaural audio, focusing on spatial content.
+
+% \clearpage
+\section*{Acknowledgments}
+The work was supported by the grants from Białystok University of Technology (WI/WI-IiT/3/2022 and WZ/WI-IIT/5/2023) and funded with resources for research by the Ministry of Science and Higher Education in Poland.
+
+\section*{Appendix}
 \label{appendix:a_hrtf}
@@
   }
-  \textbf{No.} & \textbf{Type} & \textbf{Head}                             & \textbf{Radius {[}m{]}} & \textbf{Source}                                                                                                                            & \textbf{Acronym} \\
+  \textbf{No.} & \textbf{Type} & \textbf{Head}                             & \textbf{Radius {[}m{]}} & \textbf{Source}                                                                                                                            & \textbf{Acronym} \\*
   1.
-& Human         & Human subject                             & 1.2                     & RWTH Aachen University \parencite{braren_high-resolution_2020}                                                                             & AACHEN           \\
+& Human         & Human subject                             & 1.2                     & RWTH Aachen University \parencite{braren_high-resolution_2020}                                                                             & AACHEN           \\*
   2.
@@
   3.
-& Human         & Subject 2                                 & 1.2                     & Austrian Academy of Sciences \parencite{noauthor_hrtf-database_nodate}                                                                     & ARI              \\
+& Human         & Subject 2                                 & 1.2                     & Austrian Academy of Sciences \parencite{noauthor_hrtf-database_2014}                                                                       & ARI              \\*
   4.
-& Human         & Subject 4                                 & 1.2                     &                                                                                                                                            &                  \\
+& Human         & Subject 4                                 & 1.2                     &                                                                                                                                            &                  \\*
   5.
-& Human         & Subject 10                                & 1.2                     &                                                                                                                                            &                  \\
+& Human         & Subject 10                                & 1.2                     &                                                                                                                                            &                  \\*
   6.
@@
   7.
-& Human         & Subject 012                               & 1                       & CIPIC Interface Laboratory, University of California \parencite{algazi_cipic_2001}                                                         & CIPIC            \\
+& Human         & Subject 012                               & 1                       & CIPIC Interface Laboratory, University of California \parencite{algazi_cipic_2001}                                                         & CIPIC            \\*
   8.
-& Human         & Subject 015                               & 1                       &                                                                                                                                            &                  \\
+& Human         & Subject 015                               & 1                       &                                                                                                                                            &                  \\*
   9.
@@
   10.
-& Artificial    & Neumann KU 100                            & 0.9                     & NASA (2007) \parencite{andreopoulou_inter-laboratory_2015}                                                                                 & CLUBFRITZ        \\
+& Artificial    & Neumann KU 100                            & 0.9                     & NASA (2007) \parencite{andreopoulou_inter-laboratory_2015}                                                                                 & CLUBFRITZ        \\*
   11.
@@
   12.
-& Artificial    & FABIAN                                    & 1.47                    & Technical University Berlin, Huawei Technologies, Munich Research Centre, Sennheiser Electronic \parencite{brinkmann_cross-evaluated_2019} & HUTUBS           \\
+& Artificial    & FABIAN                                    & 1.47                    & Technical University Berlin, Huawei Technologies, Munich Research Centre, Sennheiser Electronic \parencite{brinkmann_cross-evaluated_2019} & HUTUBS           \\*
   13.
-& Human         & Subject pp2                               & 1.47                    &                                                                                                                                            &                  \\
+& Human         & Subject pp2                               & 1.47                    &                                                                                                                                            &                  \\*
   14.
@@
   15.
-& Human         & Subject 1003                              & 1.95                    & IRCAM, AKG \parencite{noauthor_listen_2023}                                                                                                & LISTEN           \\
+& Human         & Subject 1003                              & 1.95                    & IRCAM, AKG \parencite{noauthor_listen_2023}                                                                                                & LISTEN           \\*
   16.
@@
   17.
-& Artificial    & KEMAR DB-4004 (DB-061)                    & 1.4                     & MIT \parencite{gardne_hrtf_1994}                                                                                                           & MIT              \\
+& Artificial    & KEMAR DB-4004 (DB-061)                    & 1.4                     & MIT \parencite{gardne_hrtf_1994}                                                                                                           & MIT              \\*
   18.
@@
   19.
-& Human         & Subject 001                               & 1.5                     & Tohoku University \parencite{watanabe_dataset_2014}                                                                                        & RIEC             \\
+& Human         & Subject 001                               & 1.5                     & Tohoku University \parencite{watanabe_dataset_2014}                                                                                        & RIEC             \\*
   20.
-& Human         & Subject 002                               & 1.5                     &                                                                                                                                            &                  \\
+& Human         & Subject 002                               & 1.5                     &                                                                                                                                            &                  \\*
   21.
@@
   22.
-& Artificial    & Neumann KU 100                            & 1.2                     & University of York \parencite{armstrong_perceptual_2018}                                                                                   & SADIE II         \\
+& Artificial    & Neumann KU 100                            & 1.2                     & University of York \parencite{armstrong_perceptual_2018}                                                                                   & SADIE II         \\*
   23.
-& Human         & Subject H3                                & 1.2                     &                                                                                                                                            &                  \\
+& Human         & Subject H3                                & 1.2                     &                                                                                                                                            &                  \\*
   24.
@@
   27.
-& Artificial    & FABIAN                                    & 1.7                     & TU Berlin \parencite{brinkmann_high_2017, wierstorf_free_2011}                                                                             & TU Berlin        \\
+& Artificial    & FABIAN                                    & 1.7                     & TU Berlin \parencite{brinkmann_high_2017, wierstorf_free_2011}                                                                             & TU Berlin        \\*
   28.
@@
   29.
-& Artificial    & GRAS 45BB-4 KEMAR - subject A attachment  & 1                       & Aalborg University; University of Iceland \newline \parencite{spagnol_viking_2019,spagnol_viking_2020}                                     & VIKING           \\
+& Artificial    & GRAS 45BB-4 KEMAR - subject A attachment  & 1                       & Aalborg University; University of Iceland \newline \parencite{spagnol_viking_2019,spagnol_viking_2020}                                     & VIKING           \\*
   30.

### 2024-07-02 src/analyze/plot.py

@@ -154,6 +154,8 @@ plt.tight_layout()
 plt.savefig(f"/app/figures/actual_vs_predicted_width.{out_img_format}")
 # plt.show()
 
+# pearsonr(results[0].actual_width, abs(results[0].predicted_width - results[0].actual_width))
+
 
 # %% actual vs predicted location
 plt.figure(figsize=(4, 3))
@@ -179,6 +181,8 @@ plt.legend(['Sample', 'One-to-one line'])
 plt.tight_layout()
 plt.savefig(f"/app/figures/actual_vs_predicted_location.{out_img_format}")
 # plt.show()
+
+# pearsonr(np.abs(results[0].actual_location), abs(results[0].predicted_location - results[0].actual_location))
 
 
 # %% Mean absolute error vs actual width

### 2024-07-02 src/analyze/score_details.py

@@ -1,8 +1,10 @@
 from pathlib import Path
 from results import get_results
 import numpy as np
+from dotmap import DotMap
+import pandas as pd
 
-model_name = "final_width_location"
+model_name = "final_width_location2"
 
 
 def get_stats(results, filter = None) -> tuple[float, float, float, float]:
@@ -13,31 +15,45 @@ def get_stats(results, filter = None) -> tuple[float, float, float, float]:
         error_widths = [result.errors_width for result in results] 
         error_locations = [result.errors_location for result in results]     
     
+    result = DotMap()
+
     error_widths_means = [np.mean(np.abs(err)) for err in error_widths]
-    mean_width = np.mean(error_widths_means)
-    std_width = np.std(error_widths_means)
+    result.min_err_width = np.min(error_widths_means)
+    result.max_err_width = np.max(error_widths_means)
+    result.mean_width = np.mean(error_widths_means)
+    result.std_width = np.std(error_widths_means)
     
     error_locations_means = [np.mean(np.abs(err)) for err in error_locations]
-    mean_location = np.mean(error_locations_means)
-    std_location = np.std(error_locations_means)
+    result.min_err_location = np.min(error_locations_means)
+    result.max_err_location = np.max(error_locations_means)
+    result.mean_location = np.mean(error_locations_means)
+    result.std_location = np.std(error_locations_means)
     
-    return (mean_width, std_width, mean_location, std_location)
+    return result
 
 
 results = get_results(model_name)
 
-mean_width, std_width, mean_location, std_location = get_stats(results)
-print(f"[all] Width MAE: {mean_width:0.2f} (std: {std_width:0.2f})")
-print(f"[all] Location MAE: {mean_location:0.2f} (std: {std_location:0.2f})")
-print(f"[all] Location MAE is better than Width MAE by {(mean_width / mean_location - 1)*100:0.2f}%")
+s = get_stats(results)
+print(f"[all] Width MAE: {s.mean_width:0.2f} (std: {s.std_width:0.2f}, min: {s.min_err_width:0.2f}, max: {s.max_err_width:0.2f})")
[…]

### 2024-07-09 paper/conver.tex
Update final version

\documentclass[a4paper,12pt]{article}

% Paweł Antoniuk \\ Email: pawel.antoniuk@sd.pb.edu.pl \\ Faculty of Computer Science \\ Białystok University of Science \\ Wiejska 45A, 15-351 Białystok \\ Poland
% \signature{Paweł Antoniuk}

\begin{document}

% \begin{tabular}{l @{}}
%     \hfill \today \bigskip                         \\ % Date
%     \hfill XXX                                     \\
%     \hfill Email: xxx@mail.ynu.edu.cn              \\
%     \hfill 2 Cuihu North Road, Wuhua District,     \\
%     \hfill Kunming, Yunnan Province, 650504, China \\ % Address
% \end{tabular}

% # Speech processing and communication systems

Dear Editor,

We are submitting a manuscript for prospective publication in Archives of Acoustics. The title of the paper is ``Estimating Ensemble Location and Width in Binaural Recordings of Music with Convolutional Neural Networks''. The manuscript is authored by Paweł Antoniuk (corresponding author) and Sławomir Krzysztof Zieliński.
\begin{itemize}
    \item We are submitting the paper to Archives of Acoustics as we believe its content fits well within the scope of the journal (the area of speech processing and communication systems).
    \item ‘Speech processing and communication systems’ constitute the field of acoustics covered by Archives of Acoustics, according to the journal website.
    \item We affirm that this manuscript has not been published elsewhere and is not under consideration by another journal.
    \item We suggest the following reviewers:
          \begin{itemize}
              \item 3 pierwszych z cover letter Antoniny
            %   \item Flip jackson profesor
          \end{itemize}
    \item In our opinion the manuscript should be classified as a ``research paper,'' considering the empirical content of the study described.
    \item We are aware that the manuscript exceeds the recommended 20-page limit. This has been caused by the need to describe the methodology at sufficient level of detail, allowing other researchers to replicate our study. Therefore, we believe that the extended content is in this case justified.
\end{itemize}

Thank you in advance for considering the manuscript.

\end{document}

% 1. tabela: signal to noise rate, MAE, korelacja, szum tyko zbior testowy, pink noise, bez modulacji, dla 10 itracji z std dev, np. 7, w db
% 2. prztestowanie modelu dla warunkow echowych

### 2024-07-09 paper/coverletter.tex
Update final version

@@
 We submit our manuscript ``Estimating Ensemble Location and Width in Binaural Recordings of Music with Convolutional Neural Networks'' for publication in Archives of Acoustics.
-Authored by Paweł Antoniuk (corresponding author) and Sławomir Krzysztof Zieliński, this paper presents a method for estimating spatial parameters in binaural audio recordings using a novel ensemble-based approach employing a spectrogram-based deep learning network technique.
+Authored by Paweł Antoniuk (corresponding author) and Sławomir Krzysztof Zieliński, this paper presents a method for estimating spatial parameters in binaural audio recordings using a novel ensemble-oriented approach employing a spectrogram-based deep learning network technique.
 
@@
           \item \textbf{prof. dr hab. inż. Bożena Kostek} \\*
-          Wydział Elektroniki, Telekomunikacji i Informatyki \\*
+          Faculty of Electronics, Telecommunications and Informatics \\*
           Gdańsk University of Technology \\*
@@
           \item \textbf{prof. dr hab. inż. Piotr Kleczkowski} \\*
-          Wydział Inżynierii Mechanicznej i Robotyki \\*
+          Faculty of Mechanical Engineering and Robotics \\*
           AGH University of Krakow \\*
@@
           \item \textbf{dr hab. inż. Grzegorz Szwoch} \\*
-          Wydział Elektroniki, Telekomunikacji i Informatyki \\*
+          Faculty of Electronics, Telecommunications and Informatics \\*
           Gdańsk University of Technology \\*
@@
           Email: \href{mailto:grzegorz.szwoch@pg.edu.pl}{grzegorz.szwoch@pg.edu.pl}
+          \item \textbf{Professor Philip J. B. Jackson} \\*
+          University of Surrey \\*
+          Guildford GU2 7XH \\*
+          United Kingdom \\*
+          Email: \href{mailto:p.jackson@surrey.ac.uk}{p.jackson@surrey.ac.uk}
         \end{itemize}

### 2024-07-09 paper/main.tex
Update final version

@@
 \DeclareFieldFormat{doi}{%
-  \href{https://doi.org/#1}{#1}%
+  doi: \href{https://doi.org/#1}{\nolinkurl{#1}}%
 }
@@
 \author[1,*]{Paweł Antoniuk\orcidlink{0000-0002-0914-8920}}
-\author[1]{Sławomir Zieliński\orcidlink{0000-0002-3205-974X}}
+\author[1]{Sławomir K. Zieliński\orcidlink{0000-0002-3205-974X}}
 \affil[1]{Faculty of Computer Science, Białystok University of Technology \protect\\ Wiejska 45A, 15-351 Białystok, Poland}
@@
 Along with its growing popularity, the quantity of publicly accessible binaural audio recordings has also expanded.
-Consequently, there is now a need for automated and objective measurements of spatial content information, with ensemble location and width being the most prominent.
+Consequently, there is now a need for automated and objective retrieval of spatial content information, with ensemble location and width being the most prominent.
 This study presents a novel method for estimating these ensemble parameters in binaural recordings of music.
@@
 Humans tend to localize groups of sound sources rather than individual ones \parencite{bregman_auditory_1990, rumsey_spatial_2002}.
-Inspired by this fact, the objective of the proposed model is to estimate the location and width of these groups, termed ensembles, instead of the positions of individual sources.
-This study is unique as it not only developed the method but also tested it on a vast, realistic music corpus.
-The corpus comprised 23,040 binaural excerpts synthesized using 192 multi-track music recordings (from a repository provided by \textcite{senior_mixing_2023}) and 30 publicly available head-related transfer functions (HRTFs) from various sources (see Table \ref{table:hrtfs} in \nameref{appendix:a_hrtf} for a detailed list).
+Inspired by this fact, the objective of the proposed model is to estimate the location and width of these groups, termed ``ensembles,'' instead of the positions of individual sources.
+This study is unique as it not only developed the method but also tested it on a relatively large, realistic music corpus.
+The corpus comprised 23,040 binaural excerpts synthesized using 192 multi-track music recordings (from a repository provided by \textcite{senior_mixing_2023}) and 30 sets of publicly available head-related transfer functions (HRTFs) acquired from various sources (see Table \ref{table:hrtfs} in \nameref{appendix:a_hrtf} for a detailed list).
 The music recordings covered many different genres, including rock, jazz, pop, and classical music.
@@
 
-The developed method has the potential to be highly beneficial in automated assessment tasks, where a significant number of binaural recordings must be measured or labeled in terms of their spatial content information.
+The developed method has the potential to be highly beneficial in automated information retrieval tasks, where a significant number of binaural recordings must be analyzed or labeled in terms of their spatial content information.
 This could be utilized in the development of a hypothetical autonomous ``web-crawler bot'' that will collect binaural recordings from publicly accessible repositories and label them according to the spatial properties of the sound sources, such as the location of the music ensemble or the sparsity of audio source positions.
@@
 Furthermore, the tracks of the input multi-track recordings were randomly assigned to sound source positions ($\theta_i$) to enhance the diversity of the final binaural corpora.
-Before the synthesis, the signals in each track were equalized to $-23$ LUFS, in accordance with the ITU-R BS.1770-5 (\citeyear{noauthor_itu-r_2023}) recommendation.
+Before the synthesis, the signals in each track were equalized to $-23$ LKFS, in accordance with the ITU-R BS.1770-5 (\citeyear{noauthor_itu-r_2023}) recommendation.
 
@@
 Although spectrograms do not directly provide information that can be translated into ensemble features, especially ensemble width, the goal of this task was to reduce the number of independent variables compared to the raw audio signal by extracting more compressed and informative data in the frequency domain.
-This step was necessary to decrease the likelihood of overfitting, reduce the number of examples needed to train the model, and thereby lower the overall computational power requirements.
-It is worth mentioning, however, that recent methods have shown that CNNs are suitable for end-to-end audio localization without the spectrogram extraction step, as shown by \textcite{vera-diaz_towards_2018, vecchiotti_end--end_2019}.
+This step was also necessary to decrease the likelihood of overfitting, reducing the number of examples needed to train the model, and thereby lower the overall computational power requirements.
+It is worth mentioning, however, that recently published studies have shown that CNNs are suitable for end-to-end audio localization without the spectrogram extraction step, as demonstrated by \textcite{vera-diaz_towards_2018, vecchiotti_end--end_2019}.
 
@@
 This repetition procedure was employed to ensure more reliable and generalizable results by assessing the model's performance across different subsets of the data.
-Additionally, it helps to account for the inherent variability in neural network training, where slight changes in initial conditions or optimization paths can lead to different model outcomes.
+Additionally, it helped to account for the inherent variability in neural network training, where slight changes in initial conditions or optimization paths can lead to different model outcomes.
 While a large and diverse dataset could mitigate this issue, the binaural excerpts used in this study were generated from only 196 multi-track music recordings.
-This limited source material raised concerns about potential significant variations between the development and test sets in each repetition.
-These concerns were validated, as the maximum observed difference in MAE between repetitions reached up to $0.85\degree$ for ensemble width.
+This limited source material raised concerns by these authors about potential significant variations between the development and test sets in each repetition.
+In hindsight, these concerns were valid, as the maximum observed difference in MAE between repetitions reached up to $0.85\degree$ for ensemble width.
 
-To ensure that the evaluation process was unbiased, this division was done in such a way that no original multi-track recordings used for synthesis were included in both the development and test sets simultaneously.
+To ensure that the evaluation process was unbiased, the data split was done in such a way that no original multi-track recordings used for synthesis were included in both the development and test sets simultaneously.
 However, this rule was not applied to HRTFs databases, allowing for the possibility of HRTF information leaking between the development and test sets.
@@
 The optimizer was configured with the following hyperparameters: an initial learning rate of $10^{-3}$, a decay rate of $10^{-6}$, and momentum parameters $\beta_1 = 0.9$ and $\beta_2 = 0.999$.
-Training was conducted using a batch size of 8.
+Training was conducted using a batch size of 8, with a maximum of 256 epochs set.
 An early stopping technique was implemented to prevent overfitting, terminating the process if no improvement was observed on the validation set for 20 consecutive epochs.
@@
 
-The computational work for this study was conducted on a workstation equipped with an RTX Nvidia GeForce 4090 GPU and a 48-core AMD Ryzen ThreadRipper processor.
+The computational work for this study was conducted on a workstation equipped with an RTX Nvidia GeForce 4090 GPU and a 48-core AMD Ryzen ThreadRipper processor (up to 4.5 GHz).
 On the software side, MATLAB \parencite{MATLAB} with the Audio Toolbox \parencite{MATLAB_Audio_Toolbox} was used for the binaural recording synthesis, while Python \parencite{python} with the SciPy package \parencite{2020SciPy-NMeth} was used for feature extraction and Keras \parencite{chollet_keras_2015} for training the CNN model.
-The complete source code for all stages is publicly available on the GitHub repository \parencite{antoniuk_software_2024}.
+The complete source code for all the experimental stages is publicly available on the GitHub repository \parencite{antoniuk_software_2024}.
 The spectrogram calculation phase required 21 minutes, data partitioning took 34 minutes, and the total training time for all iterations amounted to 40 minutes, making the entire training and evaluation process 95 minutes long.
@@
 Figure \ref{fig:actual_vs_predicted_width} compares the actual and predicted ensemble widths for each sample, showing a heteroscedastic relationship between them, with a slight bias towards predicting lower ensemble width values for higher actual widths.
-This relationship exhibits a strong positive correlation, with a Pearson r of 0.9.
-Additionally, the results indicate that the model provides more precise predictions for narrower ensemble widths, with an average Mean Absolute Error (MAE) of $5.65^\circ$ for $\omega < 30^\circ$.
+This relationship exhibits a strong positive correlation, with a Pearson coefficient $r$ of 0.90.
+Additionally, the results indicate that the model provides more precise predictions for narrower ensemble widths, with an average MAE of $5.65^\circ$ for $\omega < 30^\circ$.
 However, performance deteriorates as the ensemble width increases, resulting in an MAE of $12.44^\circ$ for $\omega > 80^\circ$.
 This effect is more visible in Figure \ref{fig:mae_width}, which highlights the impact of the actual ensemble width on the precision of prediction.
-The correlation between the actual ensemble width and prediction error shows a weak positive relationship, with a Pearson r of 0.27
+The correlation between the actual ensemble width and prediction error shows a weak positive relationship, with a Pearson coefficient $r$ of 0.27.
 
@@
 
-The correlation between actual and predicted ensemble location exhibits a high Pearson r of 0.97, as illustrated in Figure \ref{fig:actual_vs_predicted_location}.
+The correlation between the actual and predicted ensemble location values exhibits a very high degree of correlation, as illustrated in Figure \ref{fig:actual_vs_predicted_location}.
+In this case, Pearson correlation coefficient $r$ is equal to as much as 0.97.
 In contrast to the ensemble width, no significant relationship is observed between actual location and its prediction error.
@@
 Figure \ref{fig:mae_location} corroborates this observation, demonstrating relatively consistent location error across most positions, with minor increases at extreme locations.
-The negligible correlation (Pearson r = -0.03) between absolute location value and prediction error further supports the model's spatial invariance in performance.
+The negligible correlation ($r = -0.03$) between the absolute location value and prediction error further supports the model's spatial invariance in its performance.
 
@@
 
-Figure \ref{fig:map_mae_width} reveals a characteristic depression in $\omega \in [30\degree, 60\degree]$ previously shown from a different perspective in Figure \ref{fig:mae_width}.
+Figure \ref{fig:map_mae_width} reveals a characteristic performance depression in $\omega \in [30\degree, 60\degree]$ previously shown from a different perspective in Figure \ref{fig:mae_width}.
 This heatmap highlights another interesting phenomenon in its upper corners as the error in these areas is considerably higher.
@@
 The ensemble location prediction precision (MAE $=4.76\degree$) of the proposed method can be contextualized with the leading-edge binaural localization DeepEar model introduced by \textcite{yang_deepear_2022}.
-This model reported MAEs of $7.4\degree$ and $2.3\degree$ for multi-source and single-source Angle of Arrival (AoA) estimation, respectively.
+Their model reported MAEs of $7.4\degree$ and $2.3\degree$ for multi-source and single-source Angle of Arrival (AoA) estimation, respectively.
 As another promising example, the WaveLoc-CONV model developed by \textcite{vecchiotti_end--end_2019} demonstrated errors of $0\degree$ in anechoic conditions and $1.7\degree-2.4\degree$ in multi-condition scenarios.
@@
 These limitations, however, present opportunities for future research.
-Despite the current constraints, this study introduces a novel method for describing acoustic scenes in binaural recordings of music, demonstrating substantial potential for advancing binaural audio analysis.
-The method offers promising prospects for developing innovative tools that can objectively measure and analyze large repositories of binaural audio, focusing on spatial content.
+Despite the current constraints, this study introduces a novel method for characterizing acoustic scenes in binaural recordings of music, demonstrating substantial potential for advancing binaural audio analysis.
+The method offers promising prospects for developing innovative tools that can objectively analyze large repositories of binaural audio recordings, focusing on spatial content.
 
@@
 \section*{Acknowledgments}
-The work was supported by the grants from Białystok University of Technology (WI/WI-IiT/3/2022 and WZ/WI-IIT/5/2023) and funded with resources for research by the Ministry of Science and Higher Education in Poland.
+The work was supported by the grants from Białystok University of Technology (WI/WI-IIT/3/2022 and WZ/WI-IIT/5/2023) and funded with resources for research by the Ministry of Science and Higher Education in Poland.
 
